### 2022-06-14 tex/notes.tex
tex template

\documentclass[twoside]{article}

\usepackage{aistats2021}
% If your paper is accepted, change the options for the package
% aistats2021 as follows:
%
%\usepackage[accepted]{aistats2021}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
%\usepackage[round]{natbib}
%\renewcommand{\bibname}{References}
%\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

% Supplementary material: To improve readability, you must use a single-column format for the supplementary material.
\onecolumn
\aistatstitle{Notes}

\section{Transforms}

To prepare a supplementary pdf file, we ask the authors to use \texttt{aistats2021.sty} as a style file and to follow the same formatting instructions as in the main paper.
The only difference is that the supplementary material must be in a \emph{single-column} format.
You can use \texttt{supplement.tex} in our starter pack as a starting point, or append the supplementary content to the main paper and split the final PDF into two separate files.

Note that reviewers are under no obligation to examine your supplementary material.

\section{MISSING PROOFS}

The supplementary materials may contain detailed proofs of the results that are missing in the main paper.

\subsection{Proof of Lemma 3}

\textit{In this section, we present the detailed proof of Lemma 3 and then [ ... ]}

\section{ADDITIONAL EXPERIMENTS}

If you have additional experimental results, you may include them in the supplementary materials.

\subsection{The Effect of Regularization Parameter}

\textit{Our algorithm depends on the regularization parameter $\lambda$. Figure 1 below illustrates the effect of this parameter on the performance of our algorithm. As we can see, [ ... ]}

\vfill

\end{document}

### 2022-06-14 tex/sample_paper.tex
tex template

\documentclass[twoside]{article}

\usepackage{aistats2021}
% If your paper is accepted, change the options for the package
% aistats2021 as follows:
%
%\usepackage[accepted]{aistats2021}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
%\usepackage[round]{natbib}
%\renewcommand{\bibname}{References}
%\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{Transforms for constrained parameters in Hamiltonian Monte Carlo}

\aistatsauthor{ Author 1 \And Author 2 \And  Author 3 }

\aistatsaddress{ Institution 1 \And  Institution 2 \And Institution 3 } ]

\begin{abstract}
  The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
  both left and right-hand margins. Use 10~point type, with a vertical
  spacing of 11~points. The \textbf{Abstract} heading must be centered,
  bold, and in point size 12. Two line spaces precede the
  Abstract. The Abstract must be limited to one paragraph.
\end{abstract}

\section{GENERAL FORMATTING INSTRUCTIONS}

The camera-ready versions of the accepted papers are 8 pages,
plus any additional pages needed for references.

Papers are in 2 columns with the overall line width of 6.75~inches (41~picas).
Each column is 3.25~inches wide (19.5~picas).  The space
between the columns is .25~inches wide (1.5~picas).  The left margin is 0.88~inches (5.28~picas).
Use 10~point type with a vertical spacing of
11~points. Please use US Letter size paper instead of A4.

Paper title is 16~point, caps/lc, bold, centered between 2~horizontal rules.
Top rule is 4~points thick and bottom rule is 1~point thick.
Allow 1/4~inch space above and below title to rules.

Author descriptions are center-justified, initial caps.  The lead
author is to be listed first (left-most), and the Co-authors are set
to follow.  If up to three authors, use a single row of author
descriptions, each one center-justified, and all set side by side;
with more authors or unusually long names or institutions, use more
rows.

Use one-half line space between paragraphs, with no indent.

\section{FIRST LEVEL HEADINGS}

First level headings are all caps, flush left, bold, and in point size
12. Use one line space before the first level heading and one-half line space
after the first level heading.

\subsection{Second Level Heading}

Second level headings are initial caps, flush left, bold, and in point
size 10. Use one line space before the second level heading and one-half line
space after the second level heading.

\subsubsection{Third Level Heading}

Third level headings are flush left, initial caps, bold, and in point
size 10. Use one line space before the third level heading and one-half line
space after the third level heading.

\paragraph{Fourth Level Heading}

Fourth level headings must be flush left, initial caps, bold, and
Roman type.  Use one line space before the fourth level heading, and
place the section text immediately after the heading with no line
break, but an 11 point horizontal space.

%%%
\subsection{Citations, Figure, References}

\subsubsection{Citations in Text}

Citations within the text should include the author's last name and
year, e.g., (Cheesman, 1985). 
%Apart from including the author's last name and year, citations can follow any style, as long as the style is consistent throughout the paper.  
Be sure that the sentence reads
correctly if the citation is deleted: e.g., instead of ``As described
by (Cheesman, 1985), we first frobulate the widgets,'' write ``As
described by Cheesman (1985), we first frobulate the widgets.''

The references listed at the end of the paper can follow any style as long as it is used consistently.

%Be sure to avoid
%accidentally disclosing author identities through citations.

\subsubsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first
  footnote.} in the text. Use 8 point type for footnotes. Place the
footnotes at the bottom of the column in which their markers appear,
continuing to the next column if required. Precede the footnote
section of a column with a 0.5 point horizontal rule 1~inch (6~picas)
long.\footnote{Sample of the second footnote.}

\subsubsection{Figures}

All artwork must be centered, neat, clean, and legible.  All lines
should be very dark for purposes of reproduction, and art work should
not be hand-drawn.  Figures may appear at the top of a column, at the
top of a page spanning multiple columns, inline within a column, or
with text wrapped around them, but the figure number and caption
always appear immediately below the figure.  Leave 2 line spaces
between the figure and the caption. The figure caption is initial caps
and each figure should be numbered consecutively.

Make sure that the figure caption does not get separated from the
figure. Leave extra white space at the bottom of the page rather than
splitting the figure and figure caption.
\begin{figure}[h]
\vspace{.3in}
\centerline{\fbox{This figure intentionally left non-blank}}
\vspace{.3in}
\caption{Sample Figure Caption}
\end{figure}

\subsubsection{Tables}

All tables must be centered, neat, clean, and legible. Do not use hand-drawn tables.
Table number and title always appear above the table.
See Table~\ref{sample-table}.

Use one line space before the table title, one line space after the table title,
and one line space after the table. The table title must be
initial caps and each table numbered consecutively.

\begin{table}[h]
\caption{Sample Table Title} \label{sample-table}
\begin{center}
\begin{tabular}{ll}
\textbf{PART}  &\textbf{DESCRIPTION} \\
\hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

\section{SUPPLEMENTARY MATERIAL}

If you need to include additional appendices during submission, you can include them in the supplementary material file.
You can submit a single file of additional supplementary material which may be either a pdf file (such as proof details) or a zip file for other formats/more files (such as code or videos). 
Note that reviewers are under no obligation to examine your supplementary material. 
If you have only one supplementary pdf file, please upload it as is; otherwise gather everything to the single zip file.

You must use \texttt{aistats2021.sty} as a style file for your supplementary pdf file and follow the same formatting instructions as in the main paper. 
The only difference is that it must be in a \emph{single-column} format.
You can use \texttt{supplement.tex} in our starter pack as a starting point.
Alternatively, you may append the supplementary content to the main paper and split the final PDF into two separate files.

\section{SUBMISSION INSTRUCTIONS}

To submit your paper to AISTATS 2021, please follow these instructions.

\begin{enumerate}
    \item Download \texttt{aistats2021.sty}, \texttt{fancyhdr.sty}, and \texttt{sample\_paper.tex} provided in our starter pack. 
    Please, do not modify the style files as this might result in a formatting violation.

    \item Use \texttt{sample\_paper.tex} as a starting point.
    \item Begin your document with
    \begin{flushleft}
    \texttt{\textbackslash documentclass[twoside]\{article\}}\\
    \texttt{\textbackslash usepackage\{aistats2021\}}
    \end{flushleft}
    The \texttt{twoside} option for the class article allows the
    package \texttt{fancyhdr.sty} to include headings for even and odd
    numbered pages.
    \item When you are ready to submit the manuscript, compile the latex file to obtain the pdf file.
    \item Check that the content of your submission, \emph{excluding} references, is limited to \textbf{8 pages}. The number of pages containing references alone is not limited.
    \item Upload the PDF file along with other supplementary material files to the CMT website.
\end{enumerate}

\subsection{Camera-ready Papers}

%For the camera-ready paper, if you are using \LaTeX, please make sure
%that you follow these instructions.  
% (If you are not using \LaTeX,
%please make sure to achieve the same effect using your chosen
%typesetting package.)

If your papers are accepted, you will need to submit the camera-ready version. Please make sure that you follow these instructions:
\begin{enumerate}
    %\item Download \texttt{fancyhdr.sty} -- the
    %\texttt{aistats2021.sty} file will make use of it.
    \item Change the beginning of your document to
    \begin{flushleft}
    \texttt{\textbackslash documentclass[twoside]\{article\}}\\
    \texttt{\textbackslash usepackage[accepted]\{aistats2021\}}
    \end{flushleft}
    The option \texttt{accepted} for the package
    \texttt{aistats2021.sty} will write a copyright notice at the end of
    the first column of the first page. This option will also print
    headings for the paper.  For the \emph{even} pages, the title of
    the paper will be used as heading and for \emph{odd} pages the
    author names will be used as heading.  If the title of the paper
    is too long or the number of authors is too large, the style will
    print a warning message as heading. If this happens additional
    commands can be used to place as headings shorter versions of the
    title and the author names. This is explained in the next point.
    \item  If you get warning messages as described above, then
    immediately after $\texttt{\textbackslash
    begin\{document\}}$, write
    \begin{flushleft}
    \texttt{\textbackslash runningtitle\{Provide here an alternative
    shorter version of the title of your paper\}}\\
    \texttt{\textbackslash runningauthor\{Provide here the surnames of
    the authors of your paper, all separated by commas\}}
    \end{flushleft}
    Note that the text that appears as argument in \texttt{\textbackslash
      runningtitle} will be printed as a heading in the \emph{even}
    pages. The text that appears as argument in \texttt{\textbackslash
      runningauthor} will be printed as a heading in the \emph{odd}
    pages.  If even the author surnames do not fit, it is acceptable
    to give a subset of author names followed by ``et al.''

    %\item Use the file sample\_paper.tex as an example.

    \item The camera-ready versions of the accepted papers are 8
      pages, plus any additional pages needed for references.

    \item If you need to include additional appendices,
      you can include them in the supplementary
      material file.

    \item Please, do not change the layout given by the above
      instructions and by the style file.

\end{enumerate}

\subsubsection*{Acknowledgements}
All acknowledgments go at the end of the paper, including thanks to reviewers who gave useful comments, to colleagues who contributed to the ideas, and to funding agencies and corporate sponsors that provided financial support. 
To preserve the anonymity, please include acknowledgments \emph{only} in the camera-ready papers.

\subsubsection*{References}

References follow the acknowledgements.  Use an unnumbered third level
heading for the references section.  Please use the same font
size for references as for the body of the paper---remember that
references do not count against your page length total.

\begin{thebibliography}{}
\setlength{\itemindent}{-\leftmargin}
\makeatletter\renewcommand{\@biblabel}[1]{}\makeatother
\bibitem{} J.~Alspector, B.~Gupta, and R.~B.~Allen (1989).
    \newblock Performance of a stochastic learning microchip.
    \newblock In D. S. Touretzky (ed.),
    \textit{Advances in Neural Information Processing Systems 1}, 748--760.
    San Mateo, Calif.: Morgan Kaufmann.

\bibitem{} F.~Rosenblatt (1962).
    \newblock \textit{Principles of Neurodynamics.}
    \newblock Washington, D.C.: Spartan Books.

\bibitem{} G.~Tesauro (1989).
    \newblock Neurogammon wins computer Olympiad.
    \newblock \textit{Neural Computation} \textbf{1}(3):321--323.
\end{thebibliography}

\end{document}

### 2022-06-14 tex/supplement.tex
Add simplex-softmax-augmented correction derivation

--- 2022-06-14 tex/notes.tex
@@
 \usepackage{aistats2021}
+\usepackage{amssymb}
+
 % If your paper is accepted, change the options for the package
@@
 \onecolumn
-\aistatstitle{Notes}
+\aistatstitle{Instructions for Paper Submissions to AISTATS 2021: \\
+Supplementary Materials}
 
-\section{Transforms}
+\section{FORMATTING INSTRUCTIONS}
 
@@
 
-\section{MISSING PROOFS}
+\section{Derivations of density corrections}
 
-The supplementary materials may contain detailed proofs of the results that are missing in the main paper.
+\subsection{Simplex augmented softmax parameterization}
 
-\subsection{Proof of Lemma 3}
+We define the transformation $\phi: \mathbb{R} \backslash \{0\} \to \Delta^{n-1} \times \mathbb{R}_{>0}: z \mapsto (x, r)$,
+where $r = \sum_{i=1}^n e^{z_i}$ and $x_i = \frac{1}{r} e^{z_i}$.
 
-\textit{In this section, we present the detailed proof of Lemma 3 and then [ ... ]}
+First we compute the scalar derivatives:
 
-\section{ADDITIONAL EXPERIMENTS}
+$$
+\begin{aligned}
+    \frac{\mathrm{d} r}{\mathrm{d} z_j} &= e^{z_j} = r x_j\\
+    \frac{\mathrm{d} x_i}{\mathrm{d} z_j} &= \delta_{ij} \frac{1}{r} e^{z_i} - \frac{1}{r^2} e^{z_i} \frac{\mathrm{d} r}{\mathrm{d} z_j} = \delta_{ij} x_i - x_i x_j,
+\end{aligned}
+$$
 
-If you have additional experimental results, you may include them in the supplementary materials.
+where $\delta_{ij} = \begin{cases} 1 &\text{ if } i = j \\ 0 &\text{otherwise}\end{cases}$ is the Kronecker delta.
 
-\subsection{The Effect of Regularization Parameter}
+If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the elements of $x$, then we can write the Jacobian of $z \mapsto \begin{pmatrix}x \\ r \end{pmatrix}$ as
 
-\textit{Our algorithm depends on the regularization parameter $\lambda$. Figure 1 below illustrates the effect of this parameter on the performance of our algorithm. As we can see, [ ... ]}
+$$J = \begin{pmatrix}\mathrm{diag}(x) - x x^\top \\ r x^\top\end{pmatrix} = \begin{pmatrix}I_n - x \boldsymbol{1}_n^\top \\ r \boldsymbol{1}_n^\top \end{pmatrix} \mathrm{diag}(x),$$
+
+where $\boldsymbol{x}_n$ is the $n$-vector of ones.
+The Jacobian has a shape $(n+1,n)$.
+We can compute the density correction due to $\phi^{-1}$ as $\sqrt{|J^\top J}$:
+
+The matrix $P = I - x \boldsymbol{1}^\top_n$ is an orthogonal projection matrix, since if it operates twice on any vector $v$, the second action has no effect:
+
+$$Pv = v - (\boldsymbol{1}^\top_n v) x$$
+$$v - (\boldsymbol{1}^\top_n v) x - x (\boldsymbol{1}^\top_nv - (\boldsymbol{1}^\top_n v) (\boldsymbol{1}^\top_n x)) = v - (\boldsymbol{1}^\top_n v) x,$$
+
+since $\boldsymbol{1}^\top_n x = \sum_{i=1}^n x_i = 1$.
+As a result, $P^2 = P$ and $|P^\top P| = |P^2| = |P|$.
+Then,
+
+$$\sqrt{|J^\top J|} = \sqrt{|\mathrm{diag}(x)(I_n + (r^2 \boldsymbol{1}_n - x) \boldsymbol{1}_n^\top)\mathrm{diag}(x)|} = |\mathrm{diag}(x)|\sqrt{|I_n + (r^2 \boldsymbol{1}_n - x) \boldsymbol{1}_n^\top|}$$
+
+$|\operatorname{diag}(x)| = \prod_{i=1}^n x_i = r^{-n} \exp (\sum_{i=1}^n z_i)$. 
+Using Sylvester's determinant theorem, $|I_n + (r^2 \boldsymbol{1}_n - x) \boldsymbol{1}_n^\top| = 1 + \boldsymbol{1}_n^\top (r^2 \boldsymbol{1}_n - x) = 1 + n r^2 - \sum_{i=1}^n x_i = n r^2$, so the density correction is $\sqrt{n} (\sum_{i=1}^n e^{z_i})^{1-n} \exp(\sum_{i=1}^n z_i)$.
+
+To keep the target distribution proper, we must select a prior distribution for $r$.
+If we choose $r \sim \chi_n$, then the product of the correction and the density of the prior for $r$ is proportional to
+$$\exp\left(\sum_{i=1}^n z_i - \frac{1}{2}(\sum_{i=1}^n e^{z_i})^2\right).$$
 

### 2022-06-14 tex/supplement.tex
Add more details on Jacobian correction

@@
 The Jacobian has a shape $(n+1,n)$.
-We can compute the density correction due to $\phi^{-1}$ as $\sqrt{|J^\top J}$:
+Even though $\phi$ is bijective, the output of $\phi$ is represented with an extra coordinate, so that $J$ is non-square.
+It is possible to construct $\phi$ as the composition of a bijective function and a semi-orthogonal transformation that lifts the output from an $n$-dimensional vector to an $(n+1)$-dimensional vector.
+
+The Jacobian would then factorize as $J = Q H$, where $Q$ is a size $(n+1,n)$ semi-orthogonal matrix so that $Q^\top Q = I_n$.
+Orthogonal transformations leave the volume unchanged, so the density correction depends only on the unknown $H$.
+Then $\sqrt{|J^\top J|} = \sqrt{|H^\top Q^\top Q H|} = \sqrt{|H^\top H|} = |H|$ computes the density correction.
 

### 2022-06-15 simplex-stickbreaking/simplex-stickbreaking.hpp
fix simplex-stickbreaking and add code to evaluate r-hat for mixed chains with pystan and arviz

@@ -1,5 +1,5 @@
 
-// Code generated by stanc e74fdd9
+// Code generated by stanc 2803de5
 #include <stan/model/model_header.hpp>
 namespace simplex_stickbreaking_model_namespace {
 
@@ -8,25 +8,23 @@ using namespace stan::math;
 
 
 stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 18> locations_array__ = 
+static constexpr std::array<const char*, 16> locations_array__ = 
 {" (found before start of program)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 5, column 2 to column 16)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 8, column 2 to column 15)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 9, column 2 to column 16)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 11, column 6 to column 43)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 12, column 6 to column 36)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 10, column 19 to line 13, column 7)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 10, column 2 to line 13, column 7)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 14, column 2 to column 25)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 19, column 6 to column 53)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 18, column 19 to line 20, column 5)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 18, column 4 to line 20, column 5)",
- " (in '/Users/meenaljhajharia/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 17, column 19 to line 21, column 3)",
[…]

### 2022-06-16 tex/supplement.tex
Add softmax simplex derivation

@@
 
+\subsection{Simplex softmax parameterization}
+
+We define the transformation $\phi: \mathbb{R}^{n-1} \to \Delta^{n}: z \mapsto x_-$, where $x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix}$, $x_i = \frac{1}{r} e^{z_i}$ for $i \in \{1, \ldots, n-1\}$, and $r = 1 + \sum_{i=1}^{n-1} e^{z_i}$.
+
+First we compute the scalar derivatives:
+$$
+\begin{aligned}
+    \frac{\mathrm{d} r}{\mathrm{d} z_j} &= e^{z_j} = r x_j\\
+    \frac{\mathrm{d} x_i}{\mathrm{d} z_j} &= \delta_{ij} \frac{1}{r} e^{z_i} - \frac{1}{r^2} e^{z_i} \frac{\mathrm{d} r}{\mathrm{d} z_j} = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
+\end{aligned}
+$$
+
+where $\delta_{ij} = \begin{cases} 1 &\text{if } i = j \\ 0 &\text{otherwise}\end{cases}$ is the Kronecker delta.
+
+If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the elements of $x$, then the Jacobian is
+
+$$J = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top) \operatorname{diag}(x_-),$$
+
+where $\boldsymbol{1}_n$ is the $n$-vector of ones.
+
+Using Sylvester's determinant theorem, $|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = 1 - \boldsymbol{1}_{n-1}^\top x_- = 1 - \sum_{i=1}^{n-1} x_i = x_n$, so
+
+$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} z_i\right) \left(1 + \sum_{i=1}^{n-1} e^{z_i}\right)^{-n}$$
+
 \subsection{Simplex augmented softmax parameterization}

### 2022-06-16 tex/supplement.tex
Refine softmax simplex augmented derivation

@@
 
-We define the transformation $\phi: \mathbb{R} \backslash \{0\} \to \Delta^{n-1} \times \mathbb{R}_{>0}: z \mapsto (x, r)$,
+We define the transformation $\phi: \mathbb{R}^n \backslash \{0\} \to \Delta^{n-1} \times \mathbb{R}_{>0}: z \mapsto (x, r)$,
 where $r = \sum_{i=1}^n e^{z_i}$ and $x_i = \frac{1}{r} e^{z_i}$.
 
+Similar to above, to compute the density correction we use the bijective transformation $f: z \mapsto (x_-, r)$.
 First we compute the scalar derivatives:
@@
 
-where $\delta_{ij} = \begin{cases} 1 &\text{ if } i = j \\ 0 &\text{otherwise}\end{cases}$ is the Kronecker delta.
+which has the Jacobian
 
-If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the elements of $x$, then we can write the Jacobian of $z \mapsto \begin{pmatrix}x \\ r \end{pmatrix}$ as
+$$J = \begin{pmatrix}I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top & -x_- \\ r \boldsymbol{1}_{n-1}^\top & r \end{pmatrix} \mathrm{diag}(x),$$
 
-$$J = \begin{pmatrix}\mathrm{diag}(x) - x x^\top \\ r x^\top\end{pmatrix} = \begin{pmatrix}I_n - x \boldsymbol{1}_n^\top \\ r \boldsymbol{1}_n^\top \end{pmatrix} \mathrm{diag}(x),$$
+For invertible $A$, the determinant of the block matrix $\begin{pmatrix}A & B \\ C & D\end{pmatrix}$ is $|A| |D-CA^{-1}B|$.
+A square matrix is invertible iff its determinant is non-zero.
+From the previous section, $|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = x_n > 0$, so the determinant of the Jacobian is
 
-where $\boldsymbol{x}_n$ is the $n$-vector of ones.
-The Jacobian has a shape $(n+1,n)$.
-Even though $\phi$ is bijective, the output of $\phi$ is represented with an extra coordinate, so that $J$ is non-square.
-It is possible to construct $\phi$ as the composition of a bijective function and a semi-orthogonal transformation that lifts the output from an $n$-dimensional vector to an $(n+1)$-dimensional vector.
+$$|J| = x_n \left|r + r \boldsymbol{1}_{n-1}^\top (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top)^{-1} x_-\right| \prod_{i=1}^n x_i$$
 
-The Jacobian would then factorize as $J = Q H$, where $Q$ is a size $(n+1,n)$ semi-orthogonal matrix so that $Q^\top Q = I_n$.
-Orthogonal transformations leave the volume unchanged, so the density correction depends only on the unknown $H$.
-Then $\sqrt{|J^\top J|} = \sqrt{|H^\top Q^\top Q H|} = \sqrt{|H^\top H|} = |H|$ computes the density correction.
+Let $w = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top)^{-1} x_-$. Then,
 
-The matrix $P = I - x \boldsymbol{1}^\top_n$ is an orthogonal projection matrix, since if it operates twice on any vector $v$, the second action has no effect:
+$$
+\begin{aligned}
+    w - x_- \sum_{i=1}^{n-1} w_i &= x_-\\
+    w &= x_- \left(1 - \sum_{i=1}^{n-1} w_i\right)\\
+    \sum_{i=1}^{n-1} w_i &= \sum_{i=1}^{n-1} \left( x_- (1 - \sum_{i=1}^{n-1} w_i) \right) = \left(\sum_{i=1}^{n-1} x_i \right) \left(1 - \sum_{i=1}^{n-1} w_i\right) = (1 - x_n)  \left(1 - \sum_{i=1}^{n-1} w_i\right)\\
+    \sum_{i=1}^{n-1} w_i &= \frac{1 - x_n}{x_n} = \frac{1}{x_n} - 1\\
+    w &= x_- \left(1 - \left(\frac{1}{x_n} - 1\right)\right) = \frac{1}{x_n} x_-
+\end{aligned}
+$$
 
-$$Pv = v - (\boldsymbol{1}^\top_n v) x$$
-$$v - (\boldsymbol{1}^\top_n v) x - x (\boldsymbol{1}^\top_nv - (\boldsymbol{1}^\top_n v) (\boldsymbol{1}^\top_n x)) = v - (\boldsymbol{1}^\top_n v) x,$$
+Then
 
-since $\boldsymbol{1}^\top_n x = \sum_{i=1}^n x_i = 1$.
-As a result, $P^2 = P$ and $|P^\top P| = |P^2| = |P|$.
-Then,
+$$|J| = x_n r \left|1 + \frac{1}{x_n}\sum_{i=1}^{n-1} x_i\right| \prod_{i=1}^n x_i = r \prod_{i=1}^n x_i$$
 
-$$\sqrt{|J^\top J|} = \sqrt{|\mathrm{diag}(x)(I_n + (r^2 \boldsymbol{1}_n - x) \boldsymbol{1}_n^\top)\mathrm{diag}(x)|} = |\mathrm{diag}(x)|\sqrt{|I_n + (r^2 \boldsymbol{1}_n - x) \boldsymbol{1}_n^\top|}$$
+To keep the target distribution proper, we must select a prior distribution $\pi(r)$ for $r$.
+If we choose $r \sim \chi_n$, then the product of the correction and the density of the prior for $r$ is proportional to
 
-$|\operatorname{diag}(x)| = \prod_{i=1}^n x_i = r^{-n} \exp (\sum_{i=1}^n z_i)$. 
-Using Sylvester's determinant theorem, $|I_n + (r^2 \boldsymbol{1}_n - x) \boldsymbol{1}_n^\top| = 1 + \boldsymbol{1}_n^\top (r^2 \boldsymbol{1}_n - x) = 1 + n r^2 - \sum_{i=1}^n x_i = n r^2$, so the density correction is $\sqrt{n} (\sum_{i=1}^n e^{z_i})^{1-n} \exp(\sum_{i=1}^n z_i)$.
-
-To keep the target distribution proper, we must select a prior distribution for $r$.
-If we choose $r \sim \chi_n$, then the product of the correction and the density of the prior for $r$ is proportional to
-$$\exp\left(\sum_{i=1}^n z_i - \frac{1}{2}(\sum_{i=1}^n e^{z_i})^2\right).$$
+$$\mathrm{correction} = \pi(r) |J| = r^n e^{-r^2/2} \prod_{i=1}^n x_i = \exp\left(\sum_{i=1}^n z_i - \frac{1}{2}\left(\sum_{i=1}^n e^{z_i}\right)^2\right).$$
 

### 2022-06-16 tex/notes.tex
update softmax code

@@
 
-\section{Transforms}
-
-To prepare a supplementary pdf file, we ask the authors to use \texttt{aistats2021.sty} as a style file and to follow the same formatting instructions as in the main paper.
-The only difference is that the supplementary material must be in a \emph{single-column} format.
-You can use \texttt{supplement.tex} in our starter pack as a starting point, or append the supplementary content to the main paper and split the final PDF into two separate files.
-
-Note that reviewers are under no obligation to examine your supplementary material.
-
-\section{MISSING PROOFS}
-
-The supplementary materials may contain detailed proofs of the results that are missing in the main paper.
-
-\subsection{Proof of Lemma 3}
-
-\textit{In this section, we present the detailed proof of Lemma 3 and then [ ... ]}
-
-\section{ADDITIONAL EXPERIMENTS}
-
-If you have additional experimental results, you may include them in the supplementary materials.
-
-\subsection{The Effect of Regularization Parameter}
-
-\textit{Our algorithm depends on the regularization parameter $\lambda$. Figure 1 below illustrates the effect of this parameter on the performance of our algorithm. As we can see, [ ... ]}
-
-\vfill
+\section{Softmax Transform}
+\textit{Roughly writing the algebra for the idea that Bob Proposed}
 

### 2022-06-16 simplex-softmax/simplex-softmax.hpp

@@ -1,5 +1,5 @@
 
-// Code generated by stanc e74fdd9
+// Code generated by stanc 2803de5
 #include <stan/model/model_header.hpp>
 namespace simplex_softmax_model_namespace {
 
@@ -8,14 +8,15 @@ using namespace stan::math;
 
 
 stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 7> locations_array__ = 
+static constexpr std::array<const char*, 8> locations_array__ = 
 {" (found before start of program)",
- " (in '/Users/meenaljhajharia/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 1 to column 23)",
- " (in '/Users/meenaljhajharia/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 1 to column 54)",
- " (in '/Users/meenaljhajharia/transforms/simplex-softmax/simplex-softmax.stan', line 11, column 1 to column 64)",
- " (in '/Users/meenaljhajharia/transforms/simplex-softmax/simplex-softmax.stan', line 2, column 1 to column 16)",
- " (in '/Users/meenaljhajharia/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 8 to column 11)",
- " (in '/Users/meenaljhajharia/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 9 to column 10)"};
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 1 to column 19)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 1 to column 45)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 11, column 1 to column 32)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 12, column 1 to column 39)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 2, column 1 to column 16)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 8 to column 11)",
[…]

### 2022-06-16 tex/supplement.tex
Use y symbol for latent space

@@
 
-We define the transformation $\phi: \mathbb{R}^{n-1} \to \Delta^{n}: z \mapsto x_-$, where $x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix}$, $x_i = \frac{1}{r} e^{z_i}$ for $i \in \{1, \ldots, n-1\}$, and $r = 1 + \sum_{i=1}^{n-1} e^{z_i}$.
+We define the transformation $\phi: \mathbb{R}^{n-1} \to \Delta^{n}: y \mapsto x_-$, where $x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix}$, $x_i = \frac{1}{r} e^{y_i}$ for $i \in \{1, \ldots, n-1\}$, and $r = 1 + \sum_{i=1}^{n-1} e^{y_i}$.
 
@@
 \begin{aligned}
-    \frac{\mathrm{d} r}{\mathrm{d} z_j} &= e^{z_j} = r x_j\\
-    \frac{\mathrm{d} x_i}{\mathrm{d} z_j} &= \delta_{ij} \frac{1}{r} e^{z_i} - \frac{1}{r^2} e^{z_i} \frac{\mathrm{d} r}{\mathrm{d} z_j} = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
+    \frac{\mathrm{d} r}{\mathrm{d} y_j} &= e^{y_j} = r x_j\\
+    \frac{\mathrm{d} x_i}{\mathrm{d} y_j} &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i} \frac{\mathrm{d} r}{\mathrm{d} y_j} = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
 \end{aligned}
@@
 
-$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} z_i\right) \left(1 + \sum_{i=1}^{n-1} e^{z_i}\right)^{-n}$$
+$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} y_i\right) \left(1 + \sum_{i=1}^{n-1} e^{y_i}\right)^{-n}$$
 
@@
 
-We define the transformation $\phi: \mathbb{R}^n \backslash \{0\} \to \Delta^{n-1} \times \mathbb{R}_{>0}: z \mapsto (x, r)$,
-where $r = \sum_{i=1}^n e^{z_i}$ and $x_i = \frac{1}{r} e^{z_i}$.
+We define the transformation $\phi: \mathbb{R}^n \backslash \{0\} \to \Delta^{n-1} \times \mathbb{R}_{>0}: y \mapsto (x, r)$,
+where $r = \sum_{i=1}^n e^{y_i}$ and $x_i = \frac{1}{r} e^{y_i}$.
 
-Similar to above, to compute the density correction we use the bijective transformation $f: z \mapsto (x_-, r)$.
+Similar to above, to compute the density correction we use the bijective transformation $f: y \mapsto (x_-, r)$.
 First we compute the scalar derivatives:
@@
 \begin{aligned}
-    \frac{\mathrm{d} r}{\mathrm{d} z_j} &= e^{z_j} = r x_j\\
-    \frac{\mathrm{d} x_i}{\mathrm{d} z_j} &= \delta_{ij} \frac{1}{r} e^{z_i} - \frac{1}{r^2} e^{z_i} \frac{\mathrm{d} r}{\mathrm{d} z_j} = \delta_{ij} x_i - x_i x_j,
+    \frac{\mathrm{d} r}{\mathrm{d} y_j} &= e^{y_j} = r x_j\\
+    \frac{\mathrm{d} x_i}{\mathrm{d} y_j} &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i} \frac{\mathrm{d} r}{\mathrm{d} y_j} = \delta_{ij} x_i - x_i x_j,
 \end{aligned}
@@
 For invertible $A$, the determinant of the block matrix $\begin{pmatrix}A & B \\ C & D\end{pmatrix}$ is $|A| |D-CA^{-1}B|$.
-A square matrix is invertible iff its determinant is non-zero.
+A square matrix is invertible iff its determinant is non-yero.
 From the previous section, $|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = x_n > 0$, so the determinant of the Jacobian is
@@
 
-$$\mathrm{correction} = \pi(r) |J| = r^n e^{-r^2/2} \prod_{i=1}^n x_i = \exp\left(\sum_{i=1}^n z_i - \frac{1}{2}\left(\sum_{i=1}^n e^{z_i}\right)^2\right).$$
+$$\mathrm{correction} = \pi(r) |J| = r^n e^{-r^2/2} \prod_{i=1}^n x_i = \exp\left(\sum_{i=1}^n y_i - \frac{1}{2}\left(\sum_{i=1}^n e^{y_i}\right)^2\right).$$
 

### 2022-06-16 tex/supplement.tex
Add alternative choice of r prior
Notational improvements

@@
 
-We define the transformation $\phi: \mathbb{R}^{n-1} \to \Delta^{n}: y \mapsto x_-$, where $x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix}$, $x_i = \frac{1}{r} e^{y_i}$ for $i \in \{1, \ldots, n-1\}$, and $r = 1 + \sum_{i=1}^{n-1} e^{y_i}$.
+Let $\Delta^n$ indicate the unit $n$-simplex with $n$ positive elements that sum to 1 and $\Delta^n_-$ indicate the first $n-1$ elements, from which the final element can be uniquely determined.
+
+We define the transformation $\phi: \mathbb{R}^{n-1} \to \Delta^n_-: y \mapsto x_-$, where $x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix} \in \Delta^n$, $x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$, and $r = 1 + \sum_{i=1}^{n-1} e^{y_i}$.
 
@@
 
-We define the transformation $\phi: \mathbb{R}^n \backslash \{0\} \to \Delta^{n-1} \times \mathbb{R}_{>0}: y \mapsto (x, r)$,
-where $r = \sum_{i=1}^n e^{y_i}$ and $x_i = \frac{1}{r} e^{y_i}$.
+We define the transformation $\phi: \mathbb{R}^n \to \Delta^{n-1} \times \mathbb{R}_{>0}: y \mapsto (x_-, r)$,
+where $r = \sum_{i=1}^n e^{y_i}$ and $x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$..
 
-Similar to above, to compute the density correction we use the bijective transformation $f: y \mapsto (x_-, r)$.
 First we compute the scalar derivatives:
@@
 
-which has the Jacobian
+which corresponds to the Jacobian
 
-$$J = \begin{pmatrix}I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top & -x_- \\ r \boldsymbol{1}_{n-1}^\top & r \end{pmatrix} \mathrm{diag}(x),$$
+$$J = \begin{pmatrix}I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top & -x_- \\ r \boldsymbol{1}_{n-1}^\top & r \end{pmatrix} \mathrm{diag}(x).$$
 
@@
 
+Alternatively, if we choose $r \sim \mathrm{Gamma}(n, 1)$, then
+
+$$\mathrm{correction} = \pi(r) |J| = r^n e^{-r} \prod_{i=1}^n x_i = \exp\left(\sum_{i=1}^n y_i - \sum_{i=1}^n e^{y_i}\right).$$
+
+This latter correction is equivalent to the sampling procedure from the Dirichlet distribution with $\alpha_i=1$, where $z_i \sim \mathrm{Exponential}(1)$ and $y = \frac{z}{\sum_{i=1}^n z_i}$.
+
 \vfill

### 2022-06-27 tex/notes.tex
just pushing everything

@@
 \usepackage{aistats2021}
+\usepackage{amsmath}
 % If your paper is accepted, change the options for the package
@@
 
+\section{Introduction}
+
+\section{Constraints}
+\subsection{Bounded Constraints}
+\subsection{Simplex Constraints}
+\subsection{Definite Matrix constraints}
 \section{Softmax Transform}
-\textit{Roughly writing the algebra for the idea that Bob Proposed}
+
+The softmax function can be understood from Multinomial Logistic Regression employed for predicting probabilities of a Categorically distributed variable.
+Geometrically it maps $R^K$ to the boundary of a unit K-simplex(it is simply the convex hull of $k+1$ affinely independent points in $R^K$.
+Essentially it transforms a vector of size K to another vector of size K where the outputs sum to 1.
+It is worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so when a vector of size $K$ is transformed the $K_th$ vector is simply $1$- sumofk-1 vectors
+R
 

### 2022-06-27 simplex-softmax-augmented/simplex-softmax-augmented.hpp

@@ -1,26 +1,30 @@
 
 // Code generated by stanc 2803de5
 #include <stan/model/model_header.hpp>
-namespace simplex_stan_model_namespace {
+namespace simplex_softmax_augmented_model_namespace {
 
 using stan::model::model_base_crtp;
 using namespace stan::math;
 
 
 stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 7> locations_array__ = 
+static constexpr std::array<const char*, 11> locations_array__ = 
 {" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 6, column 2 to column 15)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 9, column 2 to column 39)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 2, column 2 to column 17)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 3, column 18 to column 19)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 3, column 2 to column 27)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 6, column 10 to column 11)"};
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax-augmented/simplex-softmax-augmented.stan', line 6, column 1 to column 13)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax-augmented/simplex-softmax-augmented.stan', line 9, column 1 to column 37)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax-augmented/simplex-softmax-augmented.stan', line 10, column 1 to column 30)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax-augmented/simplex-softmax-augmented.stan', line 13, column 1 to column 38)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax-augmented/simplex-softmax-augmented.stan', line 14, column 1 to column 38)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax-augmented/simplex-softmax-augmented.stan', line 2, column 1 to column 16)",
[…]

### 2022-06-27 simplex-softmax/simplex-softmax.hpp

@@ -8,15 +8,18 @@ using namespace stan::math;
 
 
 stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 8> locations_array__ = 
+static constexpr std::array<const char*, 11> locations_array__ = 
 {" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 1 to column 19)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 1 to column 45)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 11, column 1 to column 32)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 12, column 1 to column 39)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 2, column 1 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 8 to column 11)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 9 to column 10)"};
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 7, column 2 to column 20)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 11, column 2 to column 46)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 15, column 2 to column 33)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 16, column 2 to column 40)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 17, column 2 to column 39)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 2, column 2 to column 17)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 3, column 18 to column 19)",
[…]

### 2022-06-27 simplex-stickbreaking/simplex-stickbreaking.hpp

@@ -8,23 +8,26 @@ using namespace stan::math;
 
 
 stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 16> locations_array__ = 
+static constexpr std::array<const char*, 19> locations_array__ = 
 {" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 5, column 2 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 8, column 2 to column 15)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 9, column 2 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 11, column 6 to column 40)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 12, column 6 to column 36)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 10, column 19 to line 13, column 7)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 10, column 2 to line 13, column 7)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 14, column 2 to column 25)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 18, column 6 to column 63)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 17, column 19 to line 19, column 5)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 17, column 2 to line 19, column 5)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 6, column 2 to column 16)",
[…]

### 2022-07-03 simplex/softmax-augmented/dirichlet_symmetric.hpp
add scripts

@@ -1,99 +1,107 @@
 
 // Code generated by stanc 2803de5
 #include <stan/model/model_header.hpp>
-namespace simplex_stickbreaking_model_namespace {
+namespace dirichlet_symmetric_model_namespace {
 
 using stan::model::model_base_crtp;
 using namespace stan::math;
 
 
 stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 16> locations_array__ = 
+static constexpr std::array<const char*, 11> locations_array__ = 
 {" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 5, column 2 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 8, column 2 to column 15)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 9, column 2 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 11, column 6 to column 40)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 12, column 6 to column 36)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 10, column 19 to line 13, column 7)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 10, column 2 to line 13, column 7)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 14, column 2 to column 25)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 18, column 6 to column 63)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 17, column 19 to line 19, column 5)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 17, column 2 to line 19, column 5)",
[…]

### 2022-07-03 simplex/softmax/dirichlet_symmetric.hpp

@@ -1,69 +1,95 @@
 
 // Code generated by stanc 2803de5
 #include <stan/model/model_header.hpp>
-namespace simplex_stan_model_namespace {
+namespace dirichlet_symmetric_model_namespace {
 
 using stan::model::model_base_crtp;
 using namespace stan::math;
 
 
 stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 4> locations_array__ = 
+static constexpr std::array<const char*, 8> locations_array__ = 
 {" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 5, column 2 to column 15)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 2, column 2 to column 17)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 5, column 10 to column 11)"};
+<<<<<<< HEAD:simplex-softmax/simplex-softmax.hpp
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 1 to column 19)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 1 to column 45)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 11, column 1 to column 32)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 12, column 1 to column 39)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 2, column 1 to column 16)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 8 to column 11)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 9 to column 10)"};
+=======
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/softmax/dirichlet_symmetric.stan', line 7, column 2 to column 20)",
[…]

### 2022-07-03 simplex/stan/dirichlet_symmetric.hpp

@@ -1,0 +1,359 @@
+
+// Code generated by stanc 2803de5
+#include <stan/model/model_header.hpp>
+namespace dirichlet_symmetric_model_namespace {
+
+using stan::model::model_base_crtp;
+using namespace stan::math;
+
+
+stan::math::profile_map profiles__;
+static constexpr std::array<const char*, 4> locations_array__ = 
+{" (found before start of program)",
+<<<<<<< HEAD:simplex-stan/simplex-stan.hpp
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 5, column 2 to column 15)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 2, column 2 to column 17)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 5, column 10 to column 11)"};
+=======
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stan/dirichlet_symmetric.stan', line 6, column 2 to column 15)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stan/dirichlet_symmetric.stan', line 9, column 2 to column 39)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stan/dirichlet_symmetric.stan', line 2, column 2 to column 17)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stan/dirichlet_symmetric.stan', line 3, column 18 to column 19)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stan/dirichlet_symmetric.stan', line 3, column 2 to column 27)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stan/dirichlet_symmetric.stan', line 6, column 10 to column 11)"};
+>>>>>>> 51ecb2d (added scripts):simplex/stan/dirichlet_symmetric.hpp
+
+
+
+
+class dirichlet_symmetric_model final : public model_base_crtp<dirichlet_symmetric_model> {
+
+ private:
+  int K; 
+  
+ 
+ public:
+  ~dirichlet_symmetric_model() { }
+  
+  inline std::string model_name() const final { return "dirichlet_symmetric_model"; }
+
+  inline std::vector<std::string> model_compile_info() const noexcept {
[…]

### 2022-07-03 simplex/stickbreaking/dirichlet.hpp

@@ -1,0 +1,519 @@
+
+// Code generated by stanc 2803de5
+#include <stan/model/model_header.hpp>
+namespace dirichlet_model_namespace {
+
+using stan::model::model_base_crtp;
+using namespace stan::math;
+
+
+stan::math::profile_map profiles__;
+static constexpr std::array<const char*, 19> locations_array__ = 
+{" (found before start of program)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 6, column 2 to column 16)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 9, column 2 to column 15)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 10, column 2 to column 16)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 12, column 6 to column 40)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 13, column 6 to column 36)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 11, column 19 to line 14, column 7)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 11, column 2 to line 14, column 7)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 15, column 2 to column 25)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 19, column 6 to column 63)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 18, column 19 to line 20, column 5)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 18, column 2 to line 20, column 5)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex/stickbreaking/dirichlet.stan', line 21, column 2 to column 39)",
[…]

### 2022-07-03 simplex/stickbreaking/dirichlet_symmetric.hpp

@@ -1,82 +1,120 @@
 
 // Code generated by stanc 2803de5
 #include <stan/model/model_header.hpp>
-namespace simplex_softmax_model_namespace {
+namespace dirichlet_symmetric_model_namespace {
 
 using stan::model::model_base_crtp;
 using namespace stan::math;
 
 
 stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 8> locations_array__ = 
+static constexpr std::array<const char*, 16> locations_array__ = 
 {" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 1 to column 19)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 1 to column 45)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 11, column 1 to column 32)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 12, column 1 to column 39)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 2, column 1 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 8 to column 11)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 9 to column 10)"};
+<<<<<<< HEAD:simplex-stickbreaking/simplex-stickbreaking.hpp
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 5, column 2 to column 16)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 8, column 2 to column 15)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 9, column 2 to column 16)",
+ " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 11, column 6 to column 40)",
[…]

### 2022-07-03 utils.py

@@ -1,0 +1,128 @@
+import pandas as pd
+import numpy as np
+import os
+import json
+import sys
+import pickle
+from cmdstanpy import CmdStanModel
+import argparse
+import arviz as az
+from pathlib import Path
+from tqdm import tqdm
+
+
+
+def sample(transform_category, transform, evaluating_model, parameters, output_file=None, 
+                auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 1, show_progress = True):
+    '''
+    Sample from the given dictionary containing the models, transform_categories, and parameters.
+    It saves the results in a json file which is named with n_repeat, n_chains and n_iter.
+
+    Parameters
+    ----------
+    transform_category: str
+        The transform category to use.
+    
+    transform: str
+        The transform name to use.
+    
+    evaluating_model: str
+        The model to use.
+    
+    parameters: list of dict
+        List of Dictionaries containing the parameters to use.
+        e.g. parameters = [{'alpha': [0.1]*10, 'K': 10}, {'alpha': [0.1]*20, 'K': 20}]
+    
+    output_file: str
+        Custom filename.
+
+    auto_eval_all_params: bool
+        If True, all parameterizations are evaluated.
+
+    n_repeat: int
+        Number of times to repeat the sampling.
+
+    n_iter : int
+        Number of samples to be drawn.
+
+    n_chains : int
+        Number of chains to be drawn.
+
+    show_progress: bool (default = True)
+        Whether to show progress bar.
+
+    Note
+    ----
+    The directory structure for storing stan files is: transform_category/transform/model_name.stan
+    The directory structure for storing the results is: sampling_results/transform_category/transform/model_name/{n_iter}_iters_{n_chains}_chains_{n_repeat}_repeats.json
+    '''
+    if auto_eval_all_params:
+        with open(f'{transform_category}/{evaluating_model}.json', 'rb') as f:
+            parameters = pickle.load(f)
+    
+    with open(f'{transform_category}/parameters.json', 'rb') as f:
+        param_map = pickle.load(f)
+
+    path=f'sampling_results/{transform_category}/{transform}/{evaluating_model}'
+    os.makedirs(path, exist_ok=True)
+
+    for params in tqdm(parameters):
[…]

### 2022-07-03 tex/sample_paper.tex
refactor code and restructure the directories
Update readme.md

Update readme.md

restructure

refactor code and restructure directories

some changes

some changes

trying to amke stuff work

still figuring it out

@@
 \usepackage{aistats2021}
+\usepackage{cite}
+
 % If your paper is accepted, change the options for the package
@@
 
-\section{GENERAL FORMATTING INSTRUCTIONS}
+\section{Introduction}
 
-The camera-ready versions of the accepted papers are 8 pages,
-plus any additional pages needed for references.
+Sampling algorithms play an important role in Statistics, in this case we refer to drawing samples from a probability density or distribution (not survey sampling).
+Monte Carlo Markov chain methods are commonly used for high dimensional sampling.
+In this paper, we look at Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose samples for the Metropolis algorithm.
+Bayesian Inference often uses parameters with constraints, but HMC optimizes on the unconstrained space.
+It is difficult to use Monte Carlo estimation on a constrained domain ~\cite{neal2008optimal}
 
-Papers are in 2 columns with the overall line width of 6.75~inches (41~picas).
-Each column is 3.25~inches wide (19.5~picas).  The space
-between the columns is .25~inches wide (1.5~picas).  The left margin is 0.88~inches (5.28~picas).
-Use 10~point type with a vertical spacing of
-11~points. Please use US Letter size paper instead of A4.
+\section{Transforms}
 
-Paper title is 16~point, caps/lc, bold, centered between 2~horizontal rules.
-Top rule is 4~points thick and bottom rule is 1~point thick.
-Allow 1/4~inch space above and below title to rules.
+\subsection{Simplex Transforms}
 
-Author descriptions are center-justified, initial caps.  The lead
-author is to be listed first (left-most), and the Co-authors are set
-to follow.  If up to three authors, use a single row of author
-descriptions, each one center-justified, and all set side by side;
-with more authors or unusually long names or institutions, use more
-rows.
-
-Use one-half line space between paragraphs, with no indent.
-
-\section{FIRST LEVEL HEADINGS}
-
-First level headings are all caps, flush left, bold, and in point size
-12. Use one line space before the first level heading and one-half line space
-after the first level heading.
-
-\subsection{Second Level Heading}
-
-Second level headings are initial caps, flush left, bold, and in point
-size 10. Use one line space before the second level heading and one-half line
-space after the second level heading.
+\subsubsection{Softmax Transform}
+The softmax function can be understood from Multinomial Logistic Regression employed for predicting probabilities of a Categorically distributed variable.
+Geometrically it maps $R^K$ to the boundary of a unit K-simplex(it is simply the convex hull of $k+1$ affinely independent points in $R^K$.
+Essentially it transforms a vector of size K to another vector of size K where the outputs sum to 1.
+It is worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so when a vector of size $K$ is transformed the $K_th$ vector is simply $1$- sumofk-1 vectors
 
@@
 
-\begin{thebibliography}{}
-\setlength{\itemindent}{-\leftmargin}
-\makeatletter\renewcommand{\@biblabel}[1]{}\makeatother
-\bibitem{} J.~Alspector, B.~Gupta, and R.~B.~Allen (1989).
-    \newblock Performance of a stochastic learning microchip.
-    \newblock In D. S. Touretzky (ed.),
-    \textit{Advances in Neural Information Processing Systems 1}, 748--760.
-    San Mateo, Calif.: Morgan Kaufmann.
-
-\bibitem{} F.~Rosenblatt (1962).
-    \newblock \textit{Principles of Neurodynamics.}
-    \newblock Washington, D.C.: Spartan Books.
-
-\bibitem{} G.~Tesauro (1989).
-    \newblock Neurogammon wins computer Olympiad.
-    \newblock \textit{Neural Computation} \textbf{1}(3):321--323.
-\end{thebibliography}
+\bibliography{mybib}{}
+\bibliographystyle{plain}
 

### 2022-07-03 simplex-softmax/simplex-softmax.hpp

@@ -1,402 +1,0 @@
-
-// Code generated by stanc 2803de5
-#include <stan/model/model_header.hpp>
-namespace simplex_softmax_model_namespace {
-
-using stan::model::model_base_crtp;
-using namespace stan::math;
-
-
-stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 8> locations_array__ = 
-{" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 1 to column 19)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 1 to column 45)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 11, column 1 to column 32)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 12, column 1 to column 39)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 2, column 1 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 5, column 8 to column 11)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-softmax/simplex-softmax.stan', line 8, column 9 to column 10)"};
-
-
-
-
-class simplex_softmax_model final : public model_base_crtp<simplex_softmax_model> {
-
- private:
-  int K;
-  int x_unc_1dim__; 
-  
- 
- public:
-  ~simplex_softmax_model() { }
-  
-  inline std::string model_name() const final { return "simplex_softmax_model"; }
-
-  inline std::vector<std::string> model_compile_info() const noexcept {
-    return std::vector<std::string>{"stanc_version = stanc3 2803de5", "stancflags = "};
-  }
-  
-  
-  simplex_softmax_model(stan::io::var_context& context__,
-                        unsigned int random_seed__ = 0,
-                        std::ostream* pstream__ = nullptr) : model_base_crtp(0) {
-    int current_statement__ = 0;
-    using local_scalar_t__ = double ;
-    boost::ecuyer1988 base_rng__ = 
-        stan::services::util::create_rng(random_seed__, 0);
-    (void) base_rng__;  // suppress unused var warning
[…]

### 2022-07-03 simplex-stan/simplex-stan.hpp

@@ -1,349 +1,0 @@
-
-// Code generated by stanc 2803de5
-#include <stan/model/model_header.hpp>
-namespace simplex_stan_model_namespace {
-
-using stan::model::model_base_crtp;
-using namespace stan::math;
-
-
-stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 4> locations_array__ = 
-{" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 5, column 2 to column 15)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 2, column 2 to column 17)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stan/simplex-stan.stan', line 5, column 10 to column 11)"};
-
-
-
-
-class simplex_stan_model final : public model_base_crtp<simplex_stan_model> {
-
- private:
-  int K; 
-  
- 
- public:
-  ~simplex_stan_model() { }
-  
-  inline std::string model_name() const final { return "simplex_stan_model"; }
-
-  inline std::vector<std::string> model_compile_info() const noexcept {
-    return std::vector<std::string>{"stanc_version = stanc3 2803de5", "stancflags = "};
-  }
-  
-  
-  simplex_stan_model(stan::io::var_context& context__,
-                     unsigned int random_seed__ = 0,
-                     std::ostream* pstream__ = nullptr) : model_base_crtp(0) {
-    int current_statement__ = 0;
-    using local_scalar_t__ = double ;
-    boost::ecuyer1988 base_rng__ = 
-        stan::services::util::create_rng(random_seed__, 0);
-    (void) base_rng__;  // suppress unused var warning
-    static constexpr const char* function__ = "simplex_stan_model_namespace::simplex_stan_model";
-    (void) function__;  // suppress unused var warning
-    local_scalar_t__ DUMMY_VAR__(std::numeric_limits<double>::quiet_NaN());
-    (void) DUMMY_VAR__;  // suppress unused var warning
-    try {
-      int pos__ = std::numeric_limits<int>::min();
-      pos__ = 1;
-      current_statement__ = 2;
-      context__.validate_dims("data initialization","K","int",
-           std::vector<size_t>{});
-      K = std::numeric_limits<int>::min();
-      
-      
-      current_statement__ = 2;
[…]

### 2022-07-03 simplex-stickbreaking/simplex-stickbreaking.hpp

@@ -1,486 +1,0 @@
-
-// Code generated by stanc 2803de5
-#include <stan/model/model_header.hpp>
-namespace simplex_stickbreaking_model_namespace {
-
-using stan::model::model_base_crtp;
-using namespace stan::math;
-
-
-stan::math::profile_map profiles__;
-static constexpr std::array<const char*, 16> locations_array__ = 
-{" (found before start of program)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 5, column 2 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 8, column 2 to column 15)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 9, column 2 to column 16)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 11, column 6 to column 40)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 12, column 6 to column 36)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 10, column 19 to line 13, column 7)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 10, column 2 to line 13, column 7)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 14, column 2 to column 25)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 18, column 6 to column 63)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 17, column 19 to line 19, column 5)",
- " (in '/Users/meenaljhajharia/cmdstan/transforms/simplex-stickbreaking/simplex-stickbreaking.stan', line 17, column 2 to line 19, column 5)",
[…]

### 2022-07-03 utils/dirichlet_symmetric_utils.py

@@ -1,0 +1,24 @@
+import numpy as np
+from utils import sample, retrieve, cumulative_mean, rmse, rmse_leapfrog
+
+def get_dirichlet_symmetric_params():
+    alphas=[0.1,1,10]
+    Ks = [10,100,1000]
+    parameters = []
+    for alpha in alphas:
+        for K in Ks:
+            parameters.append({'alpha': [alpha]*K, 'K': K})
+            
+    return parameters
+
+def get_dirichlet_symmetric_rmse(transform, parameters, n_repeat=1):
+    
+    idata = retrieve(transform_category="simplex", transform=transform, n_repeat=n_repeat,
+                 evaluating_model="dirichlet_symmetric", parameters=parameters)
+    
+    alpha = np.asarray(parameters['alpha'])
+    true_var = alpha/sum(alpha)
+    var_name='x'
+    var_dim=0
+    
+    return rmse_leapfrog(idata, true_var, var_name, var_dim) 

### 2022-07-03 utils/utils.py

@@ -1,0 +1,130 @@
+import pandas as pd
+import numpy as np
+import os
+import json
+import sys
+import pickle
+from cmdstanpy import CmdStanModel
+import argparse
+import arviz as az
+from pathlib import Path
+from tqdm import tqdm
+
+
+
+def sample(transform_category, transform, evaluating_model, parameters, output_file=None, 
+                auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 1, show_progress = True):
+    '''
+    Sample from the given dictionary containing the models, transform_categories, and parameters.
+    It saves the results in a json file which is named with n_repeat, n_chains and n_iter.
+
+    Parameters
+    ----------
+    transform_category: str
+        The transform category to use.
+    
+    transform: str
+        The transform name to use.
+    
+    evaluating_model: str
+        The model to use.
+    
+    parameters: list of dict
+        List of Dictionaries containing the parameters to use.
+        e.g. parameters = [{'alpha': [0.1]*10, 'K': 10}, {'alpha': [0.1]*20, 'K': 20}]
+    
+    output_file: str
+        Custom filename.
+
+    auto_eval_all_params: bool
+        If True, all parameterizations are evaluated.
+
+    n_repeat: int
+        Number of times to repeat the sampling.
+
+    n_iter : int
+        Number of samples to be drawn.
+
+    n_chains : int
+        Number of chains to be drawn.
+
+    show_progress: bool (default = True)
+        Whether to show progress bar.
+
+    Note
+    ----
+    The directory structure for storing stan files is: tramsforms/transform_category/transform/evaluating_model.stan
+    The directory structure for storing the results is: sampling_results/transform_category/transform/evaluating_model/*.json
+    '''
+    if auto_eval_all_params:
+        with open(f'transforms/{transform_category}/{evaluating_model}_parameters.json', 'rb') as f:
+            parameters = pickle.load(f)
+    
+    with open(f'transforms/{transform_category}/dirichlet_symmetric_map.json', 'rb') as f:
+        param_map = pickle.load(f)
+
+    path=f'sampling_results/{transform_category}/{transform}/{evaluating_model}'
+    os.makedirs(path, exist_ok=True)
+    
+    if type(parameters)==dict:
+        parameters = [parameters]
+    for params in tqdm(parameters):
[…]

### 2022-07-06 tex/sample_paper.tex
updates

@@
 \usepackage{cite}
+\usepackage{amsmath}
+\usepackage{amsfonts}
 
@@
 
-\subsection{Simplex Transforms}
+\subsection{Simplex}
 
+A simplex is a set of non negative vectors that sum up to a constant $a$, this can be written as:
+\begin{equation*}
+\Delta=\left\{\mathbf{x} \in \mathbb{R}^{n} \mid\left\langle\mathbf{x}, \mathbf{1}_{n}\right\rangle=a, \mathbf{x} \geq 0\right\}
+\end{equation*}
+
+In computational works, we generally care about the unit simplex, this is the special case where $ a = 1$.
+Geometrically, an \textit{n-simplex} is the convex closure of $n+1$ affinely independent set of points $x_0, x_1, \dots x_n$.
+It is the simplest polytope for $n$-dimensions(in this case we will be looking at a simplex as a convex polytope instead of abstract).
+An $n-simplex$ can be represented in the barycentric coordinate system with weighted vertices, as follows:
+\begin{equation*}
+\Delta = \left\{x_{0} u_{0}+\cdots+x_{n} u_{n} \mid \sum_{i=0}^{n} x_{i}=1\\
+ \text{ and } x_{i} \geq 0 \text { for } i=0, \ldots, k\right\}
+\end{equation*}
+
+With this representation, taking $u_{i}=1$, the $n$ unit simplex can be written as: 
+\begin{equation*}
+\left\{x \in \mathbb{R}^{n}: x_{0}+\cdots+x_{n-1}=1, x_{i} \geq 0 \text { for } i=0, \ldots, k-1\right\}
+\end{equation*}
+
+A \textit{unit simplex} on $R^n$ is an $n - 1$ dimensional simplex with unit vectors as vertices. Notice the unit weights, this set only has $n-1$ affinely independent points.
+
+Sum to one constraints are useful in probabilistic inference, often used for modeling categorical responses.
+The rest of the section describes various transforms used to map a simplex to an unconstrained space ($R^n$).
 \subsubsection{Softmax Transform}
@@
 Essentially it transforms a vector of size K to another vector of size K where the outputs sum to 1.
-It is worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so when a vector of size $K$ is transformed the $K_th$ vector is simply $1$- sumofk-1 vectors
+It is worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so when a vector of size $K$ is transformed the $K_th$ vector is simply $1$- sumof k-1 vectors
 

### 2022-07-06 utils/script.py

@@ -1,0 +1,26 @@
+from utils import sample, retrieve, cumulative_mean, rmse
+from dirichlet_symmetric_utils import get_dirichlet_symmetric_rmse, get_dirichlet_symmetric_params
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os
+os.chdir('..')
+
+transforms = ['stan', 'stickbreaking', 'softmax', 'softmax-augmented']
+for transform in transforms:
+    sample(transform_category="simplex", transform=transform, evaluating_model="dirichlet_symmetric", 
+           parameters={}, auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 100)
+
+
+from utils import sample, retrieve, cumulative_mean, rmse
+from dirichlet_symmetric_utils import get_dirichlet_symmetric_rmse, get_dirichlet_symmetric_params
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os
+os.chdir('..')
+
+sample(transform_category="simplex", transform='stickbreaking', evaluating_model="dirichlet_symmetric", 
+        parameters={}, auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 100)

### 2022-07-06 tex/paper.tex
reorganized paper and directories

\documentclass[11pt]{article}

\usepackage[letterpaper, total={6.5in, 9in}]{geometry}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathpazo}
\usepackage{sourcecodepro}

\newcommand{\setcomp}[2]{\left\{ #1 \ \Big|\ #2 \right\}}
\newcommand{\rngto}[1]{1{:}#1}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\absdet}[1]{\abs{#1}}

\begin{document}

\title{Efficient Unconstraining
  Parameter Transforms for Hamiltonian Monte Carlo}
\author{Meenal Jhajaria \\ \small Flatiron Institute \and Seth Axen \\
  \small University of T\"ubingen \and Bob
  Carpenter \\ \small Flatiron Institute}
\date{DRAFT: \today}
\maketitle

\begin{abstract}
  \noindent
  This paper evaluates the the statistical and computational
  efficiency of unconstraining parameter transforms for Hamiltonian
  Monte Carlo sampling.
\end{abstract}

\section{Introduction}

In statistical computing, we often need to compute high-dimensional
integrals over densities $\pi(x)$ (e.g., Bayesian estimation or
prediction, $p$-value calcuations, etc.).  The only black-box
techniques that work for general high-dimensional integrals are
Markov chain Monte Carlo (MCMC) methods.  The most effective MCMC
method in high dimensions is Hamiltonian Monte Carlo (HMC).  HMC works
by simulating the Hamiltonian dynamics of a fictitious particle
representing the value being sampled coupled with a momentum term.

Although it's possible to write HMC samplers that work for simply
constrained values such as upper- and/or lower-bounds
\cite{neal2011mcmc} or unit vectors \cite{byrne2013geodesic}, it is
much more challenging to do the same for complex constraints such as
simplexes or positive definite matrices or for densities involving
multiple constrained values.  Instead, it is far more common to map
the constrained values to unconstrained values before sampling
\cite{}.  This presents the issue of selecting which transform to use
among an infinite set of options, which is the topic of this paper.

Sampling algorithms play an important role in Statistics, in this case
we refer to drawing samples from a probability density or distribution
(not survey sampling). Monte Carlo Markov chain methods are commonly
used for high dimensional sampling. In this paper, we look at
Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
samples for the Metropolis algorithm. Bayesian Inference often uses
parameters with constraints, but HMC optimizes on the unconstrained
space. It is difficult to use Monte Carlo estimation on a constrained
domain ~\cite{neal2008optimal}

\section{Changes of Variables}

If $X$ is a random variable with density $p_X$ and $Y = f(X)$ for a
smooth and bijective function $f$, then
\[
  p_Y(y) = p_X(f^{-1}(y)) \absdet{J_{f^{-1}}(y)},
\]
where the Jacobian of the inverse transform is defined by
\[
  J_{f^{-1}}(y) = \frac{\partial}{\partial y} \, f^{-1}(y)
\]
and
\[
  \absdet{J_{f^{-1}}(y)}
  = \abs{\det J_{f^{-1}}(y)}.
\]

\section{Unit simplex}

A unit $N$-simplex is an $N + 1$-dimensional vector of non-negative
values that sums to one.  As such, there are only $N$ degrees of
freedom, because if $x$ is an $N$-simplex, then
\[
  x_N = -(x_1 + x_2 + \cdots + x_{N-1}).
\]
Simplexes are useful for representations of multinomial probabilities
(e.g., probabilities of categories in a classification problem).

The set of unit $N$-simplexes is conventionally denoted
\[
  \Delta^N = \setcomp{x \in \mathbb{R}^{N + 1}}{\textrm{sum}(x) = 1
    \textrm{ and }
    x_n \geq 0 \textrm{ for } n \in \rngto{N}}
\]
Geometrically, an $N$-simplex is the convex closure of $N+1$ points
that are 1 in one coordinate and 0 elsewhere.  For example, the
3-simplex is the complex closure of
$\begin{bmatrix}1 & 0 & 0 \end{bmatrix},
\begin{bmatrix} 0 & 1 & 0 \end{bmatrix}$,
and $\begin{bmatrix} 0 & 0 & 1 \end{bmatrix}$.

\subsection{Additive log ratio transform}

The unconstraining transform for the identified softmax is known as
the additive log ratio (ALR) transform \cite{aitchison???}, which is a
bijection $\textrm{alr}:\Delta^{N-1} \rightarrow \mathbb{R}^{N-1}$ defined by
\[
  \textrm{alr}(x)
  = \begin{bmatrix}\displaystyle
    \log \frac{x_1}{x_N} \cdots \log \frac{x_{N-1}}{x_N}
  \end{bmatrix}.
\]
The inverse additive log ratio transform maps values in $\mathbb{R}^{N-1}$ to
$\Delta^{N-1}$ by
\[
  \textrm{alr}^{-1}(x) = \textrm{softmax}(\begin{bmatrix}x &  0\end{bmatrix}),
\]
where for $u \in \mathbb{R}^N$,
\[
  \textrm{softmax}(u) = \frac{\exp(u)}{\textrm{sum}(\exp(u))}.
\]
To calculate the change of variables adjustment, we consider only the
first $N-1$ coordinates of the result, because the last is defined in
terms of the first $N-1$.  The function operating on the first $N-1$
coordinates is $s:\mathbb{R}^{N-1} \rightarrow \mathbb{R}^{N-1}$, defined by
\[
  s(x) = \frac{\exp(x)}{\exp(x) + 1}.
\]
The change of variables is then defined so that if $p_X(x)$ is a
density over simplices $x \in \Delta^{N-1}$, then we inverse transform
to an unconstrained density defined over $Y \in R^{N-1}$ by
\[
  p_Y(y) = p_X(\textrm{alr}^{-1}(y)) \absdet{J_{\textrm{alr}^{-1}}}.
\]

\subsubsection{Softmax Transform}

The softmax function can be understood from Multinomial Logistic
Regression employed for predicting probabilities of a Categorically
distributed variable. Geometrically it maps $R^K$ to the boundary of a
unit K-simplex(it is simply the convex hull of $k+1$ affinely
independent points in $R^K$. Essentially it transforms a vector of
size K to another vector of size K where the outputs sum to 1. It is
worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
when a vector of size $K$ is transformed the $K_th$ vector is simply
$1$- sumof k-1 vectors

\subsection{Simplex softmax parameterization}

Let $\Delta^n$ indicate the unit $n$-simplex with $n$ positive
elements that sum to 1 and $\Delta^n_-$ indicate the first $n-1$
elements, from which the final element can be uniquely determined.

We define the transformation
$\phi: \mathbb{R}^{n-1} \to \Delta^n_-: y \mapsto x_-$, where
$x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix} \in \Delta^n$,
$x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$, and
$r = 1 + \sum_{i=1}^{n-1} e^{y_i}$.

First we compute the scalar derivatives:
\[
\begin{aligned}
  \frac{\mathrm{d} r}{\mathrm{d} y_j}
  &= e^{y_j} = r x_j\\
  \frac{\mathrm{d} x_i}{\mathrm{d} y_j}
  &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i}
  \frac{\mathrm{d} r}{\mathrm{d} y_j}
  = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
\end{aligned}
\]
where
$\delta_{ij} = \begin{cases} 1 &\text{if } i = j \\ 0
  &\text{otherwise}\end{cases}$ is the Kronecker delta.

If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the
elements of $x$, then the Jacobian is
\[
  J = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top) \operatorname{diag}(x_-),
\]
where $\boldsymbol{1}_n$ is the $n$-vector of ones.

Using Sylvester's determinant theorem,
$|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = 1 -
\boldsymbol{1}_{n-1}^\top x_- = 1 - \sum_{i=1}^{n-1} x_i = x_n$, so
$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} y_i\right) \left(1 + \sum_{i=1}^{n-1} e^{y_i}\right)^{-n}$$

\subsection{Simplex augmented softmax parameterization}

We define the transformation
$\phi: \mathbb{R}^n \to \Delta^{n-1} \times \mathbb{R}_{>0}: y \mapsto
(x_-, r)$, where $r = \sum_{i=1}^n e^{y_i}$ and
$x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$..

First we compute the scalar derivatives:
\[
\begin{aligned}
  \frac{\mathrm{d} r}{\mathrm{d} y_j}
  &= e^{y_j} = r x_j
  \\
  \frac{\mathrm{d} x_i}{\mathrm{d} y_j}
  &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i} \frac{\mathrm{d} r}{\mathrm{d} y_j} = \delta_{ij} x_i - x_i x_j,
\end{aligned}
\]
which corresponds to the Jacobian
\[
  J = \begin{pmatrix}I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top & -x_- \\
    r \boldsymbol{1}_{n-1}^\top & r \end{pmatrix} \mathrm{diag}(x).
\]

For invertible $A$, the determinant of the block matrix
$\begin{pmatrix}A & B \\ C & D\end{pmatrix}$ is $|A| |D-CA^{-1}B|$.  A
square matrix is invertible iff its determinant is non-yero.  From the
previous section,
\[
  |I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = x_n > 0,
\]
so the determinant of the Jacobian is
\[
  |J| = x_n \left|r + r \boldsymbol{1}_{n-1}^\top (I_{n-1} - x_-
    \boldsymbol{1}_{n-1}^\top)^{-1} x_-\right|
  \prod_{i=1}^n x_i.
\]

Let $w = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top)^{-1} x_-$. Then,
\[
\begin{aligned}
    w - x_- \sum_{i=1}^{n-1} w_i &= x_-\\
    w &= x_- \left(1 - \sum_{i=1}^{n-1} w_i\right)\\
    \sum_{i=1}^{n-1} w_i &= \sum_{i=1}^{n-1} \left( x_- (1 - \sum_{i=1}^{n-1} w_i) \right) = \left(\sum_{i=1}^{n-1} x_i \right) \left(1 - \sum_{i=1}^{n-1} w_i\right) = (1 - x_n)  \left(1 - \sum_{i=1}^{n-1} w_i\right)\\
    \sum_{i=1}^{n-1} w_i &= \frac{1 - x_n}{x_n} = \frac{1}{x_n} - 1\\
    w &= x_- \left(1 - \left(\frac{1}{x_n} - 1\right)\right) = \frac{1}{x_n} x_-
  \end{aligned}
\]

Then
\[
  |J| = x_n r \left|1 + \frac{1}{x_n}\sum_{i=1}^{n-1} x_i\right|
  \prod_{i=1}^n x_i = r \prod_{i=1}^n x_i
\]

To keep the target distribution proper, we must select a prior
distribution $\pi(r)$ for $r$.  If we choose $r \sim \chi_n$, then the
product of the correction and the density of the prior for $r$ is
proportional to

\[
  \mathrm{correction}
  = \pi(r) |J| = r^n e^{-r^2/2} \prod_{i=1}^n x_i
  = \exp\left(\sum_{i=1}^n y_i - \frac{1}{2}\left(\sum_{i=1}^n
      e^{y_i}\right)^2\right).
\]

Alternatively, if we choose $r \sim \mathrm{Gamma}(n, 1)$, then
\[
  \mathrm{correction} = \pi(r) |J| = r^n e^{-r} \prod_{i=1}^n x_i =
  \exp\left(\sum_{i=1}^n y_i - \sum_{i=1}^n e^{y_i}\right).
\]
This latter correction is equivalent to the sampling procedure from
the Dirichlet distribution with $\alpha_i=1$, where
$z_i \sim \mathrm{Exponential}(1)$ and
$y = \frac{z}{\sum_{i=1}^n z_i}$.

Both of these corrections can be captured with the generalization
\[
  \mathrm{correction}
  = \pi(r) |J|
  = r^n e^{-r^p/p} \prod_{i=1}^n x_i
  = \exp\left(\sum_{i=1}^n y_i - \frac{1}{p} \left(\sum_{i=1}^n e^{y_i}\right)^p\right),
\]
for $p > 0$, which corresponds to $r \sim \text{Generalized-Gamma}(1, n, p)$.

\subsubsection*{Acknowledgements}

We would like to thank $\ldots$.

\bibliography{all}{}
\bibliographystyle{plain}

\end{document}

### 2022-07-07 tex/paper.tex
First draft of theory of transforms

@@
 \]
+
+\section{Transforms}
+
+To draw samples from a distribution $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.
+
+Let $f\colon \mathcal{Y} \to \mathcal{X}$ be a smooth and continuous map.
+Only when $m = n$ can $f$ be bijective.
+When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = f(y)$.
+In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
+We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
+$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
+that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
+When $h$ is non-bijective at single points, we call these singularities.
+While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.
+
+Given a proper joint distribution with a smooth and continuous density $p_{X,Y}(x, y | \theta)$ for some parameters $\theta$, we can use the usual change of variables formula to write
+
+\[
+  p_Y(y | \theta) = p_{X,Z}(f(y), g(y) | \theta) |J_h(y)|.
+\]
+
+To sample from a target density $p_X(x | \theta)$, let $p_{X,Z}(x, z | \theta) = p_X(x | \theta) p_Z(z | x, \theta)$.
+Then we must choose some proper distribution $p_Z$ that is smooth and continuous.
+Given that choice, we find that
+\[
+  p_Y(y | \theta) = p_X(f(y) | \theta) p_Z(g(y) | f(y), \theta) |J_h(y)|.
+\]
+
+Note that when $m > n$, it is insufficient to specify a transform $h$.
+One must also select a prior $p_Z$, and it is desirable to establish some general properties that a given pair $h$ and $p_Z$ should have.
+In addition to the above properties, the optimal pair would
+1. discard as few degrees of freedom $|m - n|$ as possible.
+2. produce a $p_Z$ and $|J_y|$ and their gradients that are efficient to compute.
+3. produce a $p_Y$ that is easy to sample from.
+
+We can expand a bit on the latter point.
+As a general rule, the more normal $p_Y$ is, the easier it is to sample from (REF?).
+Additionally, near singularities, small perturbations to $y$ tend to result in larger changes to $h(y)$ than far from the singularity.
+This warps the volume around the singularity, which causes $|J_y(y)|$ to smoothly approach infinity near the singularity.
+If $m$ is large, this is generally not a problem, as the "curse of dimensionality" works in our favor, and there is exponentially more volume away from the singularity than near the singularity.
+However, for low $m$, it is necessary that $p_{X,Z}$ goes to 0 near the singularity faster than $|J_y(y)|$ goes to infinity.
+While $p_X$ is generally fixed by the user-specified model and may not have this necessary property, the user may then choose a $p_Z$ that has this property.
 

### 2022-07-08 utils/ess_plot.py
scripts for ess/rmse plot and rhat

@@ -1,0 +1,58 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+from cmdstanpy import CmdStanModel
+import arviz as az
+import os
+import numpy as np
+from tqdm import tqdm
+import logging
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.stats import gaussian_kde
+os.chdir('..')
+logger = logging.getLogger('cmdstanpy')
+logger.addHandler(logging.NullHandler())
+
+def get_ess_leapfrog_plot(transforms=['stickbreaking', 'softmax', 'softmax-augmented'], 
+                          params={'alpha':[0.1]*10,'N':10}, 
+                          var_name='x', var_dim=0, repeat=100, plot_format='matplotlib', fig_name='figure.png'):
+   
+    def get_ess_leapfrog_ratio(transform, params={'alpha':[0.1]*10,'N':10}, 
+                          var_name='x', var_dim=0, repeat=100):
+        model = CmdStanModel(stan_file = f'transforms/simplex/{transform}/dirichlet_symmetric.stan', 
+                         cpp_options = {'STAN_THREADS':'true'})
+        x=[]
+        for i in tqdm(range(repeat)):
+            idata = az.from_cmdstanpy(model.sample(data=params,show_progress=False ))
+            data=idata.sel(var_names=[str(var_name)], x_dim_0=var_dim)
+            ess = az.ess(data, var_names=[str(var_name)], method="bulk")[str(var_name)].values
+            leapfrog = idata.sample_stats['n_steps'].sum().values
+            x.append(ess/leapfrog)
+        return x
+
+    data={}
+    for transform in transforms:
+        data[str(transform)] = get_ess_leapfrog_ratio(transform,repeat=100)
+        
+    if plot_format=='matplotlib':
+        fig, ax = plt.subplots()
+        for transform in transforms:
+            kde = gaussian_kde(data[str(transform)])
+            dist_space = np.linspace( min(data[str(transform)]), max(data[str(transform)]), 1000 )
+            ax.plot( dist_space, kde(dist_space), label=transform)
+        ax.set_title(str(params))
+        ax.axes.yaxis.set_ticklabels([])
+        plt.xlabel('ESS/Leapfrog')
+        plt.ylabel('Density')
+        plt.legend()
+        plt.savefig(str(fig_name), dpi=300)
+        plt.show()
+
+    if plot_format=='arviz':
[…]

### 2022-07-08 utils/rhat_eval.py

@@ -1,0 +1,65 @@
+import cmdstanpy
+import pandas as pd
+import arviz as az
+import numpy as np
+import matplotlib.pyplot as plt
+import xarray as xr
+import os
+from cmdstanpy import cmdstan_path, CmdStanModel
+
+
+simplex_stan = os.path.join('simplex-stan/simplex-stan.stan')
+simplex_stickbreaking = os.path.join('simplex-stickbreaking/simplex-stickbreaking.stan')
+simplex_softmax = os.path.join('simplex-softmax/simplex-softmax.stan')
+
+def calc_rhat_mixed_chains(path_1, path_2, variables, data, force_compile=False):
+    
+    """
+    This function evaluates rhat for 4 chains from 2 different models, as in 2 chains from each.
+    
+    Parameters
+    ----------
+    
+    path_1: str
+        path for the first model
+        
+    path_2: str
+        path for the second model
+    
+    variables: list
+        Variables to evaluate rhat for
+    
+    data: dict
+        Dictionary of data
+    
+    force_compile: Bool
+        Whether to recompile or not
+    """
+    variables = ' '.join(variables)
+    #Build stan model
+    file_1 = os.path.join(path_1)
+    file_2 = os.path.join(path_2)
+    model_1 = CmdStanModel(stan_file=file_1, cpp_options={'STAN_THREADS':'true'})
+    model_2 = CmdStanModel(stan_file=file_2, cpp_options={'STAN_THREADS':'true'})
+    
+    #Recompile
+    if force_compile is True:
+        model_1.compile(force=True)
+        model_2.compile(force=True)
+    
+    #Fit stan model
+    fit_1 = model_1.sample(data=dict(K=10))
+    fit_2 = model_2.sample(data=dict(K=10))
+    
+    #Convert to idata for arviz
+    idata_1 = az.from_cmdstanpy(fit_1)
+    idata_2 = az.from_cmdstanpy(fit_2)
+    
+    #Stack the samples according to chains
+    stacked_1 = az.extract_dataset(idata_1)
+    stacked_2 = az.extract_dataset(idata_2)
+    
+    #Concatenate chains and evaluate rhat
+    chains = np.concatenate((stacked_1.sel(chain=1)[str(variables)], stacked_1.sel(chain=2)[str(variables)], stacked_2.sel(chain=1)[str(variables)], stacked_2.sel(chain=2)[str(variables)]), axis=1)
+    rhat = az.rhat(chains, var_names=variables,method="rank")
+    return rhat

### 2022-07-08 utils/rmse_plot.py
rmse_plot file was empty

@@ -1,0 +1,28 @@
+import os
+os.chdir('transforms')
+In [43]:
+from utils import sample, retrieve, cumulative_mean, rmse
+from dirichlet_symmetric_utils import get_dirichlet_symmetric_rmse, get_dirichlet_symmetric_params
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+
+transforms = ['stan', 'stickbreaking', 'softmax', 'softmax-augmented']
+for transform in transforms:
+    sample(transform_category="simplex", transform=transform, evaluating_model="dirichlet_symmetric", 
+           parameters={}, auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 100)
+plt.rcParams["figure.figsize"] = (20,10)
+
+fig, axes = plt.subplots(3,3)
+
+fig.supxlabel('Cumulative Leapfrog Steps')
+fig.supylabel('Root Mean Squared Error')
+
+for ax, params in zip(axes.flatten(),  get_dirichlet_symmetric_params()):
+    for transform in transforms:
+        x, y = get_dirichlet_symmetric_rmse(transform, params)
+        ax.plot(x,y, label = str(transform))
+    ax.set_title(f'alpha={params["alpha"][0]}, K={params["K"]}')
+
+fig.legend(labels=transforms,bbox_to_anchor = (0.6, -0.05));
+fig.savefig(str('figure.png'))

### 2022-07-08 tex/paper.tex
Jacobian determinant for ALR transform

@@
 
-\usepackage[letterpaper, total={6.5in, 9in}]{geometry}
-\usepackage{cite}
 \usepackage{amsmath}
 \usepackage{amsfonts}
+\usepackage{cite}
+\usepackage[letterpaper, total={6.5in, 9in}]{geometry}
 \usepackage{mathpazo}
 \usepackage{sourcecodepro}
+
+\usepackage{hyperref}
 
@@
 The unconstraining transform for the identified softmax is known as
-the additive log ratio (ALR) transform \cite{aitchison???}, which is a
-bijection $\textrm{alr}:\Delta^{N-1} \rightarrow \mathbb{R}^{N-1}$ defined by
+the additive log ratio (ALR) transform
+\cite{aitchison1982statistical}, which is a bijection
+$\textrm{alr}:\Delta^{N-1} \rightarrow \mathbb{R}^{N-1}$ defined for
+$x \in \Delta^{N-1}$ by
 \[
@@
 \]
-The inverse additive log ratio transform maps values in $\mathbb{R}^{N-1}$ to
-$\Delta^{N-1}$ by
+
+The inverse additive log ratio transform maps values in
+$\mathbb{R}^{N-1}$ to $\Delta^{N-1}$ defined for $y \in
+\mathbb{R}^{N-1}$ by
 \[
-  \textrm{alr}^{-1}(x) = \textrm{softmax}(\begin{bmatrix}x &  0\end{bmatrix}),
+  \textrm{alr}^{-1}(y)
+  = \textrm{softmax}(\begin{bmatrix} y &  0 \end{bmatrix}),
 \]
@@
 \]
+
 To calculate the change of variables adjustment, we consider only the
 first $N-1$ coordinates of the result, because the last is defined in
-terms of the first $N-1$.  The function operating on the first $N-1$
-coordinates is $s:\mathbb{R}^{N-1} \rightarrow \mathbb{R}^{N-1}$, defined by
+terms of the first $N-1$.  For convenience, we define a function 
+$s:\mathbb{R}^{N-1} \rightarrow \mathbb{R}^{N-1}$ that operates on
+$N-1$ unconstrained variables and returns the first $N-1$ components
+of the $\textrm{alr}^{-1}(y)$, which is
+defined for $y \in \mathbb{R}^{N-1}$ by
 \[
-  s(x) = \frac{\exp(x)}{\exp(x) + 1}.
+  s(y) = \frac{\exp(y)}{\textrm{sum}(\exp(y)) + 1}
+  = \begin{bmatrix}
+    \textrm{alr}^{-1}(y)_1
+    & \cdots &
+    \textrm{alr}^{-1}(y)_{N-1}
+    \end{bmatrix}.
 \]
-The change of variables is then defined so that if $p_X(x)$ is a
-density over simplices $x \in \Delta^{N-1}$, then we inverse transform
-to an unconstrained density defined over $Y \in R^{N-1}$ by
+Given a density $p_X(x)$ defined over simplices $x \in \Delta^{N-1}$,
+we can transform to a density over unconstrained parameters $y \in
+\mathbb{R}^{N-1}$ by applying the inverse ALR transform and adjusting
+for the change of variables, which yields
 \[
-  p_Y(y) = p_X(\textrm{alr}^{-1}(y)) \absdet{J_{\textrm{alr}^{-1}}}.
+  p_Y(y) = p_X(\textrm{alr}^{-1}(y)) \absdet{J_{s}(y)},
+\]
+where $J_{s}(y)$ is the Jacobian of the function $s$ evaluated at $y$
+and $\absdet{J_s(y)}$ is the absolute value of its determinant.
+
+To calculate the determinant of the Jacobian of the inverse transform,
+we start by noting that $s = \textrm{exp} \circ \textrm{norm}$, where
+$\textrm{exp}$ is the elementwise exponential function and
+\textrm{norm} is defined by
+\[
+  \textrm{norm}(z) = \frac{z}{\textrm{sum}(z) + 1}.
+\]
+As such, the resulting Jacobian determinant is the product of the
+Jacobian determinants of the component functions,
+\[
+  \absdet{J_s(y)}
+  = \absdet{J_{\textrm{exp}}(y)} \absdet{J_{\textrm{norm}}(z)},
+\]
+where $z = \textrm{exp}(y)$.  The Jacobian for the exponential
+function is diagonal, so the determinant is the product of the
+diagonal of the Jacobian, which for $y \in \mathbb{R}^{N-1}$ is
+\[
+  \absdet{J_{\textrm{exp}}(y)} = \textrm{prod}(\exp(y)).
+\]
+As above, let $z = \exp(y) \in (0, \inf)^{N-1}$.  We can differentiate
+$\textrm{norm}$ to derive the Jacobian,
+\[
+  J_{\textrm{norm}}
+  = \frac{1}{1 + \textrm{sum}(z)} \mathbb{I}_{N-1}
+  - \left(\frac{1}{(1 + \textrm{sum}(z))^2} \beta \right)
+  \textrm{vector}_{N-1}(1)^{\top},
+\]
+where $\mathbb{I}_{N-1}$ is the $(N - 1) \times (N - 1)$ unit matrix and
+$\textrm{vector}_{N-1}(1)$ is the $N - 1$-vector with values 1.  Using
+the matrix determinant lemma,\footnote{The matrix determinant lemma
+  is \[\textrm{det}(A + u v^{\top}) = (1 + v^{\top} A^{-1} u)
+    \textrm{det}(A).\]}
+we have
+\begin{eqnarray*}
+  \textrm{absdet}(J_{\textrm{norm}}(z))
+  & = &
+  \left(
+    1
+    + \textrm{vector}_{N-1}(1)^{\top}
+    \left(\frac{1}{1 + \textrm{sum}(z)} \mathbb{I} \right)^{-1}
+    \frac{-z}{(1 + \textrm{sum}(z))^2}
+    \right)
+    \ \textrm{det}\left(\frac{1}{1 + \textrm{sum}(z)} \mathbb{I}
+        \right)
+  \\[6pt]
+  & = &
+  \left(
+    1 
+    + \textrm{sum}\left( \frac{-(1 + \textrm{sum}(z)) z}{(1 +
+        \textrm{sum}(z))^2} \right)
+  \right)
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
+  & = &
+        \left(1 + \textrm{sum}\left(\frac{-z}{1 + \textrm{sum}(z)} \right)\right)        
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
+  & = & \left( 1 - \textrm{sum}(\textrm{norm}(z)) \right) 
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
+  & = & \left( \frac{1}{1 + \textrm{sum}(z)} \right)^N.
+\end{eqnarray*}
+Thus the entire absolute determinant of the Jacobian is defined by the
+product, 
+\[
+  \absdet{J_s(y)}
+  \ = \
+  \textrm{prod}(\exp(y))
+  \, \left( \frac{1}{1 + \textrm{sum}(\exp(y))} \right)^N.
+\]
+and our final expression for densities for unconstrained $y \in
+\mathbb{R}^{N-1}$ is
+\[
+  p_Y(y)
+  = p_X(\textrm{alr}^{-1}(y))
+  \, \textrm{prod}(\exp(y))
+  \left( \frac{1}{1 + \textrm{sum}(\exp(y))} \right)^N
 \]
 
-\subsubsection{Softmax Transform}
+\subsection{Softmax Transform}
 
@@
 
-We would like to thank $\ldots$.
+We would like to thank \url{matrixcalculus.org} for providing an
+easy-to-use symbolic matrix derivative calculator.
 

### 2022-07-11 tex/paper.tex
add stickbreaking transform

--- 2022-07-07 tex/paper.tex
@@
 
+\subsection{StickBreaking Transform}
+
+The StickBreaking transform ican be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}.
+Intuitively, this comprises of recursively breaking a piece $x_i$ from a stick of unit length, where the leftover stick in the $i^{th}$ iteration is $ 1 - \sum_{1}^{i}x$.
+Let $y = f(x)$, then we define the stick-breaking mapping $ f \colon \Delta^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  R^{N-1}$, for $1 \leq i \leq N$ as:
+\[
+y_i
+= \mathrm{logit}(z_i) - \mbox{log}\left(\frac{1}{N-i}
+   \right) \text{for break proportion} \, 
+   z_i = \frac{x_i}{1 - \sum_{i' = 1}^{i-1} x_{i'}}.
+\]
+
+The inverse transform $ f^{-1} \colon R^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  \Delta^{N-1}$ is defined as:
+\[
+x_i =
+\left( 1 - \sum_{i'=1}^{i-1} x_{i'} \right) \text{for break proportion} \, z_i = \mathrm{logit}^{-1} \left( y_i
+                             + \log \left( \frac{1}{N - i}
+                                            \right)\right).
+                                            \]
+
+An $N$ dimensional unit simplex $\Delta^{N-1}$ has $N-1$ degrees of freedom (Notice that this inverse transform only maps the first $N-1$ elements, the last element of the simplex $x_{N} = 1 - \sum_1^{N-1}{x_i}$).
+The Jacobian matrix for $f^{-1}$ is a lower-triangular diagonal matrix.
+Much like the alr transform, for the change of variables we evaluate $\mathbf{J}_{i, i}$ where $i \in 1:N-1$.
+\begin{align*}
+\mathbf{J}_{i, i} &= \frac{\partial x_i}{\partial y_i}
+=
+\frac{\partial x_i}{\partial z_i} \,
+\frac{\partial z_i}{\partial y_i}\\
+\mathbf{J}_{i, i} &= \left(
+  1 - \sum_{k' = 1}^{k-1} x_{k'}
+   \right) z_k (1 - z_k),
+\end{align*}
+
+Absolute determinant of the diagonal matrix $\mathbf{J}$ is the product of its diagonal entries:
+\begin{align*}
+	\abs{\, det \, \textbf{J} \,} = \prod_{i=1}^{N-1} \textbf{J}_{i,i}
+\end{align*}
+
+The correction term $p_Y(y) = p_X(f^{-1}(y))\,
+\prod_{i=1}^{N-1}z_i\,(1 - z_i)\left(1 - \sum_{i'=1}^{i-1} x_{i'}\right).$
 \subsubsection{Softmax Transform}

### 2022-07-11 tex/paper.tex

--- 2022-07-08 tex/paper.tex
@@
 and $\absdet{J_s(y)}$ is the absolute value of its determinant.
+
+\subsection{StickBreaking Transform}
+
+The StickBreaking transform ican be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}.
+Intuitively, this comprises of recursively breaking a piece $x_i$ from a stick of unit length, where the leftover stick in the $i^{th}$ iteration is $ 1 - \sum_{1}^{i}x$.
+Let $y = f(x)$, then we define the stick-breaking mapping $ f \colon \Delta^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  R^{N-1}$, for $1 \leq i \leq N$ as:
+\[
+y_i
+= \mathrm{logit}(z_i) - \mbox{log}\left(\frac{1}{N-i}
+   \right) \text{for break proportion} \, 
+   z_i = \frac{x_i}{1 - \sum_{i' = 1}^{i-1} x_{i'}}.
+\]
+
+The inverse transform $ f^{-1} \colon R^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  \Delta^{N-1}$ is defined as:
+\[
+x_i =
+\left( 1 - \sum_{i'=1}^{i-1} x_{i'} \right) \text{for break proportion} \, z_i = \mathrm{logit}^{-1} \left( y_i
+                             + \log \left( \frac{1}{N - i}
+                                            \right)\right).
+                                            \]
+
+An $N$ dimensional unit simplex $\Delta^{N-1}$ has $N-1$ degrees of freedom (Notice that this inverse transform only maps the first $N-1$ elements, the last element of the simplex $x_{N} = 1 - \sum_1^{N-1}{x_i}$).
+The Jacobian matrix for $f^{-1}$ is a lower-triangular diagonal matrix.
+Much like the alr transform, for the change of variables we evaluate $\mathbf{J}_{i, i}$ where $i \in 1:N-1$.
+\begin{align*}
+\mathbf{J}_{i, i} &= \frac{\partial x_i}{\partial y_i}
+=
+\frac{\partial x_i}{\partial z_i} \,
+\frac{\partial z_i}{\partial y_i}\\
+\mathbf{J}_{i, i} &= \left(
+  1 - \sum_{k' = 1}^{k-1} x_{k'}
+   \right) z_k (1 - z_k),
+\end{align*}
+
+Absolute determinant of the diagonal matrix $\mathbf{J}$ is the product of its diagonal entries:
+\begin{align*}
+	\abs{\, det \, \textbf{J} \,} = \prod_{i=1}^{N-1} \textbf{J}_{i,i}
+\end{align*}
+
+The correction term $p_Y(y) = p_X(f^{-1}(y))\,
+\prod_{i=1}^{N-1}z_i\,(1 - z_i)\left(1 - \sum_{i'=1}^{i-1} x_{i'}\right).$
+\subsubsection{Softmax Transform}
 

### 2022-07-13 utils/dirichlet_symmetric.py
refactor scripts

@@ -1,0 +1,47 @@
+import numpy as np
+import pickle
+import matplotlib.pyplot as plt
+from rmse import rmse_leapfrog
+from sample import sample
+
+def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name, n_repeat=1, n_iter=1000, 
+                                n_chains=4, show_progress=True, resample=False):
+    
+    plt.rcParams["figure.figsize"] = [20,10]
+    plt.rcParams['figure.dpi'] = 300
+    fig, axes = plt.subplots(len(parameters))
+    fig.supxlabel('Cumulative Leapfrog Steps')
+    fig.supylabel('Root Mean Squared Error')
+
+    for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
+        for transform in transforms:            
+            idata = sample(transform_category=transform_category, transform=transform, 
+                evaluating_model='dirichlet_symmetric', parameters=[params], output_file=None, 
+                auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
+                                        show_progress = show_progress, resample=resample)
+            alpha = params['alpha']
+            N = params['N']
+            true_x = [a/sum(alpha) for a in alpha]
+            x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
+            ax.plot(x,y, label = str(transform))
+
+        ax.set_title(f'alpha={alpha[0]}, N = {N}')
+    fig.legend(labels=transforms,bbox_to_anchor = (0.6, -0.05));
+    fig.savefig(f'{fig_name}', dpi=300)
+
+def create_param_map():
+    alphas = [0.1, 1, 10]
+    Ns = [10, 100, 1000]
+    keys=[1,2,3,4,5,6,7,8,9]
+    params=[]
+    for alpha in alphas:
+        for N in Ns:
+            parameters.append({'alpha': [alpha]*N, 'N': N})
+
+    param_map=dict(zip(keys, params))
+    param_map.update(dict(zip(params, keys)))
+
+    pickle.dump(param_map, open('param_map_dirichlet_symmmetric.pkl', 'wb'))
+
+def get_dirichlet_symmetric_params():
+    return pickle.load(open('param_map_dirichlet_symmmetric.pkl', 'rb'))

### 2022-07-13 utils/dirichlet_symmetric_utils.py

@@ -1,24 +1,0 @@
-import numpy as np
-from utils import sample, retrieve, cumulative_mean, rmse, rmse_leapfrog
-
-def get_dirichlet_symmetric_params():
-    alphas=[0.1,1,10]
-    Ks = [10,100,1000]
-    parameters = []
-    for alpha in alphas:
-        for K in Ks:
-            parameters.append({'alpha': [alpha]*K, 'K': K})
-            
-    return parameters
-
-def get_dirichlet_symmetric_rmse(transform, parameters, n_repeat=1):
-    
-    idata = retrieve(transform_category="simplex", transform=transform, n_repeat=n_repeat,
-                 evaluating_model="dirichlet_symmetric", parameters=parameters)
-    
-    alpha = np.asarray(parameters['alpha'])
-    true_var = alpha/sum(alpha)
-    var_name='x'
-    var_dim=0
-    
-    return rmse_leapfrog(idata, true_var, var_name, var_dim) 

### 2022-07-13 utils/ess.py

@@ -1,0 +1,25 @@
+from scipy.stats import gaussian_kde
+import numpy as np
+import matplotlib.pyplot as plt
+from cmdstanpy import CmdStanModel
+from tqdm import tqdm
+import arviz as az
+
+def get_ess_leapfrog_ratio(transform_category, transform, evaluating_model, params, var_name, var_dim, repeat=100, return_ess=False):
+    model = CmdStanModel(stan_file = f'transforms/{transform_category}/{transform}/{evaluating_model}.stan', cpp_options = {'STAN_THREADS':'true'})
+    x=[]
+    for i in tqdm(range(repeat)):
+        idata = az.from_cmdstanpy(model.sample(data=params,show_progress=False ))
+        data=idata.sel(var_names=[str(var_name)], x_dim_0=var_dim)
+        ess = az.ess(data, var_names=[str(var_name)], method="bulk")[str(var_name)].values
+        leapfrog = idata.sample_stats['n_steps'].sum().values
+        x.append(ess/leapfrog)
+    if return_ess==True:
+        return x
+    else:
+        kde = gaussian_kde(x)
+        dist_space = np.linspace( min(x), max(x), 1000)
+        return dist_space, kde(dist_space)
+
+
+

### 2022-07-13 utils/ess_plot.py

@@ -1,58 +1,0 @@
-#!/usr/bin/env python
-# coding: utf-8
-
-from cmdstanpy import CmdStanModel
-import arviz as az
-import os
-import numpy as np
-from tqdm import tqdm
-import logging
-import matplotlib.pyplot as plt
-import numpy as np
-from scipy.stats import gaussian_kde
-os.chdir('..')
-logger = logging.getLogger('cmdstanpy')
-logger.addHandler(logging.NullHandler())
-
-def get_ess_leapfrog_plot(transforms=['stickbreaking', 'softmax', 'softmax-augmented'], 
-                          params={'alpha':[0.1]*10,'N':10}, 
-                          var_name='x', var_dim=0, repeat=100, plot_format='matplotlib', fig_name='figure.png'):
-   
-    def get_ess_leapfrog_ratio(transform, params={'alpha':[0.1]*10,'N':10}, 
-                          var_name='x', var_dim=0, repeat=100):
-        model = CmdStanModel(stan_file = f'transforms/simplex/{transform}/dirichlet_symmetric.stan', 
-                         cpp_options = {'STAN_THREADS':'true'})
-        x=[]
-        for i in tqdm(range(repeat)):
-            idata = az.from_cmdstanpy(model.sample(data=params,show_progress=False ))
-            data=idata.sel(var_names=[str(var_name)], x_dim_0=var_dim)
-            ess = az.ess(data, var_names=[str(var_name)], method="bulk")[str(var_name)].values
-            leapfrog = idata.sample_stats['n_steps'].sum().values
-            x.append(ess/leapfrog)
-        return x
-
-    data={}
-    for transform in transforms:
-        data[str(transform)] = get_ess_leapfrog_ratio(transform,repeat=100)
-        
-    if plot_format=='matplotlib':
-        fig, ax = plt.subplots()
-        for transform in transforms:
-            kde = gaussian_kde(data[str(transform)])
-            dist_space = np.linspace( min(data[str(transform)]), max(data[str(transform)]), 1000 )
-            ax.plot( dist_space, kde(dist_space), label=transform)
-        ax.set_title(str(params))
-        ax.axes.yaxis.set_ticklabels([])
-        plt.xlabel('ESS/Leapfrog')
-        plt.ylabel('Density')
-        plt.legend()
-        plt.savefig(str(fig_name), dpi=300)
-        plt.show()
-
-    if plot_format=='arviz':
[…]

### 2022-07-13 utils/rhat.py

@@ -1,0 +1,63 @@
+import os
+import arviz as az
+import numpy as np
+from cmdstanpy import CmdStanModel
+import logging
+logger = logging.getLogger('cmdstanpy')
+logger.addHandler(logging.NullHandler())
+
+def calc_rhat_mixed_chains(path_1, path_2, variable, data, force_compile=False):
+    
+    """
+    This function evaluates rhat for 4 chains from 2 different models, as in 2 chains from each.
+    
+    Parameters
+    ----------
+    
+    path_1: str
+        path for the first model
+        
+    path_2: str
+        path for the second model
+    
+    variable: str
+        Variable to evaluate rhat for
+    
+    data: dict
+        Dictionary of data
+    
+    force_compile: Bool
+        Whether to recompile or not
+
+    """
+    #Build stan model
+    file_1 = os.path.join(path_1)
+    file_2 = os.path.join(path_2)
+    model_1 = CmdStanModel(stan_file=file_1, cpp_options={'STAN_THREADS':'true'})
+    model_2 = CmdStanModel(stan_file=file_2, cpp_options={'STAN_THREADS':'true'})
+    
+    #Recompile
+    if force_compile is True:
+        model_1.compile(force=True)
+        model_2.compile(force=True)
+        
+    rhat_array=[]
+    for i in range(10):
+        #Fit stan model
+        fit_1 = model_1.sample(data=data, show_progress=False, iter_sampling=1000, chains=4)
+        fit_2 = model_2.sample(data=data, show_progress=False, iter_sampling=1000, chains=4)
+        
+        #Convert to idata for arviz
+        idata_1 = az.from_cmdstanpy(fit_1)
+        idata_2 = az.from_cmdstanpy(fit_2)
+        
+        #Stack the samples according to chains
+        stacked_1 = az.extract_dataset(idata_1)
+        stacked_2 = az.extract_dataset(idata_2)
+        
+        #Concatenate chains and evaluate rhat
+        chains = np.concatenate((stacked_1.sel(chain=1)[variable], stacked_1.sel(chain=2)[variable], stacked_2.sel(chain=1)[variable], stacked_2.sel(chain=2)[variable]), axis=1)
+        rhat = az.rhat(chains, var_names=variable,method="rank")
+        rhat_array.append(rhat)
+    
+    return np.asarray(rhat_array).mean()

### 2022-07-13 utils/rhat_eval.py

@@ -1,65 +1,0 @@
-import cmdstanpy
-import pandas as pd
-import arviz as az
-import numpy as np
-import matplotlib.pyplot as plt
-import xarray as xr
-import os
-from cmdstanpy import cmdstan_path, CmdStanModel
-
-
-simplex_stan = os.path.join('simplex-stan/simplex-stan.stan')
-simplex_stickbreaking = os.path.join('simplex-stickbreaking/simplex-stickbreaking.stan')
-simplex_softmax = os.path.join('simplex-softmax/simplex-softmax.stan')
-
-def calc_rhat_mixed_chains(path_1, path_2, variables, data, force_compile=False):
-    
-    """
-    This function evaluates rhat for 4 chains from 2 different models, as in 2 chains from each.
-    
-    Parameters
-    ----------
-    
-    path_1: str
-        path for the first model
-        
-    path_2: str
-        path for the second model
-    
-    variables: list
-        Variables to evaluate rhat for
-    
-    data: dict
-        Dictionary of data
-    
-    force_compile: Bool
-        Whether to recompile or not
-    """
-    variables = ' '.join(variables)
-    #Build stan model
-    file_1 = os.path.join(path_1)
-    file_2 = os.path.join(path_2)
-    model_1 = CmdStanModel(stan_file=file_1, cpp_options={'STAN_THREADS':'true'})
-    model_2 = CmdStanModel(stan_file=file_2, cpp_options={'STAN_THREADS':'true'})
-    
-    #Recompile
-    if force_compile is True:
-        model_1.compile(force=True)
-        model_2.compile(force=True)
-    
-    #Fit stan model
-    fit_1 = model_1.sample(data=dict(K=10))
-    fit_2 = model_2.sample(data=dict(K=10))
-    
-    #Convert to idata for arviz
-    idata_1 = az.from_cmdstanpy(fit_1)
-    idata_2 = az.from_cmdstanpy(fit_2)
-    
-    #Stack the samples according to chains
-    stacked_1 = az.extract_dataset(idata_1)
-    stacked_2 = az.extract_dataset(idata_2)
-    
-    #Concatenate chains and evaluate rhat
-    chains = np.concatenate((stacked_1.sel(chain=1)[str(variables)], stacked_1.sel(chain=2)[str(variables)], stacked_2.sel(chain=1)[str(variables)], stacked_2.sel(chain=2)[str(variables)]), axis=1)
-    rhat = az.rhat(chains, var_names=variables,method="rank")
-    return rhat

### 2022-07-13 utils/rmse.py

@@ -1,0 +1,18 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from sample import sample
+
+def rmse(y_true, y_pred):
+    return np.sqrt(np.mean((y_true - y_pred)**2))
+
+def cumulative_mean(x):
+    return np.divide(np.cumsum(x), np.arange(1, len(x) + 1))
+
+def rmse_leapfrog(idata, true_var, var_name, var_dim):
+    cumulative_leapfrog_steps = np.cumsum(np.mean(idata.sample_stats.n_steps, axis=0))
+    pred_var = cumulative_mean(np.mean(idata.posterior[str(var_name)].sel(x_dim_0=var_dim), axis=0))
+    rmse_array=[]
+    for i in range(1, len(pred_var)+1):
+        rmse_array.append(rmse(true_var[var_dim], pred_var[:i]))
+    return cumulative_leapfrog_steps, rmse_array
+

### 2022-07-13 utils/rmse_plot.py

@@ -1,28 +1,0 @@
-import os
-os.chdir('transforms')
-In [43]:
-from utils import sample, retrieve, cumulative_mean, rmse
-from dirichlet_symmetric_utils import get_dirichlet_symmetric_rmse, get_dirichlet_symmetric_params
-import pickle
-import numpy as np
-import matplotlib.pyplot as plt
-
-transforms = ['stan', 'stickbreaking', 'softmax', 'softmax-augmented']
-for transform in transforms:
-    sample(transform_category="simplex", transform=transform, evaluating_model="dirichlet_symmetric", 
-           parameters={}, auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 100)
-plt.rcParams["figure.figsize"] = (20,10)
-
-fig, axes = plt.subplots(3,3)
-
-fig.supxlabel('Cumulative Leapfrog Steps')
-fig.supylabel('Root Mean Squared Error')
-
-for ax, params in zip(axes.flatten(),  get_dirichlet_symmetric_params()):
-    for transform in transforms:
-        x, y = get_dirichlet_symmetric_rmse(transform, params)
-        ax.plot(x,y, label = str(transform))
-    ax.set_title(f'alpha={params["alpha"][0]}, K={params["K"]}')
-
-fig.legend(labels=transforms,bbox_to_anchor = (0.6, -0.05));
-fig.savefig(str('figure.png'))

### 2022-07-13 utils/sample.py

@@ -1,0 +1,96 @@
+import pandas as pd
+import numpy as np
+import os
+import json
+import sys
+import pickle
+from cmdstanpy import CmdStanModel
+import argparse
+import arviz as az
+from pathlib import Path
+from tqdm import tqdm
+# import logging
+# logger = logging.getLogger('cmdstanpy')
+# logger.addHandler(logging.NullHandler())
+
+def sample(transform_category, transform, evaluating_model, parameters, output_file=None, 
+                auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 1, show_progress = True, resample=False):
+    '''
+    Sample from the given dictionary containing the models, transform_categories, and parameters.
+    It saves the results in a json file which is named with n_repeat, n_chains and n_iter.
+
+    Parameters
+    ----------
+    transform_category: str
+        The transform category to use.
+    
+    transform: str
+        The transform name to use.
+    
+    evaluating_model: str
+        The model to use.
+    
+    parameters: list of dict
+        List of Dictionaries containing the parameters to use.
+        e.g. parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*20, 'N': 20}]
+
+    output_file: str
+        Custom filename.
+
+    auto_eval_all_params: bool
+        If True, all parameterizations are evaluated.
+
+    n_repeat: int
+        Number of times to repeat the sampling.
+
+    n_iter : int
+        Number of samples to be drawn.
+
+    n_chains : int
+        Number of chains to be drawn.
+
+    show_progress: bool (default = True)
+        Whether to show progress bar.
+
+    retrieve: bool (default = False)
+        Whether to retrieve the results from the json file.
+
+    Note
+    ----
+    The directory structure for storing stan files is: tramsforms/transform_category/transform/evaluating_model.stan
+    The directory structure for storing the results is: sampling_results/transform_category/transform/evaluating_model/param_keyword*.json
+    '''
+    with open(f'transforms/{transform_category}/param_map_{evaluating_model}.json', 'rb') as f:
+        param_map = pickle.load(f)
+
+    if resample==False:
+        for params in parameters:
+            filename = f'sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.json'
+            return az.from_json(filename)
+    else:
+
+        if auto_eval_all_params:
[…]

### 2022-07-13 utils/script.py

@@ -1,26 +1,0 @@
-from utils import sample, retrieve, cumulative_mean, rmse
-from dirichlet_symmetric_utils import get_dirichlet_symmetric_rmse, get_dirichlet_symmetric_params
-import pickle
-import numpy as np
-import matplotlib.pyplot as plt
-
-import os
-os.chdir('..')
-
-transforms = ['stan', 'stickbreaking', 'softmax', 'softmax-augmented']
-for transform in transforms:
-    sample(transform_category="simplex", transform=transform, evaluating_model="dirichlet_symmetric", 
-           parameters={}, auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 100)
-
-
-from utils import sample, retrieve, cumulative_mean, rmse
-from dirichlet_symmetric_utils import get_dirichlet_symmetric_rmse, get_dirichlet_symmetric_params
-import pickle
-import numpy as np
-import matplotlib.pyplot as plt
-
-import os
-os.chdir('..')
-
-sample(transform_category="simplex", transform='stickbreaking', evaluating_model="dirichlet_symmetric", 
-        parameters={}, auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 100)

### 2022-07-13 utils/script_simplex.py

@@ -1,0 +1,13 @@
+from utils import sample, retrieve, cumulative_mean, rmse
+from dirichlet_symmetric_utils import get_dirichlet_symmetric_rmse, get_dirichlet_symmetric_params
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+
+import os
+os.chdir('..')
+
+transforms = ['stan', 'stickbreaking', 'softmax', 'softmax-augmented']
+for transform in transforms:
+    sample(transform_category="simplex", transform=transform, evaluating_model="dirichlet_symmetric", 
+           parameters={}, auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 100)

### 2022-07-13 utils/utils.py

@@ -1,130 +1,0 @@
-import pandas as pd
-import numpy as np
-import os
-import json
-import sys
-import pickle
-from cmdstanpy import CmdStanModel
-import argparse
-import arviz as az
-from pathlib import Path
-from tqdm import tqdm
-
-
-
-def sample(transform_category, transform, evaluating_model, parameters, output_file=None, 
-                auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 1, show_progress = True):
-    '''
-    Sample from the given dictionary containing the models, transform_categories, and parameters.
-    It saves the results in a json file which is named with n_repeat, n_chains and n_iter.
-
-    Parameters
-    ----------
-    transform_category: str
-        The transform category to use.
-    
-    transform: str
-        The transform name to use.
-    
-    evaluating_model: str
-        The model to use.
-    
-    parameters: list of dict
-        List of Dictionaries containing the parameters to use.
-        e.g. parameters = [{'alpha': [0.1]*10, 'K': 10}, {'alpha': [0.1]*20, 'K': 20}]
-    
-    output_file: str
-        Custom filename.
-
-    auto_eval_all_params: bool
-        If True, all parameterizations are evaluated.
-
-    n_repeat: int
-        Number of times to repeat the sampling.
-
-    n_iter : int
-        Number of samples to be drawn.
-
-    n_chains : int
-        Number of chains to be drawn.
-
-    show_progress: bool (default = True)
-        Whether to show progress bar.
-
-    Note
-    ----
-    The directory structure for storing stan files is: tramsforms/transform_category/transform/evaluating_model.stan
-    The directory structure for storing the results is: sampling_results/transform_category/transform/evaluating_model/*.json
-    '''
-    if auto_eval_all_params:
-        with open(f'transforms/{transform_category}/{evaluating_model}_parameters.json', 'rb') as f:
-            parameters = pickle.load(f)
-    
-    with open(f'transforms/{transform_category}/dirichlet_symmetric_map.json', 'rb') as f:
-        param_map = pickle.load(f)
-
-    path=f'sampling_results/{transform_category}/{transform}/{evaluating_model}'
-    os.makedirs(path, exist_ok=True)
-    
-    if type(parameters)==dict:
-        parameters = [parameters]
-    for params in tqdm(parameters):
[…]

### 2022-07-17 tex/paper.tex
Update paper.tex

@@
 
+% transforms -> change of variables
+% two kinds of transforms
+% pragprg
 \begin{document}
@@
 
-Sampling algorithms play an important role in Statistics, in this case
-we refer to drawing samples from a probability density or distribution
-(not survey sampling). Monte Carlo Markov chain methods are commonly
-used for high dimensional sampling. In this paper, we look at
-Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
-samples for the Metropolis algorithm. Bayesian Inference often uses
-parameters with constraints, but HMC optimizes on the unconstrained
-space. It is difficult to use Monte Carlo estimation on a constrained
-domain ~\cite{neal2008optimal}
+%Sampling algorithms play an important role in Statistics, in this case
+%we refer to drawing samples from a probability density or distribution
+%(not survey sampling). Monte Carlo Markov chain methods are commonly
+%used for high dimensional sampling. In this paper, we look at
+%Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
+%samples for the Metropolis algorithm. Bayesian Inference often uses
+%parameters with constraints, but HMC optimizes on the unconstrained
+%space. It is difficult to use Monte Carlo estimation on a constrained
+%domain ~\cite{neal2008optimal}
 
@@
 
+\subsection{StickBreaking Transform}
+
+The StickBreaking transform ican be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}.
+Intuitively, this comprises of recursively breaking a piece $x_i$ from a stick of unit length, where the leftover stick in the $i^{th}$ iteration is $ 1 - \sum_{1}^{i}x$.
+Let $y = f(x)$, then we define the stick-breaking mapping $ f \colon \Delta^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  R^{N-1}$, for $1 \leq i \leq N$ as:
+\[
+y_i
+= \mathrm{logit}(z_i) - \mbox{log}\left(\frac{1}{N-i}
+   \right) \text{for break proportion} \, 
+   z_i = \frac{x_i}{1 - \sum_{i' = 1}^{i-1} x_{i'}}.
+\]
+
+The inverse transform $ f^{-1} \colon R^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  \Delta^{N-1}$ is defined as:
+\[
+x_i =
+\left( 1 - \sum_{i'=1}^{i-1} x_{i'} \right) \text{for break proportion} \, z_i = \mathrm{logit}^{-1} \left( y_i
+                             + \log \left( \frac{1}{N - i}
+                                            \right)\right).
+                                            \]
+
+An $N$ dimensional unit simplex $\Delta^{N-1}$ has $N-1$ degrees of freedom (Notice that this inverse transform only maps the first $N-1$ elements, the last element of the simplex $x_{N} = 1 - \sum_1^{N-1}{x_i}$).
+The Jacobian matrix for $f^{-1}$ is a lower-triangular diagonal matrix.
+Much like the alr transform, for the change of variables we evaluate $\mathbf{J}_{i, i}$ where $i \in 1:N-1$.
+\begin{align*}
+\mathbf{J}_{i, i} &= \frac{\partial x_i}{\partial y_i}
+=
+\frac{\partial x_i}{\partial z_i} \,
+\frac{\partial z_i}{\partial y_i}\\
+\mathbf{J}_{i, i} &= \left(
+  1 - \sum_{k' = 1}^{k-1} x_{k'}
+   \right) z_k (1 - z_k),
+\end{align*}
+
+Absolute determinant of the diagonal matrix $\mathbf{J}$ is the product of its diagonal entries:
+\begin{align*}
+	\abs{\, det \, \textbf{J} \,} = \prod_{i=1}^{N-1} \textbf{J}_{i,i}
+\end{align*}
+
+The correction term $p_Y(y) = p_X(f^{-1}(y))\,
+\prod_{i=1}^{N-1}z_i\,(1 - z_i)\left(1 - \sum_{i'=1}^{i-1} x_{i'}\right).$
 \subsection{Additive log ratio transform}
@@
 
-\subsection{StickBreaking Transform}
-
-The StickBreaking transform ican be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}.
-Intuitively, this comprises of recursively breaking a piece $x_i$ from a stick of unit length, where the leftover stick in the $i^{th}$ iteration is $ 1 - \sum_{1}^{i}x$.
-Let $y = f(x)$, then we define the stick-breaking mapping $ f \colon \Delta^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  R^{N-1}$, for $1 \leq i \leq N$ as:
-\[
-y_i
-= \mathrm{logit}(z_i) - \mbox{log}\left(\frac{1}{N-i}
-   \right) \text{for break proportion} \, 
-   z_i = \frac{x_i}{1 - \sum_{i' = 1}^{i-1} x_{i'}}.
-\]
-
-The inverse transform $ f^{-1} \colon R^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  \Delta^{N-1}$ is defined as:
-\[
-x_i =
-\left( 1 - \sum_{i'=1}^{i-1} x_{i'} \right) \text{for break proportion} \, z_i = \mathrm{logit}^{-1} \left( y_i
-                             + \log \left( \frac{1}{N - i}
-                                            \right)\right).
-                                            \]
-
-An $N$ dimensional unit simplex $\Delta^{N-1}$ has $N-1$ degrees of freedom (Notice that this inverse transform only maps the first $N-1$ elements, the last element of the simplex $x_{N} = 1 - \sum_1^{N-1}{x_i}$).
-The Jacobian matrix for $f^{-1}$ is a lower-triangular diagonal matrix.
-Much like the alr transform, for the change of variables we evaluate $\mathbf{J}_{i, i}$ where $i \in 1:N-1$.
-\begin{align*}
-\mathbf{J}_{i, i} &= \frac{\partial x_i}{\partial y_i}
-=
-\frac{\partial x_i}{\partial z_i} \,
-\frac{\partial z_i}{\partial y_i}\\
-\mathbf{J}_{i, i} &= \left(
-  1 - \sum_{k' = 1}^{k-1} x_{k'}
-   \right) z_k (1 - z_k),
-\end{align*}
-
-Absolute determinant of the diagonal matrix $\mathbf{J}$ is the product of its diagonal entries:
-\begin{align*}
-	\abs{\, det \, \textbf{J} \,} = \prod_{i=1}^{N-1} \textbf{J}_{i,i}
-\end{align*}
-
-The correction term $p_Y(y) = p_X(f^{-1}(y))\,
-\prod_{i=1}^{N-1}z_i\,(1 - z_i)\left(1 - \sum_{i'=1}^{i-1} x_{i'}\right).$
 \subsubsection{Softmax Transform}
@@
 
-\subsection{Softmax Transform}
-
-The softmax function can be understood from Multinomial Logistic
-Regression employed for predicting probabilities of a Categorically
-distributed variable. Geometrically it maps $R^K$ to the boundary of a
-unit K-simplex(it is simply the convex hull of $k+1$ affinely
-independent points in $R^K$. Essentially it transforms a vector of
-size K to another vector of size K where the outputs sum to 1. It is
-worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
-when a vector of size $K$ is transformed the $K_th$ vector is simply
-$1$- sumof k-1 vectors
+%\subsection{Softmax Transform}
+%
+%The softmax function can be understood from Multinomial Logistic
+%Regression employed for predicting probabilities of a Categorically
+%distributed variable. Geometrically it maps $R^K$ to the boundary of a
+%unit K-simplex(it is simply the convex hull of $k+1$ affinely
+%independent points in $R^K$. Essentially it transforms a vector of
+%size K to another vector of size K where the outputs sum to 1. It is
+%worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
+%when a vector of size $K$ is transformed the $K_th$ vector is simply
+%$1$- sumof k-1 vectors
 

### 2022-07-18 utils/dirichlet_symmetric.py
black format all python files

@@ -4,44 +4,67 @@ import matplotlib.pyplot as plt
 from rmse import rmse_leapfrog
 from sample import sample
 
-def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name, n_repeat=1, n_iter=1000, 
-                                n_chains=4, show_progress=True, resample=False):
-    
-    plt.rcParams["figure.figsize"] = [20,10]
-    plt.rcParams['figure.dpi'] = 300
+
+def get_dirichlet_symmetric_rmse(
+    transforms,
+    transform_category,
+    parameters,
+    fig_name,
+    n_repeat=1,
+    n_iter=1000,
+    n_chains=4,
+    show_progress=True,
+    resample=False,
+):
+
+    plt.rcParams["figure.figsize"] = [20, 10]
+    plt.rcParams["figure.dpi"] = 300
     fig, axes = plt.subplots(len(parameters))
-    fig.supxlabel('Cumulative Leapfrog Steps')
-    fig.supylabel('Root Mean Squared Error')
+    fig.supxlabel("Cumulative Leapfrog Steps")
+    fig.supylabel("Root Mean Squared Error")
 
-    for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
-        for transform in transforms:            
-            idata = sample(transform_category=transform_category, transform=transform, 
-                evaluating_model='dirichlet_symmetric', parameters=[params], output_file=None, 
-                auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
-                                        show_progress = show_progress, resample=resample)
-            alpha = params['alpha']
-            N = params['N']
-            true_x = [a/sum(alpha) for a in alpha]
-            x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
-            ax.plot(x,y, label = str(transform))
+    for ax, params in zip(
+        axes.flatten() if len(parameters) > 1 else [axes], parameters
+    ):
+        for transform in transforms:
+            idata = sample(
+                transform_category=transform_category,
+                transform=transform,
+                evaluating_model="dirichlet_symmetric",
+                parameters=[params],
+                output_file=None,
+                auto_eval_all_params=False,
[…]

### 2022-07-18 utils/rmse.py

@@ -2,17 +2,21 @@ import numpy as np
 import matplotlib.pyplot as plt
 from sample import sample
 
+
 def rmse(y_true, y_pred):
-    return np.sqrt(np.mean((y_true - y_pred)**2))
+    return np.sqrt(np.mean((y_true - y_pred) ** 2))
+
 
 def cumulative_mean(x):
     return np.divide(np.cumsum(x), np.arange(1, len(x) + 1))
 
+
 def rmse_leapfrog(idata, true_var, var_name, var_dim):
     cumulative_leapfrog_steps = np.cumsum(np.mean(idata.sample_stats.n_steps, axis=0))
-    pred_var = cumulative_mean(np.mean(idata.posterior[str(var_name)].sel(x_dim_0=var_dim), axis=0))
-    rmse_array=[]
-    for i in range(1, len(pred_var)+1):
+    pred_var = cumulative_mean(
+        np.mean(idata.posterior[str(var_name)].sel(x_dim_0=var_dim), axis=0)
+    )
+    rmse_array = []
+    for i in range(1, len(pred_var) + 1):
         rmse_array.append(rmse(true_var[var_dim], pred_var[:i]))
     return cumulative_leapfrog_steps, rmse_array
-

### 2022-07-18 utils/sample.py

@@ -9,13 +9,27 @@ import argparse
 import arviz as az
 from pathlib import Path
 from tqdm import tqdm
+
 # import logging
 # logger = logging.getLogger('cmdstanpy')
 # logger.addHandler(logging.NullHandler())
 
-def sample(transform_category, transform, evaluating_model, parameters, output_file=None, 
-                auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 1, show_progress = True, resample=False, return_idata=True):
-    '''
+
+def sample(
+    transform_category,
+    transform,
+    evaluating_model,
+    parameters,
+    output_file=None,
+    auto_eval_all_params=False,
+    n_iter=1000,
+    n_chains=4,
+    n_repeat=1,
+    show_progress=True,
+    resample=False,
+    return_idata=True,
+):
+    """
     Sample from the given dictionary containing the models, transform_categories, and parameters.
     It saves the results in a json file which is named with n_repeat, n_chains and n_iter.
 
@@ -23,13 +37,13 @@ def sample(transform_category, transform, evaluating_model, parameters, output_f
     ----------
     transform_category: str
         The transform category to use.
-    
+
     transform: str
         The transform name to use.
-    
+
     evaluating_model: str
         The model to use.
-    
+
     parameters: list of dict
         List of Dictionaries containing the parameters to use.
         e.g. parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*20, 'N': 20}]
@@ -57,7 +71,7 @@ def sample(transform_category, transform, evaluating_model, parameters, output_f
 
     resample: bool (default = False)
         Whether to resample the data.
-    
+
     return_idata: bool (default = True)
         Whether to return the idata.
 
@@ -65,37 +79,55 @@ def sample(transform_category, transform, evaluating_model, parameters, output_f
     ----
     The directory structure for storing stan files is: stan_models/{transform}_{evaluating_model}.stan
     The directory structure for storing the results is: sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_mapping}.json
-    '''
-    with open(f'target_densities/param_map_{evaluating_model}.json', 'rb') as f:
+    """
+    with open(f"target_densities/param_map_{evaluating_model}.json", "rb") as f:
         param_map = pickle.load(f)
 
-    if resample==False:
+    if resample == False:
         for params in parameters:
[…]

### 2022-07-18 utils/rhat.py
black format all python files

@@ -3,61 +3,75 @@ import arviz as az
 import numpy as np
 from cmdstanpy import CmdStanModel
 import logging
-logger = logging.getLogger('cmdstanpy')
+
+logger = logging.getLogger("cmdstanpy")
 logger.addHandler(logging.NullHandler())
 
+
 def calc_rhat_mixed_chains(path_1, path_2, variable, data, force_compile=False):
-    
+
     """
     This function evaluates rhat for 4 chains from 2 different models, as in 2 chains from each.
-    
+
     Parameters
     ----------
-    
+
     path_1: str
         path for the first model
-        
+
     path_2: str
         path for the second model
-    
+
     variable: str
         Variable to evaluate rhat for
-    
+
     data: dict
         Dictionary of data
-    
+
     force_compile: Bool
         Whether to recompile or not
 
     """
-    #Build stan model
+    # Build stan model
     file_1 = os.path.join(path_1)
     file_2 = os.path.join(path_2)
-    model_1 = CmdStanModel(stan_file=file_1, cpp_options={'STAN_THREADS':'true'})
-    model_2 = CmdStanModel(stan_file=file_2, cpp_options={'STAN_THREADS':'true'})
-    
-    #Recompile
+    model_1 = CmdStanModel(stan_file=file_1, cpp_options={"STAN_THREADS": "true"})
+    model_2 = CmdStanModel(stan_file=file_2, cpp_options={"STAN_THREADS": "true"})
+
+    # Recompile
     if force_compile is True:
         model_1.compile(force=True)
         model_2.compile(force=True)
-        
-    rhat_array=[]
+
+    rhat_array = []
     for i in range(10):
-        #Fit stan model
-        fit_1 = model_1.sample(data=data, show_progress=False, iter_sampling=1000, chains=4)
-        fit_2 = model_2.sample(data=data, show_progress=False, iter_sampling=1000, chains=4)
-        
-        #Convert to idata for arviz
+        # Fit stan model
+        fit_1 = model_1.sample(
+            data=data, show_progress=False, iter_sampling=1000, chains=4
+        )
+        fit_2 = model_2.sample(
+            data=data, show_progress=False, iter_sampling=1000, chains=4
+        )
+
+        # Convert to idata for arviz
         idata_1 = az.from_cmdstanpy(fit_1)
         idata_2 = az.from_cmdstanpy(fit_2)
-        
-        #Stack the samples according to chains
+
+        # Stack the samples according to chains
[…]

### 2022-07-18 tex/paper.tex
Draft section on methods for computing Jacobian determinants

@@
 \newcommand{\absdet}[1]{\abs{#1}}
+\newcommand{\dv}[1]{\mathrm{d}{#1}}
+\newcommand{\vect}{\mathrm{vec}}
+\newcommand{\vectu}{\mathrm{vecu}}
 
@@
 While $p_X$ is generally fixed by the user-specified model and may not have this necessary property, the user may then choose a $p_Z$ that has this property.
+
+\section{Preliminaries}
+
+[SA: TODO: eliminate or merge with following section]
+
+To compute Jacobians of functions with matrix-valued inputs or outputs, we need to implicitly or explicitly choose for the inputs and outputs a set of coordinates with dimension equal to the number of degrees of freedom of the matrix
+Unless otherwise specified, we choose the coordinates to be the vectorization of the matrix.
+
+We define the (bijective) vectorization of a matrix $X \in \mathbb{R}^{N \times K}$ with columns $x_k$ for $k \in \{1, ..., K\}$ as the vector
+\[\vect(X) = \begin{bmatrix}x_1 \\ x_2 \\ \vdots \\ x_M\end{bmatrix}\]
+formed by stacking the columns of $X$.
+
+For $X \in \mathbb{R}^{N \times K}$, let $\vectu(X)$ be the map that stacks only the parts of the columns of $X$ above and including the diagonal.
+This map is sometimes called the half-vectorization and is bijective for upper-triangular or symmetric $X$.
+
+Similarly, $\vectu_-(X)$ extracts the same elements except the diagonal.
+When $X$ is strictly upper-triangular or skew-symmetric (i.e. $X = -X^\top$), then this map is bijective.
+
+\section{Computing Jacobian determinants}
+
+[SA: TODO: remove everything we do not use later, add strategy for making non-square Jacobian square]
+
+Here we introduce several strategies for computing Jacobian determinants.
+
+The vectorization has the useful property that for all matrices $A$, $X$, and $B$, $\vect(AYB) = (B^\top \otimes A) \vect(Y)$, where $\otimes$ is the Kronecker product.
+This allows us to convert expressions of matrices to expressions of vectors, from which the Jacobian can be read off.
+For example, let $X = AYB$. Then $\dv{X} = A \dv{Y} B$, and $\vect(\dv{X}) = (B^\top \otimes A) \vect(\dv{Y})$, and the Jacobian is $J = B^\top \otimes A$.
+
+When the previously discussed strategies are insufficient to compute the Jacobian, we bypass the Jacobian computation entirely and use the exterior product $\wedge$.
+For example, let $X \in \mathbb{R}^{N \times K}$.
+The exterior product
+\[
+  \dv{X_{11} }\wedge \dv{X_{21}} \wedge \ldots \wedge \dv{X_{NK}} = \bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{X}_{ij}
+\]
+is another more explicit way of writing the integration measure $\dv{X_{11}} \dv{X_{21}} \dots \dv{X_{NK}}$ that has useful properties for computing Jacobian determinants.
+The basic properties we will use are
+
+\begin{itemize}
+  \item $a\dv{x} \wedge b\dv{y} = (ab)(\dv{x} \wedge \dv{y})$ for $a,b \in \mathbb{R}$
+  \item $(\dv{x} + \dv{y}) \wedge \dv{z} = \dv{x} \wedge \dv{z} + \dv{y} \wedge \dv{z}$
+  \item $\dv{x} \wedge \dv{y} = - \dv{y} \wedge \dv{x}$
+  \item $\dv{x} \wedge \dv{x} = 0$
+\end{itemize}
+
+The connection between the exterior product and the Jacobian determinant is given by the following expression.
+Given $f: y \mapsto x$ for $x,y \in \mathbb{R}^N$,
+\[\bigwedge_{i=1}^N \dv{x_i} = \det(J_f(y)) \bigwedge_{i=1}^N \dv{y_i}.\]
+
+We can then find the Jacobian determinant by first writing all differential forms $\dv{x_i}$ or $\dv{y_i}$ and then computing the exterior product until the expression looks like the above.
+For our purposes, we will ignore the sign of the Jacobian determinant.
+
+A related useful property is that for $X,Y,B \in \mathbb{R}^{N \times K}$, if $\dv{X} = B \dv{Y}$, then
+\[\bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{X_i} = \det(B)^K \bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{Y_i}\]
+
+Following [REF, Edelman], we use the notation $(\cdot)^\wedge$ as a shorthand to represent the exterior product of the unique elements of $\cdot$.
+
+For example,
+\begin{itemize}
+  \item Rectangular matrix $\{X \in \mathbb{R}^{N \times K}\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{X_{ij}}$
+  \item Upper-triangular rectangular matrix $\{X \in \mathbb{R}^{N \times K} | X_{ij} = 0\ \forall\ i > j\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{\min(j, N)} \dv{X_{ij}}$
+  \item Diagonal matrix $\{X \in \mathbb{R}^{N \times K} | X_{ij} = 0\ \forall\ i \ne j\}$: $(\dv{X})^\wedge = \bigwedge_{i=1}^{\min(N, K)} \dv{X_{ii}}$
+  \item Symmetric matrix $\{X \in \mathbb{R}^{N \times N} | X_{ij} = X_{ji}\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{j} \dv{X_{ij}}$
+  \item Skew-symmetric matrix $\{X \in \mathbb{R}^{N \times N} | X_{ij} = -X_{ji}, X_{ii} = 0\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{j-1} \dv{X_{ij}}$
+\end{itemize}
+
+Using this notation, for example, $(B \dv{X})^\wedge = \det(B)^K (\dv{X})^\wedge$
 

### 2022-07-18 tex/paper.tex
Draft section on methods for computing Jacobian determinants

@@
 
-% transforms -> change of variables
-% two kinds of transforms
-% pragprg
 \begin{document}
@@
 \cite{neal2011mcmc} or unit vectors \cite{byrne2013geodesic}, it is
-much more challenging to do the same for complex constraints such as
-simplexes or positive definite matrices or for densities involving
-multiple constrained values.  Instead, it is far more common to map
-the constrained values to unconstrained values before sampling
-\cite{}.  This presents the issue of selecting which transform to use
-among an infinite set of options, which is the topic of this paper.
+far more challenging to do the same for complex constraints such as
+simplices or positive definite matrices or for densities involving
+multiple constrained values.  Instead, usually constrained values are mapped to unconstrained values before sampling
+\cite{}.
+A change of variables adjustment is made to the target density for these transforms to work.
+The tricky part is- there can be more than one suitable mapping for every constraint.
+This leads us to the issue of selecting a transform.
 
-%Sampling algorithms play an important role in Statistics, in this case
-%we refer to drawing samples from a probability density or distribution
-%(not survey sampling). Monte Carlo Markov chain methods are commonly
-%used for high dimensional sampling. In this paper, we look at
-%Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
-%samples for the Metropolis algorithm. Bayesian Inference often uses
-%parameters with constraints, but HMC optimizes on the unconstrained
-%space. It is difficult to use Monte Carlo estimation on a constrained
-%domain ~\cite{neal2008optimal}
+\section{Transforms}
 
-\section{Changes of Variables}
+To draw samples from a distribution $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.
+
+Let $f\colon \mathcal{Y} \to \mathcal{X}$ be a smooth and continuous map.
+Only when $m = n$ can $f$ be bijective.
+When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = f(y)$.
+In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
+We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
+$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
+that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
+When $h$ is non-bijective at single points, we call these singularities.
+While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.
+
+\subsection{Changes of Variables}
 
@@
 
-\section{Transforms}
-
-To draw samples from a distribution $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.
-
-Let $f\colon \mathcal{Y} \to \mathcal{X}$ be a smooth and continuous map.
-Only when $m = n$ can $f$ be bijective.
-When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = f(y)$.
-In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
-We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
-$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
-that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
-When $h$ is non-bijective at single points, we call these singularities.
-While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.
-
+\subsection{Target Density}
 Given a proper joint distribution with a smooth and continuous density $p_{X,Y}(x, y | \theta)$ for some parameters $\theta$, we can use the usual change of variables formula to write
@@
 
-The StickBreaking transform ican be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}.
+The StickBreaking transform can be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}.
 Intuitively, this comprises of recursively breaking a piece $x_i$ from a stick of unit length, where the leftover stick in the $i^{th}$ iteration is $ 1 - \sum_{1}^{i}x$.
@@
   \\[6pt]
-  & = &
-  \left(
-    1 
-    + \textrm{sum}\left( \frac{-(1 + \textrm{sum}(z)) z}{(1 +
-        \textrm{sum}(z))^2} \right)
-  \right)
-        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
-  \\[6pt]
-  & = &
-        \left(1 + \textrm{sum}\left(\frac{-z}{1 + \textrm{sum}(z)} \right)\right)        
-        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
-  \\[6pt]
-  & = & \left( 1 - \textrm{sum}(\textrm{norm}(z)) \right) 
-        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
-  \\[6pt]
   & = & \left( \frac{1}{1 + \textrm{sum}(z)} \right)^N.
@@
 \]
-
-%\subsection{Softmax Transform}
-%
-%The softmax function can be understood from Multinomial Logistic
-%Regression employed for predicting probabilities of a Categorically
-%distributed variable. Geometrically it maps $R^K$ to the boundary of a
-%unit K-simplex(it is simply the convex hull of $k+1$ affinely
-%independent points in $R^K$. Essentially it transforms a vector of
-%size K to another vector of size K where the outputs sum to 1. It is
-%worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
-%when a vector of size $K$ is transformed the $K_th$ vector is simply
-%$1$- sumof k-1 vectors
-
-\subsection{Simplex softmax parameterization}
-
-Let $\Delta^n$ indicate the unit $n$-simplex with $n$ positive
-elements that sum to 1 and $\Delta^n_-$ indicate the first $n-1$
-elements, from which the final element can be uniquely determined.
-
-We define the transformation
-$\phi: \mathbb{R}^{n-1} \to \Delta^n_-: y \mapsto x_-$, where
-$x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix} \in \Delta^n$,
-$x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$, and
-$r = 1 + \sum_{i=1}^{n-1} e^{y_i}$.
-
-First we compute the scalar derivatives:
-\[
-\begin{aligned}
-  \frac{\mathrm{d} r}{\mathrm{d} y_j}
-  &= e^{y_j} = r x_j\\
-  \frac{\mathrm{d} x_i}{\mathrm{d} y_j}
-  &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i}
-  \frac{\mathrm{d} r}{\mathrm{d} y_j}
-  = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
-\end{aligned}
-\]
-where
-$\delta_{ij} = \begin{cases} 1 &\text{if } i = j \\ 0
-  &\text{otherwise}\end{cases}$ is the Kronecker delta.
-
-If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the
-elements of $x$, then the Jacobian is
-\[
-  J = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top) \operatorname{diag}(x_-),
-\]
-where $\boldsymbol{1}_n$ is the $n$-vector of ones.
-
-Using Sylvester's determinant theorem,
-$|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = 1 -
-\boldsymbol{1}_{n-1}^\top x_- = 1 - \sum_{i=1}^{n-1} x_i = x_n$, so
-$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} y_i\right) \left(1 + \sum_{i=1}^{n-1} e^{y_i}\right)^{-n}$$
 
@@
 $\begin{pmatrix}A & B \\ C & D\end{pmatrix}$ is $|A| |D-CA^{-1}B|$.  A
-square matrix is invertible iff its determinant is non-yero.  From the
+square matrix is invertible iff its determinant is non-zero.  From the
 previous section,

### 2022-07-18 tex/paper.tex
Draft section on semi-orthogonal matrices

@@
 
+\subsection{Semi-orthogonal matrices (Stiefel manifold)}
+
+A matrix $X \in \mathbb{R}^{N \times K}$ for $K \le N$ is semi-orthogonal if and only if $X^\top X = I_K$.
+The semi-orthogonal matrices form a space called the Stiefel manifold.
+When $K=N$, $X \in O(N)$ is an orthogonal matrix and also satisfies $X X^\top = I_N$ and $\det(X) = \pm 1$.
+When $K=1$, $X$ is a point on the $(N-1)$-sphere $\mathbb{S}^{N-1} \subset \mathbb{R}^N$.
+
+By choosing a set of rules such as Gram-Schmidt orthonormalization of the canonical basis in $\mathbb{R}^{N \times N}$, it is possible to analytically compute from $X$ an orthogonal complement $X^\perp \in \mathbb{R}^{N \times (N - K)}$, such that $\tilde{X}^\top \tilde{X} = I_N$, where $\tilde{X} = \begin{bmatrix}X & X^\perp \end{bmatrix} \in O(N)$.
+
+Note that if we differentiate both sides of the constraint $X^\top X = I_K$, we get $\dv{X}^\top X + X^\top \dv{X} = 0$, which leads to $(X^\top \dv{X})^\top = - X^\top \dv{X}$.
+Therefore, $X^\top \dv{X}$ is a $K \times K$ skew-symmetric matrix with $\frac{K(K-1)}{2}$ unique entries in the strict upper triangle.
+
+Densities of distributions on the Stiefel manifold are usually written with respect to the invariant metric on the Stiefel manifold, which we will define separately for the two cases $K=N$ and $K < N$.
+
+For $X \in O(N)$ with the $i$th column written $x_i$, we define the invariant measure as the exterior product of the unique elements of this skew-symmetric matrix:
+\[
+  (X^\top \dv{X})^\wedge \equiv \bigwedge_{j=1}^N \bigwedge_{i=j+1}^N (X^\top \dv{X})_{ij} = \bigwedge_{j=1}^N \bigwedge_{i=j+1}^N x_i^\top \dv{x}_j.
+\]
+This measure is called invariant because if $f(X) = A X B$ for $A,B \in O(N)$ and $\mathcal{S} \subset O(N)$ then $\int_{\mathcal{S}} (X^\top \dv{X}) = \int_{f(\mathcal{S})} (X^\top \dv{X})$.
+It is also called the Haar measure on $O(N)$ \cite{muirhead2009aspects}..
+
+To choose an invariant measure for $K < N$, we first choose an $X^\perp$ analytically computed from $X$.
+Because $X^\perp$ is analytically computed, $\dv{X^\perp_{ij}} = 0$.
+We can then construct the invariant measure for $K<N$ from the invariant measure for $O(N)$:
+
+\[
+  (X^\top \dv{X})^\wedge \equiv \bigwedge_{j=1}^N \bigwedge_{i=j+1}^N x_i^\top \dv{x}_j = \bigwedge_{j=1}^K \bigwedge_{i=j+1}^N x_i^\top \dv{x}_j,
+\]
+
+Though we needed to choose a $X^\perp$, the invariant measure does not depend on the choice of $X^\perp$.
+Moreover, this measure is invariant to left-actions by $O(N)$ and right-actions by $O(K)$.
+For more details, see \cite{muirhead2009aspects}.
+
+When $K=1$, the invariant measure is equivalent to the usual Hausdorff measure on the sphere.
+
+\subsubsection{Givens parameterization of Stiefel manifold}
+
+\subsubsection{Cayley parameterization of Stiefel manifold}
+
+\subsubsection{Householder parameterization of Stiefel manifold}
+
+\subsubsection{Matrix exponential parameterization of Orthogonal group}
+
 \subsubsection*{Acknowledgements}

### 2022-07-18 tex/paper.tex
Draft section on positive-definite matrices

--- 2022-07-18 tex/paper.tex
@@
 \newcommand{\dv}[1]{\mathrm{d}{#1}}
+\newcommand{\Exp}{\mathrm{Exp}}
 \newcommand{\vect}{\mathrm{vec}}
@@
 
+\subsection{Positive definite matrices}
+
+[SA: all transformations below can be generalized to SPD matrices of fixed rank, but we'd need to know what the common reference measure is for densities of SPD matrices, and I can't recall seeing one.]
+
+An $N \times N$ positive definite (PD) matrix $X$ is a symmetric matrix that satisfies $v^\top X v \ge 0$ for all non-zero vectors $v$.
+Covariance matrices are examples of PD matrices.
+
+Density functions for distributions of PD matrices, such as the Wishart distribution, are generally defined with respect to the Lebesgue measure of the upper triangular elements $(\dv{X})^\wedge$.
+
+\subsubsection{Cholesky parameterization}
+
+Given a positive (semi-)definite $X$ of rank $K$, we can uniquely decompose it as $X = Z^\top Z$ where $Z$ is an $N \times N$ upper triangular matrix where $Z_{ii} > 0$ for all $i \le K$ and $Z_{ij} = 0$ for all $i > K$.
+$Z$ has $\frac{K(K+1)}{2} + (N - K)K$ degrees of freedom.
+When $K=N$, $X$ is positive-definite.
+
+We can construct $X$ by composing 2 bijective maps.
+The first, $f_1$, enforces the constraint of the positive diagonal:
+
+\[Z_{ij} = f_1(Y)_{ij} = \begin{cases} \exp(Y_{ij}) & \text{ if } i = j \le K \\ Y_{ij} & \text{otherwise} \end{cases}\]
+
+The density correction for this map is $\absdet{J_{f_1}(Y)} = \exp(\sum_{i=1}^K Y_{ii})$.
+
+The second map is $f_2: Z \mapsto Z^\top Z = X$.
+To compute the Jacobian of the map, we use the half-vectorization $\vectu(X)$.
+Note that $X_{ij} = \sum_{k=1}^{\min(i, j)} Z_{ki} Z_{kj}$.
+$X_{ij}$ is thus only dependent on elements of $Z$ that are above and to the left of the $(i, j)$ position.
+As a result, the Jacobian of $\vectu \circ f_2 \circ \\vectu^{-1}$ is lower-triangular, and its determinant is only the product of its diagonal elements: $\absdet{J_{f_2}} = \prod_{1 \le i \le j \le N} \frac{\partial X_{ij}}{\partial Z_{ij}}$.
+
+In this index range, $\frac{\partial X_{ij}}{\partial Z_{ij}} = (1 + \delta_{ij}) Z_{ii}$.
+As a result
+
+\[
+  \absdet{J_{f_2}(Z)} = 2^N \prod_{j = 1}^N \prod_{i = 1}^j Z_{ii} = 2^N \prod_{i=1}^N |Z_{ii}|^{N - i + 1} \\
+\]
+
+Combining the two bijective maps as $f = f_2 \circ f_1$, the combined density correction is
+
+\[
+  \absdet{J_{f}(Y)} = 2^N \exp\left( \sum_{i=1}^N (N - i + 2) Y_{ii}\right).
+\]
+
+\subsubsection{Matrix exponential parameterization}
+
+For every PD matrix $X \in \mathbb{R}^{N \times N}$, there is a unique symmetric matrix $Y$ such that $X = \Exp(Y)$, where $\operatorname{Exp}(\cdot)$ is the matrix exponential, defined as
+\[\Exp(Y) = \sum_{k=0}^\infty \frac{1}{k!} Y^k.\]
+
+Let $Y = U \Lambda U^\top$ be the symmetric eigendecomposition of $Y$, where the matrix of eigenvectors $U$ is orthogonal, and the $\Lambda = \operatorname{diag}(\lambda)$ for $\lambda \in \mathbb{R}^N$.
+Then $Y^2 = U \Lambda U^\top U \Lambda U^\top = U \Lambda^2 U^\top$.
+In fact, $Y^c = U \Lambda^c U^\top$ for all real $c$.
+As a result,
+\[\Exp(Y) = U \Exp(\Lambda) U^\top,\]
+where $\Exp(\Lambda)_{ij} = e^{\lambda_i} \delta_{ij}$.
+
+$\Exp(Y)$ is one example of a matrix function $f$, which we can roughly define for symmetric matrices as any function that obeys
+\[f(Y) = U f(\Lambda) U^\top,\]
+where $f(\Lambda)_{ij} = f(\lambda_i)\delta_{ij}$.
+\footnote{
+  In a similar way, we can generalize other scalar functions like the inverse, trigonometric, and hyperbolic functions to functions of matrices.
+  All results in this section apply to these functions as well.
+}
+
+For such functions $f$, we can compute a common form of the Jacobian determinant.
+Differentiating the expression, we find
+\[
+\begin{aligned}
+  \dv{(f(Y))} &= \dv{U} f(\Lambda) U^\top + U f'(\Lambda) \dv{\Lambda} U^\top + U f(\Lambda) \dv{U}^\top\\
+  U^\top \dv{(f(Y))} U &= U^\top \dv{U} f(\Lambda) + f(\Lambda) \dv{U}^\top U + f'(\Lambda)\dv{\Lambda},
+\end{aligned}
+\]
+where $f'(\Lambda)_{ij} = \frac{\dv{(f(\lambda_i))}}{\dv{\lambda_i}}$.
+
+Recall that $U^\top \dv{U}$ is skew-symmetric.
+Note also that for skew-symmetrix $Z = -Z^\top$ and diagonal matrix $\operatorname{diag}(v)$,
+\[(Z\operatorname{diag}(v) + \operatorname{diag}(v)Z^\top)_{ij} = Z_{ij} v_j + v_i Z_{ji} = Z_{ij} v_j - v_i Z_{ij} = Z_{ij} (v_j - v_i) = Z \circ V,\]
+where $V_{ij} = v_j - v_i$ and $\circ$ is the Hadamard (elementwise) product.
+
+From the right-hand side, note then that
+\[(U^\top \dv{(f(Y))} U)_{ij} = \begin{cases} (U^\top \dv{U})_{ij} (f(\lambda_j) - f(\lambda_i)) & \text{ if } i \ne j \\ f'(\lambda_i) \dv{f(\lambda_i)} &\text{ if } i = j \end{cases}\]
+
+The exterior product of the diagonal part is
+\[
+  \bigwedge_{i=1}^N f'(\lambda_i) \dv{\lambda_i} = \left(\prod_{i=1}^N f'(\lambda_i)\right) (\dv{\Lambda})^\wedge,
+\]
+while the exterior product of the super-diagonal part is
+\[
+  \bigwedge_{j=1}^N \bigwedge_{i=j+1}^N (U^\top \dv{U})_{ij} (f(\lambda_j) - f(\lambda_i)) = \left( \prod_{j=1}^N \prod_{i=1}^{j-1} (f(\lambda_j) - f(\lambda_i)) \right) \left(U^\top \dv{U} \right)^\wedge
+\]
+
+Lastly, note for any symmetric $Z$ and orthogonal matrix $U$,
+\[(U^\top \dv{Z} U)^\wedge = \det(U^\top)^N \det(U)^N (\dv{Z})^\wedge = (\dv{Z})^\wedge.\]
+
+As a result, we can compute $(\dv(f(Y)))^\wedge$ by taking the exterior product of the super-diagonal and diagonal parts: 
+\[(\dv{(f(Y))})^\wedge = \left(\prod_{j=1}^N f'(\lambda_j) \prod_{i=1}^{j-1} (f(\lambda_j) - f(\lambda_i))\right) ((U^\top \dv{U})^\wedge \wedge \dv{\Lambda})\]
+
+If $f$ is the identity map $f(Y) = Y$, then
+\[ (\dv{Y})^\wedge = \left(\prod_{j=1}^N \prod_{i=1}^{j-1} (\lambda_j - \lambda_i)\right) \left((U^\top \dv{U})^\wedge \wedge \dv{\Lambda}\right)\]
+
+Combining the latter two expressions:
+\[(\dv{(f(Y))})^\wedge = \left(\prod_{j=1}^N f'(\lambda_j) \prod_{i=1}^{j-1} \frac{f(\lambda_j) - f(\lambda_i)}{\lambda_j - \lambda_i}\right) (\dv{Y})^\wedge\]
+
+[REF, Edelman's notes] previously found this result.
+
+For $X = f(Y) = \Exp(Y)$ then,
+\[ (\dv{X})^\wedge = \left(\prod_{j=1}^N e^{\lambda_j} \prod_{i=1}^{j-1} \frac{e^{\lambda_j} - e^{\lambda_i}}{\lambda_j - \lambda_i}\right) (\dv{Y})^\wedge,\]
+
+so the Jacobian determinant of the exponential map is $\absdet{J_{\Exp}(Y)} = \prod_{j=1}^N e^{\lambda_j} \prod_{i=1}^{j-1} \left|\frac{e^{\lambda_j} - e^{\lambda_i}}{\lambda_j - \lambda_i}\right|$.
+
+Note that this expression is undefined when any eigenvalues are repeated.
+This is not a problem for MCMC, since the eigenvalues are non-unique with probability 0, but non-unique eigenvalues could occur when maximizing a log-density.
+
+Because $\lim_{h \to 0} \frac{f(x) - f(x + h)}{h} = f'(x)$, we can then modify the expression to handle such cases:
+\[
+  \absdet{J_{\Exp}(Y)} = \prod_{j=1}^N \prod_{i=1}^j \begin{cases}
+    \left|\frac{e^{\lambda_j} - e^{\lambda_i}}{\lambda_j - \lambda_i}\right| &\text{ if } \lambda_i \ne \lambda_j \\ 
+    e^{\lambda_i} &\text{ if } \lambda_i = \lambda_j
+  \end{cases}.
+\]
+
+While the cost of computing the Jacobian determinant is $O(N^2)$, the cost of computing the eigendecomposition naively is $O(N^3)$, so this transformation is expensive.
+An attractive alternative would be to define the transform directly from the eigendecomposition of $X$.
+
+\subsubsection{Eigendecomposition parameterization}
+
+To define a transform directly from the symmetric eigendecomposition of $X$, we first need to note that the eigendecomposition is not unique; if we swap any two columns $i$ and $j$ of $U$ as well as swapping the eigenvalues $\lambda_i$ and $\lambda_j$, the result is the same matrix.
+
+When the eigenvalues are all unique, we can make the eigendecomposition also unique by enforcing the ordering of eigenvalues $\lambda_i > \lambda_j > 0$ for $j > i$.
+This is the ordering generally used for symmetric singular value decomposition.
+
+So we can construct a bijective map with the maps $f_1$ that produces an ordered vector of positive eigenvalues and $f_2$ that produces an orthogonal matrix.
+Consider $\lambda = f_1(y_1)$ and $U = f_2(y_2)$.
+
+Let $f_1$ be the map that computes each $\lambda_i$ as
+\[
+  \lambda_i = \begin{cases}
+    e^{(y_1)_N} &\text{ if } i = N\\
+    \lambda_{i-1} + e^{(y_1)_i} &\text{ if } 1 \le i < N
+  \end{cases}
+\]
+The Jacobian determinant of this map is $\absdet{J_{f_1}(y_1)} = \exp\left(\sum_{i=1}^N (y_1)_i\right)$.
+
+We can then choose for $f_2$ any of the transforms to the orthogonal matrices previously discussed.
+The combined Jacobian determinant of $f: (y_1, y_2) \mapsto f_2(y_2) \operatorname{diag}(f_1(y_1)) f_2(y_2)^\top = X$ is then 
+\[ \absdet{J_{f}(y_1, y_2)} = \absdet{J_{f_2}(y_2)} \exp\left(\sum_{i=1}^N (y_1)_i\right).\]
+
 \subsubsection*{Acknowledgements}

### 2022-07-18 tex/paper.tex
Draft section on positive-definite matrices

@@
 \cite{neal2011mcmc} or unit vectors \cite{byrne2013geodesic}, it is
-far more challenging to do the same for complex constraints such as
-simplices or positive definite matrices or for densities involving
-multiple constrained values.  Instead, usually constrained values are mapped to unconstrained values before sampling
-\cite{}.
-A change of variables adjustment is made to the target density for these transforms to work.
-The tricky part is- there can be more than one suitable mapping for every constraint.
-This leads us to the issue of selecting a transform.
+much more challenging to do the same for complex constraints such as
+simplexes or positive definite matrices or for densities involving
+multiple constrained values.  Instead, it is far more common to map
+the constrained values to unconstrained values before sampling
+\cite{}.  This presents the issue of selecting which transform to use
+among an infinite set of options, which is the topic of this paper.
 
-\section{Transforms}
+%Sampling algorithms play an important role in Statistics, in this case
+%we refer to drawing samples from a probability density or distribution
+%(not survey sampling). Monte Carlo Markov chain methods are commonly
+%used for high dimensional sampling. In this paper, we look at
+%Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
+%samples for the Metropolis algorithm. Bayesian Inference often uses
+%parameters with constraints, but HMC optimizes on the unconstrained
+%space. It is difficult to use Monte Carlo estimation on a constrained
+%domain ~\cite{neal2008optimal}
 
-To draw samples from a distribution $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.
-
-Let $f\colon \mathcal{Y} \to \mathcal{X}$ be a smooth and continuous map.
-Only when $m = n$ can $f$ be bijective.
-When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = f(y)$.
-In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
-We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
-$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
-that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
-When $h$ is non-bijective at single points, we call these singularities.
-While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.
-
-\subsection{Changes of Variables}
+\section{Changes of Variables}
 
@@
 
-\subsection{Target Density}
+\section{Transforms}
+
+To draw samples from a distribution $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.
+
+Let $f\colon \mathcal{Y} \to \mathcal{X}$ be a smooth and continuous map.
+Only when $m = n$ can $f$ be bijective.
+When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = f(y)$.
+In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
+We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
+$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
+that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
+When $h$ is non-bijective at single points, we call these singularities.
+While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.
+
 Given a proper joint distribution with a smooth and continuous density $p_{X,Y}(x, y | \theta)$ for some parameters $\theta$, we can use the usual change of variables formula to write
@@
 
-The StickBreaking transform can be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}.
+The StickBreaking transform ican be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}.
 Intuitively, this comprises of recursively breaking a piece $x_i$ from a stick of unit length, where the leftover stick in the $i^{th}$ iteration is $ 1 - \sum_{1}^{i}x$.
@@
   \\[6pt]
+  & = &
+  \left(
+    1 
+    + \textrm{sum}\left( \frac{-(1 + \textrm{sum}(z)) z}{(1 +
+        \textrm{sum}(z))^2} \right)
+  \right)
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
+  & = &
+        \left(1 + \textrm{sum}\left(\frac{-z}{1 + \textrm{sum}(z)} \right)\right)        
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
+  & = & \left( 1 - \textrm{sum}(\textrm{norm}(z)) \right) 
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
   & = & \left( \frac{1}{1 + \textrm{sum}(z)} \right)^N.
@@
 \]
+
+%\subsection{Softmax Transform}
+%
+%The softmax function can be understood from Multinomial Logistic
+%Regression employed for predicting probabilities of a Categorically
+%distributed variable. Geometrically it maps $R^K$ to the boundary of a
+%unit K-simplex(it is simply the convex hull of $k+1$ affinely
+%independent points in $R^K$. Essentially it transforms a vector of
+%size K to another vector of size K where the outputs sum to 1. It is
+%worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
+%when a vector of size $K$ is transformed the $K_th$ vector is simply
+%$1$- sumof k-1 vectors
+
+\subsection{Simplex softmax parameterization}
+
+Let $\Delta^n$ indicate the unit $n$-simplex with $n$ positive
+elements that sum to 1 and $\Delta^n_-$ indicate the first $n-1$
+elements, from which the final element can be uniquely determined.
+
+We define the transformation
+$\phi: \mathbb{R}^{n-1} \to \Delta^n_-: y \mapsto x_-$, where
+$x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix} \in \Delta^n$,
+$x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$, and
+$r = 1 + \sum_{i=1}^{n-1} e^{y_i}$.
+
+First we compute the scalar derivatives:
+\[
+\begin{aligned}
+  \frac{\mathrm{d} r}{\mathrm{d} y_j}
+  &= e^{y_j} = r x_j\\
+  \frac{\mathrm{d} x_i}{\mathrm{d} y_j}
+  &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i}
+  \frac{\mathrm{d} r}{\mathrm{d} y_j}
+  = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
+\end{aligned}
+\]
+where
+$\delta_{ij} = \begin{cases} 1 &\text{if } i = j \\ 0
+  &\text{otherwise}\end{cases}$ is the Kronecker delta.
+
+If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the
+elements of $x$, then the Jacobian is
+\[
+  J = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top) \operatorname{diag}(x_-),
+\]
+where $\boldsymbol{1}_n$ is the $n$-vector of ones.
+
+Using Sylvester's determinant theorem,
+$|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = 1 -
+\boldsymbol{1}_{n-1}^\top x_- = 1 - \sum_{i=1}^{n-1} x_i = x_n$, so
+$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} y_i\right) \left(1 + \sum_{i=1}^{n-1} e^{y_i}\right)^{-n}$$
 
@@
 $\begin{pmatrix}A & B \\ C & D\end{pmatrix}$ is $|A| |D-CA^{-1}B|$.  A
-square matrix is invertible iff its determinant is non-zero.  From the
+square matrix is invertible iff its determinant is non-yero.  From the
 previous section,

### 2022-07-19 tex/paper.tex
cleaning up paper a bit

--- 2022-07-18 tex/paper.tex
@@
 \newcommand{\absdet}[1]{\abs{#1}}
-\newcommand{\dv}[1]{\mathrm{d}{#1}}
-\newcommand{\vect}{\mathrm{vec}}
-\newcommand{\vectu}{\mathrm{vecu}}
 
-% transforms -> change of variables
-% two kinds of transforms
-% pragprg
 \begin{document}
@@
 
-\section{Preliminaries}
-
-[SA: TODO: eliminate or merge with following section]
-
-To compute Jacobians of functions with matrix-valued inputs or outputs, we need to implicitly or explicitly choose for the inputs and outputs a set of coordinates with dimension equal to the number of degrees of freedom of the matrix
-Unless otherwise specified, we choose the coordinates to be the vectorization of the matrix.
-
-We define the (bijective) vectorization of a matrix $X \in \mathbb{R}^{N \times K}$ with columns $x_k$ for $k \in \{1, ..., K\}$ as the vector
-\[\vect(X) = \begin{bmatrix}x_1 \\ x_2 \\ \vdots \\ x_M\end{bmatrix}\]
-formed by stacking the columns of $X$.
-
-For $X \in \mathbb{R}^{N \times K}$, let $\vectu(X)$ be the map that stacks only the parts of the columns of $X$ above and including the diagonal.
-This map is sometimes called the half-vectorization and is bijective for upper-triangular or symmetric $X$.
-
-Similarly, $\vectu_-(X)$ extracts the same elements except the diagonal.
-When $X$ is strictly upper-triangular or skew-symmetric (i.e. $X = -X^\top$), then this map is bijective.
-
-\section{Computing Jacobian determinants}
-
-[SA: TODO: remove everything we do not use later, add strategy for making non-square Jacobian square]
-
-Here we introduce several strategies for computing Jacobian determinants.
-
-The vectorization has the useful property that for all matrices $A$, $X$, and $B$, $\vect(AYB) = (B^\top \otimes A) \vect(Y)$, where $\otimes$ is the Kronecker product.
-This allows us to convert expressions of matrices to expressions of vectors, from which the Jacobian can be read off.
-For example, let $X = AYB$. Then $\dv{X} = A \dv{Y} B$, and $\vect(\dv{X}) = (B^\top \otimes A) \vect(\dv{Y})$, and the Jacobian is $J = B^\top \otimes A$.
-
-When the previously discussed strategies are insufficient to compute the Jacobian, we bypass the Jacobian computation entirely and use the exterior product $\wedge$.
-For example, let $X \in \mathbb{R}^{N \times K}$.
-The exterior product
-\[
-  \dv{X_{11} }\wedge \dv{X_{21}} \wedge \ldots \wedge \dv{X_{NK}} = \bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{X}_{ij}
-\]
-is another more explicit way of writing the integration measure $\dv{X_{11}} \dv{X_{21}} \dots \dv{X_{NK}}$ that has useful properties for computing Jacobian determinants.
-The basic properties we will use are
-
-\begin{itemize}
-  \item $a\dv{x} \wedge b\dv{y} = (ab)(\dv{x} \wedge \dv{y})$ for $a,b \in \mathbb{R}$
-  \item $(\dv{x} + \dv{y}) \wedge \dv{z} = \dv{x} \wedge \dv{z} + \dv{y} \wedge \dv{z}$
-  \item $\dv{x} \wedge \dv{y} = - \dv{y} \wedge \dv{x}$
-  \item $\dv{x} \wedge \dv{x} = 0$
-\end{itemize}
-
-The connection between the exterior product and the Jacobian determinant is given by the following expression.
-Given $f: y \mapsto x$ for $x,y \in \mathbb{R}^N$,
-\[\bigwedge_{i=1}^N \dv{x_i} = \det(J_f(y)) \bigwedge_{i=1}^N \dv{y_i}.\]
-
-We can then find the Jacobian determinant by first writing all differential forms $\dv{x_i}$ or $\dv{y_i}$ and then computing the exterior product until the expression looks like the above.
-For our purposes, we will ignore the sign of the Jacobian determinant.
-
-A related useful property is that for $X,Y,B \in \mathbb{R}^{N \times K}$, if $\dv{X} = B \dv{Y}$, then
-\[\bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{X_i} = \det(B)^K \bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{Y_i}\]
-
-Following [REF, Edelman], we use the notation $(\cdot)^\wedge$ as a shorthand to represent the exterior product of the unique elements of $\cdot$.
-
-For example,
-\begin{itemize}
-  \item Rectangular matrix $\{X \in \mathbb{R}^{N \times K}\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{X_{ij}}$
-  \item Upper-triangular rectangular matrix $\{X \in \mathbb{R}^{N \times K} | X_{ij} = 0\ \forall\ i > j\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{\min(j, N)} \dv{X_{ij}}$
-  \item Diagonal matrix $\{X \in \mathbb{R}^{N \times K} | X_{ij} = 0\ \forall\ i \ne j\}$: $(\dv{X})^\wedge = \bigwedge_{i=1}^{\min(N, K)} \dv{X_{ii}}$
-  \item Symmetric matrix $\{X \in \mathbb{R}^{N \times N} | X_{ij} = X_{ji}\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{j} \dv{X_{ij}}$
-  \item Skew-symmetric matrix $\{X \in \mathbb{R}^{N \times N} | X_{ij} = -X_{ji}, X_{ii} = 0\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{j-1} \dv{X_{ij}}$
-\end{itemize}
-
-Using this notation, for example, $(B \dv{X})^\wedge = \det(B)^K (\dv{X})^\wedge$
-
 \section{Unit simplex}
@@
 
-\subsection{Semi-orthogonal matrices (Stiefel manifold)}
-
-A matrix $X \in \mathbb{R}^{N \times K}$ for $K \le N$ is semi-orthogonal if and only if $X^\top X = I_K$.
-The semi-orthogonal matrices form a space called the Stiefel manifold.
-When $K=N$, $X \in O(N)$ is an orthogonal matrix and also satisfies $X X^\top = I_N$ and $\det(X) = \pm 1$.
-When $K=1$, $X$ is a point on the $(N-1)$-sphere $\mathbb{S}^{N-1} \subset \mathbb{R}^N$.
-
-By choosing a set of rules such as Gram-Schmidt orthonormalization of the canonical basis in $\mathbb{R}^{N \times N}$, it is possible to analytically compute from $X$ an orthogonal complement $X^\perp \in \mathbb{R}^{N \times (N - K)}$, such that $\tilde{X}^\top \tilde{X} = I_N$, where $\tilde{X} = \begin{bmatrix}X & X^\perp \end{bmatrix} \in O(N)$.
-
-Note that if we differentiate both sides of the constraint $X^\top X = I_K$, we get $\dv{X}^\top X + X^\top \dv{X} = 0$, which leads to $(X^\top \dv{X})^\top = - X^\top \dv{X}$.
-Therefore, $X^\top \dv{X}$ is a $K \times K$ skew-symmetric matrix with $\frac{K(K-1)}{2}$ unique entries in the strict upper triangle.
-
-Densities of distributions on the Stiefel manifold are usually written with respect to the invariant metric on the Stiefel manifold, which we will define separately for the two cases $K=N$ and $K < N$.
-
-For $X \in O(N)$ with the $i$th column written $x_i$, we define the invariant measure as the exterior product of the unique elements of this skew-symmetric matrix:
-\[
-  (X^\top \dv{X})^\wedge \equiv \bigwedge_{j=1}^N \bigwedge_{i=j+1}^N (X^\top \dv{X})_{ij} = \bigwedge_{j=1}^N \bigwedge_{i=j+1}^N x_i^\top \dv{x}_j.
-\]
-This measure is called invariant because if $f(X) = A X B$ for $A,B \in O(N)$ and $\mathcal{S} \subset O(N)$ then $\int_{\mathcal{S}} (X^\top \dv{X}) = \int_{f(\mathcal{S})} (X^\top \dv{X})$.
-It is also called the Haar measure on $O(N)$ \cite{muirhead2009aspects}..
-
-To choose an invariant measure for $K < N$, we first choose an $X^\perp$ analytically computed from $X$.
-Because $X^\perp$ is analytically computed, $\dv{X^\perp_{ij}} = 0$.
-We can then construct the invariant measure for $K<N$ from the invariant measure for $O(N)$:
-
-\[
-  (X^\top \dv{X})^\wedge \equiv \bigwedge_{j=1}^N \bigwedge_{i=j+1}^N x_i^\top \dv{x}_j = \bigwedge_{j=1}^K \bigwedge_{i=j+1}^N x_i^\top \dv{x}_j,
-\]
-
-Though we needed to choose a $X^\perp$, the invariant measure does not depend on the choice of $X^\perp$.
-Moreover, this measure is invariant to left-actions by $O(N)$ and right-actions by $O(K)$.
-For more details, see \cite{muirhead2009aspects}.
-
-When $K=1$, the invariant measure is equivalent to the usual Hausdorff measure on the sphere.
-
-\subsubsection{Givens parameterization of Stiefel manifold}
-
-\subsubsection{Cayley parameterization of Stiefel manifold}
-
-\subsubsection{Householder parameterization of Stiefel manifold}
-
-\subsubsection{Matrix exponential parameterization of Orthogonal group}
-
 \subsubsection*{Acknowledgements}

### 2022-07-19 simplex.py
maintenance (#32)

@@ -1,0 +1,21 @@
+import os
+from tqdm import tqdm
+os.chdir('transforms')
+
+import sys
+sys.path.insert(1, 'utils')
+
+from sample import sample
+
+parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*100, 'N': 100}, {'alpha': [0.1]*1000, 'N': 1000},
+{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000},
+{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000}]
+
+for transform in tqdm(transforms):
+    for params in tqdm(parameters):
+        sample(transform_category='simplex', transform=transforms, 
+            evaluating_model='dirichlet_symmetric', parameters=[params], output_file='/mnt/ceph/users/mjhajaria/sampling_results/stan_1.json', 
+            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
+                            show_progress = True, resample=True, return_idata=False)
+
+

### 2022-07-19 utils/dirichlet_symmetric.py

@@ -4,52 +4,30 @@ import matplotlib.pyplot as plt
 from rmse import rmse_leapfrog
 from sample import sample
 
+def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name, n_repeat=1, n_iter=1000, 
+                                n_chains=4, show_progress=True, resample=False):
+    
+    plt.rcParams["figure.figsize"] = [20,10]
+    plt.rcParams['figure.dpi'] = 300
+    fig, axes = plt.subplots(3,3)
+    fig.supxlabel('Cumulative Leapfrog Steps')
+    fig.supylabel('Root Mean Squared Error')
 
-def get_dirichlet_symmetric_rmse(
-    transforms,
-    transform_category,
-    parameters,
-    fig_name,
-    n_repeat=1,
-    n_iter=1000,
-    n_chains=4,
-    show_progress=True,
-    resample=False,
-):
+    for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
+        for transform in transforms:            
+            idata = sample(transform_category=transform_category, transform=transform, 
+                evaluating_model='dirichlet_symmetric', parameters=[params], output_file=None, 
+                auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
+                                        show_progress = show_progress, resample=resample)
+            alpha = params['alpha']
+            N = params['N']
+            true_x = [a/sum(alpha) for a in alpha]
+            x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
+            ax.plot(x,y, label = str(transform))
 
-    plt.rcParams["figure.figsize"] = [20, 10]
-    plt.rcParams["figure.dpi"] = 300
-    fig, axes = plt.subplots(len(parameters))
-    fig.supxlabel("Cumulative Leapfrog Steps")
-    fig.supylabel("Root Mean Squared Error")
-
-    for ax, params in zip(
-        axes.flatten() if len(parameters) > 1 else [axes], parameters
-    ):
-        for transform in transforms:
-            idata = sample(
-                transform_category=transform_category,
-                transform=transform,
-                evaluating_model="dirichlet_symmetric",
-                parameters=[params],
-                output_file=None,
-                auto_eval_all_params=False,
-                n_iter=n_iter,
[…]

### 2022-07-19 utils/sample.py
maintenance (#32)

@@ -85,8 +85,8 @@ def sample(
 
     if resample == False:
         for params in parameters:
-            filename = f"sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.json"
-            return az.from_json(filename)
+            filename = f'sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            return az.load_arviz_data(filename)
     else:
 
         if auto_eval_all_params:
@@ -94,10 +94,9 @@ def sample(
                 f"target_densities/{evaluating_model}_parameters.json", "rb"
             ) as f:
                 parameters = pickle.load(f)
-
-        stan_filename = f"stan_models/{transform}_{evaluating_model}.json"
-
-        if type(parameters) == dict:
+        stan_filename=f'stan_models/{transform}_{evaluating_model}.stan'
+        
+        if type(parameters)==dict:
             parameters = [parameters]
         for params in tqdm(parameters):
             model = CmdStanModel(
@@ -121,13 +120,9 @@ def sample(
                 )
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 
-            filename = (
-                output_file
-                if output_file
-                else f"sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.json"
-            )
-            idata.to_json(filename)
-            if return_idata == True:
-                return idata
+            filename = output_file if output_file else f'sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            idata.to_netcdf(filename)
+        if return_idata==True:
+            return idata
 
         pass

### 2022-07-19 simplex.py
maintenance (#32)
more maintenance

vectorize

misc

misc

-

 -

@@ -1,0 +1,18 @@
+import os
+from tqdm import tqdm
+os.chdir('transforms')
+
+import sys
+sys.path.insert(1, 'utils')
+
+from sample import sample
+
+transforms=['stan', 'stickbreaking', 'softmax', 'softmax-augmented']
+
+for transform in tqdm(transforms):
+        sample(transform_category='simplex', transform=transform, 
+            evaluating_model='dirichlet_symmetric', parameters={}, 
+            auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 1, 
+                            show_progress = True, resample=True, return_idata=False)
+
+

### 2022-07-19 utils/dirichlet_symmetric.py

@@ -4,52 +4,30 @@ import matplotlib.pyplot as plt
 from rmse import rmse_leapfrog
 from sample import sample
 
+def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name, n_repeat=1, n_iter=1000, 
+                                n_chains=4, show_progress=True, resample=True):
+    
+    plt.rcParams["figure.figsize"] = [20,10]
+    plt.rcParams['figure.dpi'] = 300
+    fig, axes = plt.subplots(3,3)
+    fig.supxlabel('Cumulative Leapfrog Steps')
+    fig.supylabel('Root Mean Squared Error')
 
-def get_dirichlet_symmetric_rmse(
-    transforms,
-    transform_category,
-    parameters,
-    fig_name,
-    n_repeat=1,
-    n_iter=1000,
-    n_chains=4,
-    show_progress=True,
-    resample=False,
-):
+    for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
+        for transform in transforms:            
+            idata = sample(transform_category=transform_category, transform=transform, 
+                evaluating_model='dirichlet_symmetric', parameters=[params], 
+                auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
+                                        show_progress = show_progress, resample=resample)
+            alpha = params['alpha']
+            N = params['N']
+            true_x = [a/sum(alpha) for a in alpha]
+            x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
+            ax.plot(x,y, label = str(transform))
 
-    plt.rcParams["figure.figsize"] = [20, 10]
-    plt.rcParams["figure.dpi"] = 300
-    fig, axes = plt.subplots(len(parameters))
-    fig.supxlabel("Cumulative Leapfrog Steps")
-    fig.supylabel("Root Mean Squared Error")
-
-    for ax, params in zip(
-        axes.flatten() if len(parameters) > 1 else [axes], parameters
-    ):
-        for transform in transforms:
-            idata = sample(
-                transform_category=transform_category,
-                transform=transform,
-                evaluating_model="dirichlet_symmetric",
-                parameters=[params],
-                output_file=None,
-                auto_eval_all_params=False,
-                n_iter=n_iter,
[…]

### 2022-07-19 utils/sample.py

@@ -20,7 +20,7 @@ def sample(
     transform,
     evaluating_model,
     parameters,
-    output_file=None,
+    output_dir='/mnt/sdceph/users/mjhajaria',
     auto_eval_all_params=False,
     n_iter=1000,
     n_chains=4,
@@ -85,8 +85,8 @@ def sample(
 
     if resample == False:
         for params in parameters:
-            filename = f"sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.json"
-            return az.from_json(filename)
+            filename = f'sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            return az.load_arviz_data(filename)
     else:
 
         if auto_eval_all_params:
@@ -94,10 +94,9 @@ def sample(
                 f"target_densities/{evaluating_model}_parameters.json", "rb"
             ) as f:
                 parameters = pickle.load(f)
-
-        stan_filename = f"stan_models/{transform}_{evaluating_model}.json"
-
-        if type(parameters) == dict:
+        stan_filename=f'stan_models/{transform}_{evaluating_model}.stan'
+        
+        if type(parameters)==dict:
             parameters = [parameters]
         for params in tqdm(parameters):
             model = CmdStanModel(
@@ -121,13 +120,9 @@ def sample(
                 )
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 
-            filename = (
-                output_file
-                if output_file
-                else f"sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.json"
-            )
-            idata.to_json(filename)
-            if return_idata == True:
-                return idata
+            filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            idata.to_netcdf(filename)
+        if return_idata==True:
+            return idata
 
         pass

### 2022-07-19 simplex.py
vectorize
misc

misc

-

@@ -7,15 +7,12 @@ sys.path.insert(1, 'utils')
 
 from sample import sample
 
-parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*100, 'N': 100}, {'alpha': [0.1]*1000, 'N': 1000},
-{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000},
-{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000}]
+transforms=['stan', 'stickbreaking', 'softmax', 'softmax-augmented']
 
 for transform in tqdm(transforms):
-    for params in tqdm(parameters):
-        sample(transform_category='simplex', transform=transforms, 
-            evaluating_model='dirichlet_symmetric', parameters=[params], 
-            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
+        sample(transform_category='simplex', transform=transform, 
+            evaluating_model='dirichlet_symmetric', parameters={}, 
+            auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 1, 
                             show_progress = True, resample=True, return_idata=False)
 
 

### 2022-07-19 utils/dirichlet_symmetric.py

@@ -5,7 +5,7 @@ from rmse import rmse_leapfrog
 from sample import sample
 
 def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name, n_repeat=1, n_iter=1000, 
-                                n_chains=4, show_progress=True, resample=False):
+                                n_chains=4, show_progress=True, resample=True):
     
     plt.rcParams["figure.figsize"] = [20,10]
     plt.rcParams['figure.dpi'] = 300
@@ -16,7 +16,7 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
     for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
         for transform in transforms:            
             idata = sample(transform_category=transform_category, transform=transform, 
-                evaluating_model='dirichlet_symmetric', parameters=[params], output_file=None, 
+                evaluating_model='dirichlet_symmetric', parameters=[params], 
                 auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
                                         show_progress = show_progress, resample=resample)
             alpha = params['alpha']

### 2022-07-19 simplex.py
more maintenance

@@ -14,7 +14,7 @@ parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*100, 'N': 100}, {'al
 for transform in tqdm(transforms):
     for params in tqdm(parameters):
         sample(transform_category='simplex', transform=transforms, 
-            evaluating_model='dirichlet_symmetric', parameters=[params], output_file='/mnt/ceph/users/mjhajaria/sampling_results/stan_1.json', 
+            evaluating_model='dirichlet_symmetric', parameters=[params], 
             auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
                             show_progress = True, resample=True, return_idata=False)
 

### 2022-07-19 utils/sample.py

@@ -20,7 +20,7 @@ def sample(
     transform,
     evaluating_model,
     parameters,
-    output_file=None,
+    output_dir='/mnt/sdceph/users/mjhajaria',
     auto_eval_all_params=False,
     n_iter=1000,
     n_chains=4,
@@ -120,7 +120,7 @@ def sample(
                 )
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 
-            filename = output_file if output_file else f'sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
             idata.to_netcdf(filename)
         if return_idata==True:
             return idata

### 2022-07-20 tex/dummypaper.tex

--- 2022-07-18 tex/paper.tex
@@
 simplices or positive definite matrices or for densities involving
-multiple constrained values.  Instead, usually constrained values are mapped to unconstrained values before sampling
-\cite{}.
+multiple constrained values.
+Instead, usually constrained values are mapped to unconstrained values before sampling.
+These transform mappings are formed out of the constraints on the model parameters.
 A change of variables adjustment is made to the target density for these transforms to work.
@@
 This leads us to the issue of selecting a transform.
+
+We evaluate the efficiency of unconstraining parameter transforms using metrics like effective sample size, leapfrog steps and squared error.
+Different transforms work well for different models, this paper aims to provide a useful analysis of the same, so practitioners can choose transforms that work well for their models.
 

### 2022-07-20 tex/notes.tex

@@
-\documentclass[twoside]{article}
+Sampling algorithms play an important role in Statistics, in this case
+we refer to drawing samples from a probability density or distribution
+(not survey sampling). Monte Carlo Markov chain methods are commonly
+used for high dimensional sampling. In this paper, we look at
+Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
+samples for the Metropolis algorithm. Bayesian Inference often uses
+parameters with constraints, but HMC optimizes on the unconstrained
+space. It is difficult to use Monte Carlo estimation on a constrained
+domain ~\cite{neal2008optimal}
 
-\usepackage{aistats2021}
-\usepackage{amsmath}
-% If your paper is accepted, change the options for the package
-% aistats2021 as follows:
-%
-%\usepackage[accepted]{aistats2021}
-%
-% This option will print headings for the title of your paper and
-% headings for the authors names, plus a copyright note at the end of
-% the first column of the first page.
+The softmax function can be understood from Multinomial Logistic
+Regression employed for predicting probabilities of a Categorically
+distributed variable. Geometrically it maps $R^K$ to the boundary of a
+unit K-simplex(it is simply the convex hull of $k+1$ affinely
+independent points in $R^K$. Essentially it transforms a vector of
+size K to another vector of size K where the outputs sum to 1. It is
+worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
+when a vector of size $K$ is transformed the $K_th$ vector is simply
+$1$- sum of k-1 vectors
 
-% If you set papersize explicitly, activate the following three lines:
-%\special{papersize = 8.5in, 11in}
-%\setlength{\pdfpageheight}{11in}
-%\setlength{\pdfpagewidth}{8.5in}
+\subsection{Simplex softmax parameterization}
 
-% If you use natbib package, activate the following three lines:
-%\usepackage[round]{natbib}
-%\renewcommand{\bibname}{References}
-%\renewcommand{\bibsection}{\subsubsection*{\bibname}}
+Let $\Delta^n$ indicate the unit $n$-simplex with $n$ positive
+elements that sum to 1 and $\Delta^n_-$ indicate the first $n-1$
+elements, from which the final element can be uniquely determined.
 
-% If you use BibTeX in apalike style, activate the following line:
-%\bibliographystyle{apalike}
+We define the transformation
+$\phi: \mathbb{R}^{n-1} \to \Delta^n_-: y \mapsto x_-$, where
+$x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix} \in \Delta^n$,
+$x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$, and
+$r = 1 + \sum_{i=1}^{n-1} e^{y_i}$.
 
-\begin{document}
+First we compute the scalar derivatives:
+\[
+\begin{aligned}
+  \frac{\mathrm{d} r}{\mathrm{d} y_j}
+  &= e^{y_j} = r x_j\\
+  \frac{\mathrm{d} x_i}{\mathrm{d} y_j}
+  &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i}
+  \frac{\mathrm{d} r}{\mathrm{d} y_j}
+  = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
+\end{aligned}
+\]
+where
+$\delta_{ij} = \begin{cases} 1 &\text{if } i = j \\ 0
+  &\text{otherwise}\end{cases}$ is the Kronecker delta.
 
-% If your paper is accepted and the title of your paper is very long,
-% the style will print as headings an error message. Use the following
-% command to supply a shorter title of your paper so that it can be
-% used as headings.
-%
-%\runningtitle{I use this title instead because the last one was very long}
+If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the
+elements of $x$, then the Jacobian is
+\[
+  J = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top) \operatorname{diag}(x_-),
+\]
+where $\boldsymbol{1}_n$ is the $n$-vector of ones.
 
-% If your paper is accepted and the number of authors is large, the
-% style will print as headings an error message. Use the following
-% command to supply a shorter version of the authors names so that
-% they can be used as headings (for example, use only the surnames)
-%
-%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
+Using Sylvester's determinant theorem,
+$|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = 1 -
+\boldsymbol{1}_{n-1}^\top x_- = 1 - \sum_{i=1}^{n-1} x_i = x_n$, so
+$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} y_i\right) \left(1 + \sum_{i=1}^{n-1} e^{y_i}\right)^{-n}$$
 
-% Supplementary material: To improve readability, you must use a single-column format for the supplementary material.
-\onecolumn
-\aistatstitle{Notes}
+\subsubsection{Softmax Transform}
 
-\section{Introduction}
-
-\section{Constraints}
-\subsection{Bounded Constraints}
-\subsection{Simplex Constraints}
-\subsection{Definite Matrix constraints}
-\section{Softmax Transform}
-
-The softmax function can be understood from Multinomial Logistic Regression employed for predicting probabilities of a Categorically distributed variable.
-Geometrically it maps $R^K$ to the boundary of a unit K-simplex(it is simply the convex hull of $k+1$ affinely independent points in $R^K$.
-Essentially it transforms a vector of size K to another vector of size K where the outputs sum to 1.
-It is worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so when a vector of size $K$ is transformed the $K_th$ vector is simply $1$- sumofk-1 vectors
-R
-
-\end{document}
+To calculate the determinant of the Jacobian of the inverse transform,
+we start by noting that $s = \textrm{exp} \circ \textrm{norm}$, where
+$\textrm{exp}$ is the elementwise exponential function and
+\textrm{norm} is defined by
+\[
+  \textrm{norm}(z) = \frac{z}{\textrm{sum}(z) + 1}.
+\]
+As such, the resulting Jacobian determinant is the product of the
+Jacobian determinants of the component functions,
+\[
+  \absdet{J_s(y)}
+  = \absdet{J_{\textrm{exp}}(y)} \absdet{J_{\textrm{norm}}(z)},
+\]
+where $z = \textrm{exp}(y)$.  The Jacobian for the exponential
+function is diagonal, so the determinant is the product of the
+diagonal of the Jacobian, which for $y \in \mathbb{R}^{N-1}$ is
+\[
+  \absdet{J_{\textrm{exp}}(y)} = \textrm{prod}(\exp(y)).
+\]
+As above, let $z = \exp(y) \in (0, \inf)^{N-1}$.  We can differentiate
+$\textrm{norm}$ to derive the Jacobian,
+\[
+  J_{\textrm{norm}}
+  = \frac{1}{1 + \textrm{sum}(z)} \mathbb{I}_{N-1}
+  - \left(\frac{1}{(1 + \textrm{sum}(z))^2} \beta \right)
+  \textrm{vector}_{N-1}(1)^{\top},
+\]
+where $\mathbb{I}_{N-1}$ is the $(N - 1) \times (N - 1)$ unit matrix and
+$\textrm{vector}_{N-1}(1)$ is the $N - 1$-vector with values 1.  Using
+the matrix determinant lemma,\footnote{The matrix determinant lemma
+  is \[\textrm{det}(A + u v^{\top}) = (1 + v^{\top} A^{-1} u)
+    \textrm{det}(A).\]}
+we have
+\begin{eqnarray*}
+  \textrm{absdet}(J_{\textrm{norm}}(z))
+  & = &
+  \left(
+    1
+    + \textrm{vector}_{N-1}(1)^{\top}
+    \left(\frac{1}{1 + \textrm{sum}(z)} \mathbb{I} \right)^{-1}
+    \frac{-z}{(1 + \textrm{sum}(z))^2}
+    \right)
+    \ \textrm{det}\left(\frac{1}{1 + \textrm{sum}(z)} \mathbb{I}
+        \right)
+  \\[6pt]
+  & = &
+  \left(
+    1 
+    + \textrm{sum}\left( \frac{-(1 + \textrm{sum}(z)) z}{(1 +
+        \textrm{sum}(z))^2} \right)
+  \right)
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
+  & = &
+        \left(1 + \textrm{sum}\left(\frac{-z}{1 + \textrm{sum}(z)} \right)\right)        
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
+  & = & \left( 1 - \textrm{sum}(\textrm{norm}(z)) \right) 
+        \ \left( \frac{1}{1 + \textrm{sum}(z)} \right)^{N-1}
+  \\[6pt]
+  & = & \left( \frac{1}{1 + \textrm{sum}(z)} \right)^N.
+\end{eqnarray*}
+Thus the entire absolute determinant of the Jacobian is defined by the
+product, 
+\[
+  \absdet{J_s(y)}
+  \ = \
+  \textrm{prod}(\exp(y))
+  \, \left( \frac{1}{1 + \textrm{sum}(\exp(y))} \right)^N.
+\]
+and our final expression for densities for unconstrained $y \in
+\mathbb{R}^{N-1}$ is
+\[
+  p_Y(y)
+  = p_X(\textrm{alr}^{-1}(y))
+  \, \textrm{prod}(\exp(y))
+  \left( \frac{1}{1 + \textrm{sum}(\exp(y))} \right)^N
+\]

### 2022-07-20 tex/dummypaper.tex
make paper for simplex and add randomseed

--- 2022-07-19 tex/paper.tex
@@
 \newcommand{\absdet}[1]{\abs{#1}}
+\newcommand{\dv}[1]{\mathrm{d}{#1}}
+\newcommand{\Exp}{\mathrm{Exp}}
+\newcommand{\vect}{\mathrm{vec}}
+\newcommand{\vectu}{\mathrm{vecu}}
 
@@
 
+We evaluate the efficiency of unconstraining parameter transforms using metrics like effective sample size, leapfrog steps and squared error.
+Different transforms work well for different models, this paper aims to provide a useful analysis of the same, so practitioners can choose transforms that work well for their models.
+
 \section{Transforms}
+The transform from a constrained space $\mathcal{X}$ to an unconstrained space $\mathcal{Y}$ is a smooth and bijective function $f: \mathcal{X} \to \mathcal{Y}$.
+We aim to induce a density $h(y)$ on $\mathcal{Y}$ and define a smooth and continous map $g\colon \mathcal{Y} \to \mathcal{X}$, such that $h(y)$ is uniform over $\mathcal{X}$.
+%review and make this better
+Uniformity is defined here to mean, if $y$, $y`$ $\sim$ $h(y)$,  then $g(y)$, $g(y`)$ $\sim$ uniform($\mathcal{X}$).
 
-To draw samples from a distribution $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.
+% check m,n relationship
+%To draw samples from a $y$ $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.
 
-Let $f\colon \mathcal{Y} \to \mathcal{X}$ be a smooth and continuous map.
-Only when $m = n$ can $f$ be bijective.
-When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = f(y)$.
-In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
-We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
-$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
-that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
-When $h$ is non-bijective at single points, we call these singularities.
-While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.
+%Only when $m = n$ can $f$ be bijective.
+%When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = g(y)$.
+
+%In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
+%We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
+%$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
+%that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
+%When $h$ is non-bijective at single points, we call these singularities.
+%While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.
 
@@
 
-If $X$ is a random variable with density $p_X$ and $Y = f(X)$ for a
-smooth and bijective function $f$, then
+If $Y \in \mathbb{R^N}$ is a random variable with density $p_Y(y)$, where $Y = g^{-1}(X)$, where $X \in \mathbb{R^M}$.
+Then for a smooth and bijective function $g$ which is the inverse transform from an unconstrained space to a constrained space, we can define its density as:
 \[
-  p_Y(y) = p_X(f^{-1}(y)) \absdet{J_{f^{-1}}(y)},
+  p_Y(y) = p_X(g(y)) \absdet{J_{g}(y)},
 \]
@@
 \[
-  J_{f^{-1}}(y) = \frac{\partial}{\partial y} \, f^{-1}(y)
+  J_{g}(y) = \frac{\partial}{\partial y} \, g(y)
 \]
@@
 \[
-  \absdet{J_{f^{-1}}(y)}
-  = \abs{\det J_{f^{-1}}(y)}.
+  \absdet{J_{g}(y)}
+  = \abs{\det J_{g}(y)}.
 \]
-
-\subsection{Target Density}
-Given a proper joint distribution with a smooth and continuous density $p_{X,Y}(x, y | \theta)$ for some parameters $\theta$, we can use the usual change of variables formula to write
-
-\[
-  p_Y(y | \theta) = p_{X,Z}(f(y), g(y) | \theta) |J_h(y)|.
-\]
-
-To sample from a target density $p_X(x | \theta)$, let $p_{X,Z}(x, z | \theta) = p_X(x | \theta) p_Z(z | x, \theta)$.
-Then we must choose some proper distribution $p_Z$ that is smooth and continuous.
-Given that choice, we find that
-\[
-  p_Y(y | \theta) = p_X(f(y) | \theta) p_Z(g(y) | f(y), \theta) |J_h(y)|.
-\]
-
-Note that when $m > n$, it is insufficient to specify a transform $h$.
-One must also select a prior $p_Z$, and it is desirable to establish some general properties that a given pair $h$ and $p_Z$ should have.
-In addition to the above properties, the optimal pair would
-1. discard as few degrees of freedom $|m - n|$ as possible.
-2. produce a $p_Z$ and $|J_y|$ and their gradients that are efficient to compute.
-3. produce a $p_Y$ that is easy to sample from.
-
-We can expand a bit on the latter point.
-As a general rule, the more normal $p_Y$ is, the easier it is to sample from (REF?).
-Additionally, near singularities, small perturbations to $y$ tend to result in larger changes to $h(y)$ than far from the singularity.
-This warps the volume around the singularity, which causes $|J_y(y)|$ to smoothly approach infinity near the singularity.
-If $m$ is large, this is generally not a problem, as the "curse of dimensionality" works in our favor, and there is exponentially more volume away from the singularity than near the singularity.
-However, for low $m$, it is necessary that $p_{X,Z}$ goes to 0 near the singularity faster than $|J_y(y)|$ goes to infinity.
-While $p_X$ is generally fixed by the user-specified model and may not have this necessary property, the user may then choose a $p_Z$ that has this property.
-
 \section{Unit simplex}
@@
 for $p > 0$, which corresponds to $r \sim \text{Generalized-Gamma}(1, n, p)$.
-
-\subsubsection*{Acknowledgements}
+\subsection*{Acknowledgements}
 

### 2022-07-20 utils/sample.py

@@ -117,6 +117,7 @@ def sample(
                     show_progress=show_progress,
                     iter_sampling=n_iter,
                     chains=n_chains,
+                    seed=i
                 )
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 

### 2022-07-20 utils/sample.py
slurm scripts
maintenance (#32)

more maintenance

vectorize

misc

misc

-

 -

make paper for simplex and add randomseed

fix indent

indent ugh

syntax

haha syntax again rip

scripts for slurm

@@ -1,6 +1,7 @@
 import pandas as pd
 import numpy as np
 import os
+import time
 import json
 import sys
 import pickle
@@ -94,10 +95,15 @@ def sample(
                 f"target_densities/{evaluating_model}_parameters.json", "rb"
             ) as f:
                 parameters = pickle.load(f)
+       
         stan_filename=f'stan_models/{transform}_{evaluating_model}.stan'
-        
+        start_time = time.time()
         if type(parameters)==dict:
             parameters = [parameters]
+        
+        stan_filename=f'stan_models/{transform}_{evaluating_model}.stan'
+        start_time = time.time()
+
         for params in tqdm(parameters):
             model = CmdStanModel(
                 stan_file=stan_filename, cpp_options={"STAN_THREADS": "true"}
@@ -110,18 +116,25 @@ def sample(
                     chains=n_chains,
                 )
             )
-
-            for i in range(n_repeat - 1):
+	    
+            for i in tqdm(range(n_repeat - 1)):
                 fit = model.sample(
                     data=params,
                     show_progress=show_progress,
                     iter_sampling=n_iter,
                     chains=n_chains,
+                    seed=i
                 )
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 
             filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
             idata.to_netcdf(filename)
+<<<<<<< HEAD
+	    with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
+=======
+        with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
+>>>>>>> b0b57e1 (fix indent)
+	        f.write(str(time.time() - start_time))
         if return_idata==True:
             return idata
 

### 2022-07-20 utils/sample.py
slurm scripts
maintenance (#32)

more maintenance

vectorize

misc

misc

-

 -

make paper for simplex and add randomseed

fix indent

indent ugh

syntax

haha syntax again rip

scripts for slurm

@@ -137,6 +137,6 @@ def sample(
 	            f.write(str(time.time() - start_time))
             if return_idata==True:
                 return idata
-
-        pass
+        with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
+            f.write(str(time.time() - start_time))
 

### 2022-07-20 plot.py
slurm scripts
maintenance (#32)

more maintenance

vectorize

misc

misc

-

 -

make paper for simplex and add randomseed

fix indent

indent ugh

syntax

haha syntax again rip

scripts for slurm

@@ -22,7 +22,7 @@ parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*100, 'N': 100}, {'al
 {'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000},
 {'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000}]
 
-transforms = ['stan', 'softmax-augmented','stickbreaking', 'softmax']
+transforms = ['softmax', 'softmax-augmented','stickbreaking', 'stan']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
                              parameters, fig_name='rmse_example.png')

### 2022-07-20 simplex.py

@@ -6,13 +6,22 @@ import sys
 sys.path.insert(1, 'utils')
 
 from sample import sample
+import pickle
+import argparse
+parser = argparse.ArgumentParser()
+parser.add_argument('--parameters', type=int, required=True)
+parser.add_argument('--transform', type=str, required=True)
+args = parser.parse_args()
 
-transforms=['stan', 'stickbreaking', 'softmax', 'softmax-augmented']
+with open(f"target_densities/param_map_dirichlet_symmetric.json", "rb") as f:
+    param_map = pickle.load(f)
 
-for transform in tqdm(transforms):
-        sample(transform_category='simplex', transform=transform, 
-            evaluating_model='dirichlet_symmetric', parameters={}, 
-            auto_eval_all_params=True, n_iter = 1000, n_chains = 4, n_repeat = 1, 
+alpha = param_map[args.parameters]
+N = len(alpha)
+
+sample(transform_category='simplex', transform=args.transform, evaluating_model='dirichlet_symmetric', 
+parameters={'alpha':alpha, 'N':N}, 
+            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
                             show_progress = True, resample=True, return_idata=False)
 
 

### 2022-07-21 plot.py
syncing stuff

@@ -1,4 +1,5 @@
 
+
 import os
 os.chdir('transforms')
 
@@ -6,6 +7,17 @@ import sys
 sys.path.insert(1, 'utils')
 from sample import sample
 
+import pickle
+import argparse
+parser = argparse.ArgumentParser()
+parser.add_argument('--parameters', type=int, required=True)
+args = parser.parse_args()
+
+with open(f"target_densities/param_map_dirichlet_symmetric.json", "rb") as f:
+    param_map = pickle.load(f)
+
+alpha = param_map[args.parameters]
+N = len(alpha)
 
 import numpy as np
 import matplotlib.pyplot as plt
@@ -18,11 +30,8 @@ from ess import get_ess_leapfrog_ratio
 transform_category='simplex'
 evaluating_model='dirichlet_symmetric'
 
-parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*100, 'N': 100}, {'alpha': [0.1]*1000, 'N': 1000},
-{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000},
-{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000}]
-
 transforms = ['softmax', 'softmax-augmented','stickbreaking', 'stan']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
-                             parameters, fig_name='rmse_example.png')
+                             parameters=[{'alpha':alpha, 'N':N}], fig_name='rmse_1.png')
+

### 2022-07-21 utils/sample.py

@@ -129,11 +129,7 @@ def sample(
 
             filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
             idata.to_netcdf(filename)
-<<<<<<< HEAD
 	    with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
-=======
-        with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
->>>>>>> b0b57e1 (fix indent)
 	        f.write(str(time.time() - start_time))
         if return_idata==True:
             return idata

### 2022-07-21 utils/sample.py
whitespace adjustment

@@ -129,7 +129,7 @@ def sample(
 
             filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
             idata.to_netcdf(filename)
-	    with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
+        with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
 	        f.write(str(time.time() - start_time))
         if return_idata==True:
             return idata

### 2022-07-21 utils/dirichlet_symmetric.py
plot file fix

@@ -13,7 +13,7 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
     fig.supxlabel('Cumulative Leapfrog Steps')
     fig.supylabel('Root Mean Squared Error')
 
-    for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
+    for ax, params in zip([axes],  parameters):
         for transform in transforms:            
             idata = sample(transform_category=transform_category, transform=transform, 
                 evaluating_model='dirichlet_symmetric', parameters=[params], 

### 2022-07-21 utils/dirichlet_symmetric.py
stuff

@@ -13,7 +13,7 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
     fig.supxlabel('Cumulative Leapfrog Steps')
     fig.supylabel('Root Mean Squared Error')
 
-    for ax, params in zip([axes],  parameters):
+    for ax, params in zip(axes.T.flatten(),  parameters):
         for transform in transforms:            
             idata = sample(transform_category=transform_category, transform=transform, 
                 evaluating_model='dirichlet_symmetric', parameters=[params], 

### 2022-07-21 plot.py
indent shit

@@ -33,5 +33,5 @@ evaluating_model='dirichlet_symmetric'
 transforms = ['softmax', 'softmax-augmented','stickbreaking', 'stan']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
-                             parameters=[{'alpha':alpha, 'N':N}], fig_name='rmse_1.png')
+                             parameters=[{'alpha':alpha, 'N':N}],resample=False, fig_name='rmse_1.png', n_repeat=100)
 

### 2022-07-21 utils/dirichlet_symmetric.py

@@ -9,12 +9,11 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
     
     plt.rcParams["figure.figsize"] = [20,10]
     plt.rcParams['figure.dpi'] = 300
-    fig, axes = plt.subplots(3,3)
+    fig, ax = plt.subplots(1)
     fig.supxlabel('Cumulative Leapfrog Steps')
     fig.supylabel('Root Mean Squared Error')
 
-    for ax, params in zip(axes.T.flatten(),  parameters):
-        for transform in transforms:            
+    for transform in transforms:            
             idata = sample(transform_category=transform_category, transform=transform, 
                 evaluating_model='dirichlet_symmetric', parameters=[params], 
                 auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
@@ -22,7 +21,7 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
             alpha = params['alpha']
             N = params['N']
             true_x = [a/sum(alpha) for a in alpha]
-            x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
+    x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
             ax.plot(x,y, label = str(transform))
 
         ax.set_title(f'alpha={alpha[0]}, N = {N}')

### 2022-07-21 utils/sample.py

@@ -86,7 +86,7 @@ def sample(
 
     if resample == False:
         for params in parameters:
-            filename = f'sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
             return az.load_arviz_data(filename)
     else:
 

### 2022-07-21 utils/dirichlet_symmetric.py
stuff

@@ -14,15 +14,15 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
     fig.supylabel('Root Mean Squared Error')
 
     for transform in transforms:            
-            idata = sample(transform_category=transform_category, transform=transform, 
-                evaluating_model='dirichlet_symmetric', parameters=[params], 
-                auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
-                                        show_progress = show_progress, resample=resample)
-            alpha = params['alpha']
-            N = params['N']
-            true_x = [a/sum(alpha) for a in alpha]
-    x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
-            ax.plot(x,y, label = str(transform))
+        idata = sample(transform_category=transform_category, transform=transform, 
+            evaluating_model='dirichlet_symmetric', parameters=parameters, 
+            auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
+                                    show_progress = show_progress, resample=resample)
+        alpha = parameters['alpha']
+        N = parameters['N']
+        true_x = [a/sum(alpha) for a in alpha]
+        x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
+        ax.plot(x,y, label = str(transform))
 
         ax.set_title(f'alpha={alpha[0]}, N = {N}')
     fig.legend(labels=transforms,bbox_to_anchor = (0.6, -0.05));

### 2022-07-21 plot.py
indent shit

@@ -30,7 +30,7 @@ from ess import get_ess_leapfrog_ratio
 transform_category='simplex'
 evaluating_model='dirichlet_symmetric'
 
-transforms = ['softmax', 'softmax-augmented','stickbreaking', 'stan']
+transforms = ['stan', 'softmax-augmented','stickbreaking', 'softmax']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
                              parameters=[{'alpha':alpha, 'N':N}],resample=False, fig_name='rmse_1.png', n_repeat=100)

### 2022-07-21 utils/dirichlet_symmetric.py

@@ -18,6 +18,7 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
             evaluating_model='dirichlet_symmetric', parameters=parameters, 
             auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
                                     show_progress = show_progress, resample=resample)
+	print(parameters)
         alpha = parameters['alpha']
         N = parameters['N']
         true_x = [a/sum(alpha) for a in alpha]

### 2022-07-21 utils/sample.py

@@ -87,7 +87,7 @@ def sample(
     if resample == False:
         for params in parameters:
             filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
-            return az.load_arviz_data(filename)
+            return az.from_netcdf(filename)
     else:
 
         if auto_eval_all_params:

### 2022-07-21 utils/dirichlet_symmetric.py
indent

@@ -18,7 +18,7 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
             evaluating_model='dirichlet_symmetric', parameters=parameters, 
             auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
                                     show_progress = show_progress, resample=resample)
-	print(parameters)
+        print(parameters)
         alpha = parameters['alpha']
         N = parameters['N']
         true_x = [a/sum(alpha) for a in alpha]

### 2022-07-21 utils/dirichlet_symmetric.py
add figure

@@ -19,8 +19,8 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
             auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
                                     show_progress = show_progress, resample=resample)
         print(parameters)
-        alpha = parameters['alpha']
-        N = parameters['N']
+        alpha = parameters[0]['alpha']
+        N = parameters[0]['N']
         true_x = [a/sum(alpha) for a in alpha]
         x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
         ax.plot(x,y, label = str(transform))

### 2022-07-21 simplex.py
i hate indents

@@ -21,7 +21,7 @@ N = len(alpha)
 
 sample(transform_category='simplex', transform=args.transform, evaluating_model='dirichlet_symmetric', 
 parameters={'alpha':alpha, 'N':N}, 
-            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
+            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 2, 
                             show_progress = True, resample=True, return_idata=False)
 
 

### 2022-07-21 utils/sample.py

@@ -128,6 +128,7 @@ def sample(
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 
             filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+		print(filename)
             idata.to_netcdf(filename)
         with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
 	        f.write(str(time.time() - start_time))

### 2022-07-21 utils/sample.py
whitespace

@@ -128,8 +128,8 @@ def sample(
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 
             filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
-		print(filename)
-            idata.to_netcdf(filename)
+        print(filename)
+        idata.to_netcdf(filename)
         with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
 	        f.write(str(time.time() - start_time))
         if return_idata==True:

### 2022-07-21 utils/dirichlet_symmetric.py
-

@@ -33,13 +33,13 @@ def create_param_map():
     alphas = [0.1, 1, 10]
     Ns = [10, 100, 1000]
     keys = [1, 2, 3, 4, 5, 6, 7, 8, 9]
-    params = []
+    parameters = []
     for alpha in alphas:
         for N in Ns:
-            parameters.append({"alpha": [alpha] * N, "N": N})
+            parameters.append({"alpha": [alpha]*N, "N": N})
 
-    param_map = dict(zip(keys, params))
-    param_map.update(dict(zip(params, keys)))
+    param_map = dict(zip(keys, parameters))
+    param_map.update(dict(zip(parameters, keys)))
 
     pickle.dump(param_map, open("param_map_dirichlet_symmmetric.pkl", "wb"))
 

### 2022-07-21 utils/dirichlet_symmetric.py
more changes

@@ -37,7 +37,7 @@ def create_param_map():
     for alpha in alphas:
         for N in Ns:
             parameters.append({"alpha": [alpha]*N, "N": N})
-
+    
     param_map = dict(zip(keys, parameters))
     param_map.update(dict(zip(parameters, keys)))
 

### 2022-07-21 utils/sample.py

@@ -81,7 +81,7 @@ def sample(
     The directory structure for storing stan files is: stan_models/{transform}_{evaluating_model}.stan
     The directory structure for storing the results is: sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_mapping}.json
     """
-    with open(f"target_densities/param_map_{evaluating_model}.json", "rb") as f:
+    with open(f"target_densities/param_map_{evaluating_model}.pkl", "rb") as f:
         param_map = pickle.load(f)
 
     if resample == False:

### 2022-07-21 plot.py
fixmylife

@@ -13,7 +13,7 @@ parser = argparse.ArgumentParser()
 parser.add_argument('--parameters', type=int, required=True)
 args = parser.parse_args()
 
-with open(f"target_densities/param_map_dirichlet_symmetric.json", "rb") as f:
+with open(f"target_densities/param_map_dirichlet_symmetric.pkl", "rb") as f:
     param_map = pickle.load(f)
 
 alpha = param_map[args.parameters]

### 2022-07-21 simplex.py

@@ -13,7 +13,7 @@ parser.add_argument('--parameters', type=int, required=True)
 parser.add_argument('--transform', type=str, required=True)
 args = parser.parse_args()
 
-with open(f"target_densities/param_map_dirichlet_symmetric.json", "rb") as f:
+with open(f"target_densities/param_map_dirichlet_symmetric.pkl", "rb") as f:
     param_map = pickle.load(f)
 
 alpha = param_map[args.parameters]

### 2022-07-21 simplex.py
minor things

@@ -23,5 +23,4 @@ sample(transform_category='simplex', transform=args.transform, evaluating_model=
 parameters={'alpha':alpha, 'N':N}, 
             auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 2, 
                             show_progress = True, resample=True, return_idata=False)
-
-
+                            

### 2022-07-21 utils/sample.py

@@ -21,7 +21,7 @@ def sample(
     transform,
     evaluating_model,
     parameters,
-    output_dir='/mnt/sdceph/users/mjhajaria',
+    output_dir='/mnt/sdceph/users/mjhajaria/',
     auto_eval_all_params=False,
     n_iter=1000,
     n_chains=4,
@@ -127,10 +127,10 @@ def sample(
                 )
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 
-            filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            filename = f'{output_dir}sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
         print(filename)
         idata.to_netcdf(filename)
-        with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
+        with open(f'{output_dir}sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
 	        f.write(str(time.time() - start_time))
         if return_idata==True:
             return idata

### 2022-07-21 utils/dirichlet_symmetric.py
fix bugs again

@@ -37,7 +37,7 @@ def create_param_map():
     for alpha in alphas:
         for N in Ns:
             parameters.append({"alpha": [alpha]*N, "N": N})
-    
+
     param_map = dict(zip(keys, parameters))
     param_map.update(dict(zip(parameters, keys)))
 
@@ -46,3 +46,5 @@ def create_param_map():
 
 def get_dirichlet_symmetric_params():
     return pickle.load(open("param_map_dirichlet_symmmetric.pkl", "rb"))
+
+

### 2022-07-22 plot.py
add plots

@@ -33,5 +33,5 @@ evaluating_model='dirichlet_symmetric'
 transforms = ['stan', 'softmax-augmented','stickbreaking', 'softmax']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
-                             parameters=[{'alpha':alpha, 'N':N}],resample=False, fig_name='rmse_1.png', n_repeat=100)
+                             parameters=[{'alpha':alpha, 'N':N}],resample=False, fig_name=f'rmse_{args.parameters}.png', n_repeat=100)
 

### 2022-07-22 simplex.py

@@ -21,6 +21,6 @@ N = len(alpha)
 
 sample(transform_category='simplex', transform=args.transform, evaluating_model='dirichlet_symmetric', 
 parameters={'alpha':alpha, 'N':N}, 
-            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 2, 
+            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
                             show_progress = True, resample=True, return_idata=False)
                             

### 2022-07-22 utils/dirichlet_symmetric.py
slight changes

@@ -24,7 +24,7 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
         true_x = [a/sum(alpha) for a in alpha]
         x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
         ax.plot(x,y, label = str(transform))
-
+        ax.legend()
         ax.set_title(f'alpha={alpha[0]}, N = {N}')
     fig.legend(labels=transforms,bbox_to_anchor = (0.6, -0.05));
     fig.savefig(f'figures/{fig_name}', dpi=300)

### 2022-07-22 Untitled.py
add time

@@ -1,0 +1,60 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+import sys
+sys.path.insert(1, 'utils')
+
+import numpy as np
+import matplotlib.pyplot as plt
+import arviz as az
+import time
+import pickle 
+
+from sample import sample
+from dirichlet_symmetric import get_dirichlet_symmetric_rmse
+from rhat import calc_rhat_mixed_chains
+from ess import get_ess_leapfrog_ratio
+
+def get_param_map(param):
+    with open(f"target_densities/param_map_dirichlet_symmetric.pkl", "rb") as f:
+        param_map = pickle.load(f)
+    alpha = param_map[param]
+    N = len(alpha)
+    return alpha[0], N
+
+def timer(transform, param):
+    with open(f"target_densities/param_map_dirichlet_symmetric.pkl", "rb") as f:
+        param_map = pickle.load(f)
+    alpha = param_map[param]
+    N = len(alpha)
+    start_time = time.time()
+    
+    sample(transform_category='simplex', transform=transform, evaluating_model='dirichlet_symmetric', 
+    parameters={'alpha':alpha, 'N':N}, 
+                auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 10, 
+                                show_progress = True, resample=True, return_idata=False, output_dir='')
+
+    print(time.time() - start_time)
+    return time.time() - start_time
+
+
+
+
+transforms = ['stickbreaking']
+timedict={}
+
+for param in [1,2,3,4,5,6,7,8,9]:
+    for transform in transforms:
+        timetaken = timer(transform,param)
+        print(str(transform), timetaken)
+        timedict[f'{transform}_{param}'] = timetaken
+
+
+
+
+# import json
+    
+# with open("time.json", "w") as f:
+#     json.dump(timedict, f)
+
+

### 2022-07-22 Untitled.py
replot

@@ -1,6 +1,6 @@
-#!/usr/bin/env python
-# coding: utf-8
 
+import os
+os.chdir('transforms')
 import sys
 sys.path.insert(1, 'utils')
 
@@ -32,7 +32,7 @@ def timer(transform, param):
     sample(transform_category='simplex', transform=transform, evaluating_model='dirichlet_symmetric', 
     parameters={'alpha':alpha, 'N':N}, 
                 auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 10, 
-                                show_progress = True, resample=True, return_idata=False, output_dir='')
+                                show_progress = True, resample=True, return_idata=False)
 
     print(time.time() - start_time)
     return time.time() - start_time
@@ -52,9 +52,9 @@ for param in [1,2,3,4,5,6,7,8,9]:
 
 
 
-# import json
+import json
     
-# with open("time.json", "w") as f:
-#     json.dump(timedict, f)
+with open("time_sb.json", "w") as f:
+    json.dump(timedict, f)
 
 

### 2022-07-22 utils/dirichlet_symmetric.py

@@ -18,7 +18,6 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
             evaluating_model='dirichlet_symmetric', parameters=parameters, 
             auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
                                     show_progress = show_progress, resample=resample)
-        print(parameters)
         alpha = parameters[0]['alpha']
         N = parameters[0]['N']
         true_x = [a/sum(alpha) for a in alpha]

### 2022-07-22 plot.py
plotmylife

@@ -33,5 +33,5 @@ evaluating_model='dirichlet_symmetric'
 transforms = ['stan', 'softmax-augmented','stickbreaking', 'softmax']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
-                             parameters=[{'alpha':alpha, 'N':N}],resample=False, fig_name=f'rmse_{args.parameters}.png', n_repeat=100)
+                             parameters=[{'alpha':alpha, 'N':N}],resample=True, fig_name=f'rmse_{args.parameters}.png', n_repeat=1)
 

### 2022-07-22 utils/sample.py

@@ -86,7 +86,7 @@ def sample(
 
     if resample == False:
         for params in parameters:
-            filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            filename = f'{output_dir}sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
             return az.from_netcdf(filename)
     else:
 

### 2022-07-22 plot.py
Revert "plotmylife"
This reverts commit bde7b392ee8ddcc5837802f7b43cb8580b74b646.

@@ -33,5 +33,5 @@ evaluating_model='dirichlet_symmetric'
 transforms = ['stan', 'softmax-augmented','stickbreaking', 'softmax']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
-                             parameters=[{'alpha':alpha, 'N':N}],resample=True, fig_name=f'rmse_{args.parameters}.png', n_repeat=1)
+                             parameters=[{'alpha':alpha, 'N':N}],resample=False, fig_name=f'rmse_{args.parameters}.png', n_repeat=100)
 

### 2022-07-22 utils/sample.py

@@ -86,7 +86,7 @@ def sample(
 
     if resample == False:
         for params in parameters:
-            filename = f'{output_dir}sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
             return az.from_netcdf(filename)
     else:
 

### 2022-07-22 plot.py
plots (#40)
* syncing stuff

* whitespace adjustment

* plot file fix

* stuff

* indent shit

* stuff

* indent shit

* indent

* add figure

* i hate indents

* whitespace

* -

* more changes

* fixmylife

* minor things

* fix bugs again

* add plots

* slight changes

* --

* add time

* replot

* plotmylife

* Revert "plotmylife"

This reverts commit bde7b392ee8ddcc5837802f7b43cb8580b74b646.

@@ -1,4 +1,5 @@
 
+
 import os
 os.chdir('transforms')
 
@@ -6,6 +7,17 @@ import sys
 sys.path.insert(1, 'utils')
 from sample import sample
 
+import pickle
+import argparse
+parser = argparse.ArgumentParser()
+parser.add_argument('--parameters', type=int, required=True)
+args = parser.parse_args()
+
+with open(f"target_densities/param_map_dirichlet_symmetric.pkl", "rb") as f:
+    param_map = pickle.load(f)
+
+alpha = param_map[args.parameters]
+N = len(alpha)
 
 import numpy as np
 import matplotlib.pyplot as plt
@@ -18,11 +30,8 @@ from ess import get_ess_leapfrog_ratio
 transform_category='simplex'
 evaluating_model='dirichlet_symmetric'
 
-parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*100, 'N': 100}, {'alpha': [0.1]*1000, 'N': 1000},
-{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000},
-{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000}]
-
-transforms = ['softmax', 'softmax-augmented','stickbreaking', 'stan']
+transforms = ['stan', 'softmax-augmented','stickbreaking', 'softmax']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
-                             parameters, fig_name='rmse_example.png')
+                             parameters=[{'alpha':alpha, 'N':N}],resample=False, fig_name=f'rmse_{args.parameters}.png', n_repeat=100)
+

### 2022-07-22 simplex.py

@@ -13,7 +13,7 @@ parser.add_argument('--parameters', type=int, required=True)
 parser.add_argument('--transform', type=str, required=True)
 args = parser.parse_args()
 
-with open(f"target_densities/param_map_dirichlet_symmetric.json", "rb") as f:
+with open(f"target_densities/param_map_dirichlet_symmetric.pkl", "rb") as f:
     param_map = pickle.load(f)
 
 alpha = param_map[args.parameters]
@@ -23,5 +23,4 @@ sample(transform_category='simplex', transform=args.transform, evaluating_model=
 parameters={'alpha':alpha, 'N':N}, 
             auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
                             show_progress = True, resample=True, return_idata=False)
-
-
+                            

### 2022-07-22 utils/dirichlet_symmetric.py

@@ -9,22 +9,21 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
     
     plt.rcParams["figure.figsize"] = [20,10]
     plt.rcParams['figure.dpi'] = 300
-    fig, axes = plt.subplots(3,3)
+    fig, ax = plt.subplots(1)
     fig.supxlabel('Cumulative Leapfrog Steps')
     fig.supylabel('Root Mean Squared Error')
 
-    for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
-        for transform in transforms:            
-            idata = sample(transform_category=transform_category, transform=transform, 
-                evaluating_model='dirichlet_symmetric', parameters=[params], 
-                auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
-                                        show_progress = show_progress, resample=resample)
-            alpha = params['alpha']
-            N = params['N']
-            true_x = [a/sum(alpha) for a in alpha]
-            x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
-            ax.plot(x,y, label = str(transform))
-
+    for transform in transforms:            
+        idata = sample(transform_category=transform_category, transform=transform, 
+            evaluating_model='dirichlet_symmetric', parameters=parameters, 
+            auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
+                                    show_progress = show_progress, resample=resample)
+        alpha = parameters[0]['alpha']
+        N = parameters[0]['N']
+        true_x = [a/sum(alpha) for a in alpha]
+        x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
+        ax.plot(x,y, label = str(transform))
+        ax.legend()
         ax.set_title(f'alpha={alpha[0]}, N = {N}')
     fig.legend(labels=transforms,bbox_to_anchor = (0.6, -0.05));
     fig.savefig(f'figures/{fig_name}', dpi=300)
@@ -33,16 +32,18 @@ def create_param_map():
     alphas = [0.1, 1, 10]
     Ns = [10, 100, 1000]
     keys = [1, 2, 3, 4, 5, 6, 7, 8, 9]
-    params = []
+    parameters = []
     for alpha in alphas:
         for N in Ns:
[…]

### 2022-07-22 utils/sample.py

@@ -21,7 +21,7 @@ def sample(
     transform,
     evaluating_model,
     parameters,
-    output_dir='/mnt/sdceph/users/mjhajaria',
+    output_dir='/mnt/sdceph/users/mjhajaria/',
     auto_eval_all_params=False,
     n_iter=1000,
     n_chains=4,
@@ -81,13 +81,13 @@ def sample(
     The directory structure for storing stan files is: stan_models/{transform}_{evaluating_model}.stan
     The directory structure for storing the results is: sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_mapping}.json
     """
-    with open(f"target_densities/param_map_{evaluating_model}.json", "rb") as f:
+    with open(f"target_densities/param_map_{evaluating_model}.pkl", "rb") as f:
         param_map = pickle.load(f)
 
     if resample == False:
         for params in parameters:
-            filename = f'sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
-            return az.load_arviz_data(filename)
+            filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+            return az.from_netcdf(filename)
     else:
 
         if auto_eval_all_params:
@@ -127,13 +127,10 @@ def sample(
                 )
                 idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
 
-            filename = f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
-            idata.to_netcdf(filename)
-<<<<<<< HEAD
-	    with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
-=======
-        with open(f'{output_dir}/sampling_results/{transform_category}/{transform}/{evaluating_model}/time_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.txt', 'w') as f:
->>>>>>> b0b57e1 (fix indent)
+            filename = f'{output_dir}sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_map[tuple(list(params.values())[0])]}_{n_repeat}.nc'
+        print(filename)
[…]

### 2022-07-22 tex/paper.tex
plots (#40)
* syncing stuff

* whitespace adjustment

* plot file fix

* stuff

* indent shit

* stuff

* indent shit

* indent

* add figure

* i hate indents

* whitespace

* -

* more changes

* fixmylife

* minor things

* fix bugs again

* add plots

* slight changes

* --

* add time

* replot

* plotmylife

* Revert "plotmylife"

This reverts commit bde7b392ee8ddcc5837802f7b43cb8580b74b646.

plots and other fixes

* syncing stuff

* whitespace adjustment

* plot file fix

* stuff

* indent shit

* stuff

* indent shit

* indent

* add figure

* i hate indents

* whitespace

* -

* more

--- 2022-07-18 tex/paper.tex
@@
   Parameter Transforms for Hamiltonian Monte Carlo}
-\author{Meenal Jhajaria \\ \small Flatiron Institute \and Seth Axen \\
-  \small University of T\"ubingen \and Bob
-  Carpenter \\ \small Flatiron Institute}
+\author{Meenal Jhajaria \\ \small Flatiron Institute 
+\and Seth Axen \\ \small University of T\"ubingen 
+\and Adam Haber \\ \small Weizmann Institute 
+\and Sean Pinkney \\ \small Omnicom Media Group
+\and Bob Carpenter \\ \small Flatiron Institute}
 \date{DRAFT: \today}
@@
   \noindent
-  This paper evaluates the the statistical and computational
+  This paper evaluates the statistical and computational
   efficiency of unconstraining parameter transforms for Hamiltonian
@@
 integrals over densities $\pi(x)$ (e.g., Bayesian estimation or
-prediction, $p$-value calcuations, etc.).  The only black-box
+prediction, $p$-value calculations, etc.).  The only black-box
 techniques that work for general high-dimensional integrals are
@@
 
-\section{Changes of Variables}
+\section{Transforms}
+
+To draw samples from a distribution $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.
+
+Let $f\colon \mathcal{Y} \to \mathcal{X}$ be a smooth and continuous map.
+Only when $m = n$ can $f$ be bijective.
+When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = f(y)$.
+In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
+We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
+$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
+that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
+When $h$ is non-bijective at single points, we call these singularities.
+While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.
+
+\subsection{Changes of Variables}
 
 If $X$ is a random variable with density $p_X$ and $Y = f(X)$ for a
-smooth and bijective function $f$, then
+smooth, monotonic, bijection $f$, then
 \[
@@
   \item Rectangular matrix $\{X \in \mathbb{R}^{N \times K}\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^N \dv{X_{ij}}$
-  \item Upper-triangular rectangular matrix $\{X \in \mathbb{R}^{N \times K} | X_{ij} = 0\ \forall\ i > j\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{\min(j, N)} \dv{X_{ij}}$
+  \item Lower-triangular rectangular matrix $\{X \in \mathbb{R}^{N \times K} | X_{ij} = 0\ \forall\ i < j\}$: $(\dv{X})^\wedge = \bigwedge_{i=1}^N \bigwedge_{j=1}^{\min(i, K)} \dv{X_{ij}}$
   \item Diagonal matrix $\{X \in \mathbb{R}^{N \times K} | X_{ij} = 0\ \forall\ i \ne j\}$: $(\dv{X})^\wedge = \bigwedge_{i=1}^{\min(N, K)} \dv{X_{ii}}$
-  \item Symmetric matrix $\{X \in \mathbb{R}^{N \times N} | X_{ij} = X_{ji}\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{j} \dv{X_{ij}}$
-  \item Skew-symmetric matrix $\{X \in \mathbb{R}^{N \times N} | X_{ij} = -X_{ji}, X_{ii} = 0\}$: $(\dv{X})^\wedge = \bigwedge_{j=1}^K \bigwedge_{i=1}^{j-1} \dv{X_{ij}}$
+  \item Symmetric matrix $\{X \in \mathbb{R}^{N \times N} | X_{ij} = X_{ji}\}$: $(\dv{X})^\wedge = \bigwedge_{i=1}^N \bigwedge_{j=1}^i \dv{X_{ij}}$
+  \item Skew-symmetric matrix $\{X \in \mathbb{R}^{N \times N} | X_{ij} = -X_{ji}, X_{ii} = 0\}$: $(\dv{X})^\wedge = \bigwedge_{i=1}^N \bigwedge_{j=1}^{i-1} \dv{X_{ij}}$
 \end{itemize}
@@
 
-\subsection{Simplex augmented softmax parameterization}
+\subsection{Simplex softmax parameterization through parameter expansion}
 
@@
 
-\subsection{Semi-orthogonal matrices (Stiefel manifold)}
+\section{Semi-orthogonal matrices (Stiefel manifold)}
 
@@
 The semi-orthogonal matrices form a space called the Stiefel manifold.
-When $K=N$, $X \in O(N)$ is an orthogonal matrix and also satisfies $X X^\top = I_N$ and $\det(X) = \pm 1$.
 When $K=1$, $X$ is a point on the $(N-1)$-sphere $\mathbb{S}^{N-1} \subset \mathbb{R}^N$.
+We will consider such unit vectors in the next section.
+When $K=N$, $X$ is an element of the orthogonal group $O(N)$ and additionally satisfies $\det(X) = \pm 1$.
+$O(N)$ is comprised of two disconnected submanifolds, one being the Special Orthogonal group $SO(N)$ with $\det(X) = +1$, and the other having $\det(X) = -1$.
+$SO(2)$ and $SO(3)$ are the groups of rotation matrices of 2D and 3D vectors, respectively.
 
@@
 
-\subsubsection{Givens parameterization of Stiefel manifold}
+\subsection{Givens parameterization of Stiefel manifold}
 
-\subsubsection{Cayley parameterization of Stiefel manifold}
+\subsection{Cayley parameterization of Stiefel manifold}
 
-\subsubsection{Householder parameterization of Stiefel manifold}
+\subsection{Polar parameterization of Stiefel manifold}
 
-\subsubsection{Matrix exponential parameterization of Orthogonal group}
+\subsection{Householder parameterization of Stiefel manifold}
 
-\subsection{Positive definite matrices}
+\subsection{Matrix exponential parameterization of Special Orthogonal group}
+
+\section{Unit sphere}
+
+The unit sphere $\mathbb{S}^{N-1} \subset \mathbb{R}^N$ consists of all unit vectors $x$ of length 1, i.e. $x^\top x = 1$.
+It is a special case of the Stiefel manifold.
+
+Densities on the sphere are generally defined with respect to the spherically uniform measure.
+This measure can be defined in many ways that are equivalent to within a constant factor that is not relevant for MCMC.
+We will take the same approach as with the Stiefel manifold and define the spherically uniform measure as the measure that is invariant to transformations by $N \times N$ orthogonal matrices (rotations and reflections).
+As discussed in the previous section, if $\tilde{X} \in O(N)$ is an orthogonal matrix whose first column is $x$, then $\tilde{X}^\top x = e_1$, where $e_1$ is the unit vector whose first element is 1, and $(\tilde{X}^\top \dv{x})_1 = 0$.
+For convenience, we will assume that the rules used to produce $\tilde{X}$ enforce that $\det(\tilde{X}) = +1$.
+The spherically uniform measure is then written
+\[
+    (\dv{x})^\wedge \equiv (\tilde{X}^\top \dv{x})^\wedge = \bigwedge_{i=2}^N (\tilde{X}^\top \dv{x})_i.
+\]
+
+\subsection{Embedding in real space using parameter expansion}
+
+Let $f\colon y \mapsto (r, x)$ for $y \in \mathbb{R}^N$, $r > 0$, and $x \in \mathbb{S}^{N-1}$.
+This map is bijective everywhere except $y=0$.
+
+To compute the Jacobian determinant of the map, we first compute the differential form of $y=rx$:
+
+\[\dv{y} = \dv{r} x + r \dv{x}.\]
+
+Multiplying both sides by $\tilde{X}^\top$, we find
+
+\[
+    \tilde{X}^\top \dv{y} = \dv{r} \tilde{X}^\top x + r \tilde{X}^\top \dv{x} = \dv{r} e_1 + r (\tilde{X}^\top \dv{x}).
+    \]
+
+Then
+
+\[
+    (\tilde{X}^\top \dv{y})_i = \begin{cases}
+        \dv{r} & \text{ if } i = 1 \\
+        r (\tilde{X}^\top \dv{x})_i & \text{ if } i \ne 1
+    \end{cases}.
+\]
+
+Note that $(\tilde{X}^\top \dv{y})^\wedge = \det({\tilde{X}}) (\dv{y})^\wedge = (\dv{y})^\wedge$.
+
+Taking the exterior product of both sides, we then find
+
+\[
+\begin{aligned}
+    (\dv{y})^\wedge &= \dv{r} \wedge \left(\bigwedge_{i=2}^N r (\tilde{X}^\top \dv{x})_i \right) = r^{N-1} \left(\dv{r} \wedge (\dv{x})^\wedge\right)\\
+    \dv{r} \wedge (\dv{x})^\wedge &= r^{1-N} (\dv{y})^\wedge = \lVert y \rVert^{1-N} (\dv{y})^\wedge
+\end{aligned}
+\]
+
+So the absolute Jacobian determinant is $\absdet{J_f(y)} = \lVert y \rVert^{1-N}$.
+
+In the constrained space there is a free radial parameter $r$ that is unconstrained on the positive reals.
+As a result, nothing prevents $r$ from growing arbitrarily large, which would cause the posterior to be improper, i.e. to have infinite mass.
+Thus, we must also specify a prior $\pi(r)$.
+While we may want to customize $\pi(r)$, a reasonable default choice is $r \sim \chi_N$, which has the density $\pi(r) \propto r^{N-1} \exp\left(-\frac{1}{2}r^2\right) $.
+
+The combined log density correction is then
+\[
+  \absdet{J_f(y)} \pi(r) = \absdet{J_f(y)} \pi(\lVert y \rVert) \propto \lVert y \rVert^{1-N} \lVert y \rVert^{N-1} \exp\left(-\frac{1}{2}\lVert y \rVert^2\right) = \exp\left(-\frac{1}{2}\lVert y \rVert^2\right).
+\]
+
+That is, if $x$ is uniformly distributed on the sphere and $r$ follows this prior, then $y$ has a zero-centered standard multivariate normal distribution.
+
+Since HMC is especially efficient at sampling normal distributions, this parameterization is very efficient when the marginal posterior distribution of $x$ is very diffuse on the sphere.
+However, as the target distribution becomes concentrated on $x$, the geometry near the singularity $y=0$ may become problematic.
+
+\subsection{Parameterization using hyperspherical coordinates}
+
+Given $z \in \mathbb{R}^N$, $r > 0$ and angles $\phi=\{\phi_1, \ldots, \phi_{N-1}\}$, where $\phi_{N-1} \in (0, 2\pi]$ and $\phi_i \in (0, \pi]$ for $i < N-1$, we define the relationship between $z$ and $(r, \phi)$ as
+
+\[
+  z_i = \left(r \prod_{j=1}^{i-1} \sin(\phi_j)\right) \begin{cases}
+    1 & \text{ if } i = N\\
+    \cos(\phi_i) & \text{otherwise}
+  \end{cases}.
+\]
+
+For the plane $N=2$ this works out the usual polar coordinate system:
+\[
+    \begin{aligned}
+        z_1 &= r \cos\phi_1\\
+        z_2 &= r \sin\phi_1
+    \end{aligned}
+\]
+
+For $N=3$ this is \footnote{
+    Note that this definition of hyperspherical coordinates differs for $N=3$ from the usual definition of spherical coordinates.
+    It can be made the same through the coordinate permutation $(z_1, z_2, z_3) \to (z_2, z_3, z_1)$.
+}
+\[
+    \begin{aligned}
+        z_1 &= r \cos\phi_1\\
+        z_2 &= r \sin\phi_1\cos\phi_2\\
+        z_3 &= r \sin\phi_1\sin\phi_2
+    \end{aligned}
+\]
+
+Our strategy is to compute the Jacobian determinant of the bijective map $g: (r, \phi) \mapsto z$ and relate it to the Jacobian determinant from the previous section to find the Jacobian determinant of the bijective map $f_2: \phi \mapsto x$ for $x \in \mathbb{S}^{N-1}$.
+
+To do so, note first that
+\[
+  z_i^2 = \left(r^2 \prod_{j=1}^{i-1} \sin^2(\phi_j)\right) \begin{cases}
+    1 & \text{ if } i = N\\
+    \cos^2(\phi_i) & \text{otherwise}
+  \end{cases}.
+\]
+
+Note also that
+\[
+    \begin{aligned}
+        z_N^2 &= r^2 \prod_{j=1}^{N-1} \sin^2(\phi_j)\\
+        z_{N-1}^2 + z_N^2 &= \left(r^2 \prod_{j=1}^{N-2} \sin^2(\phi_j)\right) \cos^2(\phi_{N-1}) + \left(r^2 \prod_{j=1}^{N-2} \sin^2(\phi_j) \right) \sin^2(\phi_{N-1}) = r^2 \prod_{j=1}^{N-2} \sin^2(\phi_j)\\
+        z_{N-2}^2 + z_{N-1}^2 + z_N^2 &= \left(r^2 \prod_{j=1}^{N-3} \sin^2(\phi_j)\right) \cos^2(\phi_{N-2}) + \left(r^2 \prod_{j=1}^{N-3} \sin^2(\phi_j)\right) \sin^2(\phi_{N-2}) = r^2 \prod_{j=1}^{N-3} \sin^2(\phi_j)\\
+        &\vdots \\
+        \sum_{k=1}^i z_{N - i + k}^2 &= r^2 \prod_{j=1}^{N - i} \sin^2(\phi_j)\\
+        &\vdots \\
+        \sum_{k=1}^N z_{N - i + k}^2 &= r^2
+    \end{aligned}
+\]
+
+We then differentiate both sides of this expression to get the differential form:
+
+\[
+    \begin{aligned}
+        2 z_N \dv{z_N} &= 2r\dv{r} + 2 r^2 \left(\prod_{j=1}^{N-1} \sin^2(\phi_j)\right)\left( \sum_{j=1}^{N-1} \cot(\phi_j) \dv{\phi_j} \right)\\
+        2z_{N-1}\dv{z_{N-1}} + 2z_{N}\dv{z_{N}} &= 2r \dv{r} + 2 r^2 \left(\prod_{j=1}^{N-2} \sin^2(\phi_j)\right) \left(\sum_{j=1}^{N-2} \cot(\phi_j)\right)\\
+        2z_{N-2}\dv{z_{N-2}} + 2z_{N-1}\dv{z_{N-1}} + 2z_{N}\dv{z_{N}} &= 2r \dv{r} + 2 r^2 \left(\prod_{j=1}^{N-3} \sin^2(\phi_j)\right) \left(\sum_{j=1}^{N-3} \cot(\phi_j)\right)\\
+        &\vdots \\
+        2\sum_{k=1}^i z_{N - i + k} \dv{z_{N - i + k}} &= 2 r \dv{r} + 2 r^2 \left(\prod_{j=1}^{N - i} \sin^2(\phi_j)\right) \left(\sum_{k=1}^{N-i} \cot(\phi_k) \dv{\phi_k}\right)\\
+        &\vdots \\
+        2\sum_{k=1}^N z_{N - i + k} \dv{z_{N - i + k}} &= 2r\dv{r}
+    \end{aligned}
+\]
+
+For each side separately, we now take the exterior product for all $i \in 1\ldots N$.
+On the left side, we start at the top with $(2z_N dz_n) \wedge (2 z_{N-1} \dv{z_{N-1}} + 2 z_N \dv{z_N}) = 2^2 z_{N} z_{N-1} (\dv{z_N} \wedge \dv{z_{N-1}})$.
+Note that the $\dv{z_N}$ from the second equation is eliminated since the exterior product already contains that term.
+This trend continues as we proceed to the bottom, and the result is
+\[
+    \bigwedge_{i=1}^N 2z_{N-i+1} \dv{z_{N-i+1}} = \bigwedge_{i=1}^N 2z_i \dv{z_i} = 2^N \left(\prod_{i=1}^N z_i \right) (\dv{z})^\wedge
+\]
+
+Note that
+
+\[
+  \prod_{i=1}^N z_i = r \prod_{i=1}^{N-1} r \sin(\phi_i) \cos(\phi_i) \prod_{j=1}^{i-1} \sin(\phi_j) = r^N \prod_{i=1}^{N-1} \sin(\phi_i) \cos(\phi_i) \prod_{j=1}^{i-1} \sin(\phi_j)
+\]
+
+On the right side, we start at the bottom instead with
+\[
+(2r\dv{r} + 2 r^2 \sin(\phi_1) \cos(\phi_1) \dv{\phi_1}) \wedge (2r\dv{r}) = 4 r^3 \sin(\phi_1) \cos(\phi_1) (\dv{\phi_1} \wedge \dv{r}).
+\]
+Again, the $\dv{r}$ term in the second-from-bottom equation cancels since $\dv{r}$ is already included in the bottom term.
+This trend also continues as we proceed to the top, and the result is then
+
+\[
+\begin{aligned}
+    \left(\bigwedge_{i=1}^{N - 1} 2 r^2 \left(\prod_{j=1}^{N - i - 1} \sin^2(\phi_j)\right) \sin(\phi_{N-i}) \cos(\phi_{N-i}) \dv{\phi_{N-i}} \right) \wedge (2r\dv{r}) \\
+    = 2^N r^{2N-1} \left(\prod_{i=1}^{N-1} \sin(\phi_{N-i}) \cos(\phi_{N-i}) \prod_{j=1}^{N-i-1} \sin^2(\phi_j)\right) \left((\dv{\phi})^\wedge \wedge \dv{r}\right)\\
+    = 2^N r^{2N-1} \left(\prod_{i=1}^{N-1} \sin(\phi_i) \cos(\phi_i) \prod_{j=1}^{i-1} \sin^2(\phi_j)\right) \left((\dv{\phi})^\wedge \wedge \dv{r}\right)\\
+    = 2^N r^{N-1} \left(\prod_{i=1}^N z_i\right) \left(\prod_{j=1}^{N-2} \sin(\phi_j)^{N-j-1}\right) \left((\dv{\phi})^\wedge \wedge \dv{r}\right)
+\end{aligned}
+\]
+
+Equating the exterior products of each side, we have
+
+\[
+\begin{aligned}
+    2^N \left(\prod_{i=1}^N z_i \right) (\dv{z})^\wedge &= 2^N r^{N-1} \left(\prod_{i=1}^N z_i\right) \left(\prod_{j=1}^{N-2} \sin(\phi_j)^{N-j-1}\right) \left((\dv{\phi})^\wedge \wedge \dv{r}\right)\\
+    (\dv{z})^\wedge &= r^{N-1} \left(\prod_{j=1}^{N-2} \sin(\phi_j)^{N-j-1}\right) \left((\dv{\phi})^\wedge \wedge \dv{r}\right)
+\end{aligned}
+\]
+
+As a result, the Jacobian determinant is
+
+\[
+    \absdet{J_g(r, \phi)} = r^{N-1} \prod_{i=1}^{N-2} \sin(\phi_i)^{N-i-1}
+\]
+
+As a quick check, for the circle ($N=1$), $\absdet{J_g(r, \phi)} = r$, and for the sphere ($N=2$), $\absdet{J_g(r, \phi)} = r^2 \sin(\phi_1)$, which are the expected results.
+A similar result was given in \cite{muirhead2009aspects}.
+
+Comparing with the previous section, we find
+
+\[
+\begin{aligned}
+    (\dv{z})^\wedge &= r^{N-1} \left(\prod_{i=1}^{N-2} \sin(\phi_i)^{N-i-1}\right) \left((\dv{\phi})^\wedge \wedge \dv{r}\right) = r^{N-1} \left(\dv{r} \wedge (\dv{x})^\wedge \right) \\
+    &= (-1)^{N-1} r^{N-1} \left((\dv{x})^\wedge \wedge \dv{r}\right)\\
+    (\dv{x})^\wedge &= (-1)^{N-1}\left(\prod_{i=1}^{N-2} \sin(\phi_i)^{N-i-1}\right) (\dv{\phi})^\wedge
+\end{aligned},
+\]
+
+so the absolute Jacobian determinant for the bijective map from hyperspherical coordinates to the sphere is
+\[
+    \prod_{i=1}^{N-2} \sin(\phi_i)^{N-i-1}.
+\]
+
+However, we're not done yet.
+Recall that $\phi_j \in (0, \pi]$ for $i < N-1$ and $\phi_{N-1} \in (0, 2\pi]$.
+So far we've essentially changed the problem of sampling on the hypersphere in $N$ dimensions to the problem of sampling on $N-1$ intervals, or, alternatively, $N-2$ hemicircles and 1 circle.
+
+There are a few considerations when choosing the remaining parameterizations.
+First, and most importantly, we want a smooth geometry.
+Second, we want to avoid unnecessary work.
+For example, if we choose parameter expansion for these parameters, we will end up with $N-1$ extra degrees of freedom.
+Third, we want to avoid costly functions if possible.
+For example, computing $\sin$ and $\cos$ is about 3 times more costly than computing $\exp$ and about 20 times more costly than computing a square.
+
+\section{Cholesky factors of correlation matrices}
+
+A $K \times K$ correlation matrix $\Omega$ is positive definite, has a unit diagonal, and all off-diagonal entries are between -1 and 1 (exclusive).
+Because it is positive definite, it can be Cholesky factored (see \ref{cholesky}) to a $K \times K$ lower-triangular matrix $L$
+with positive diagonal elements such that $\Omega = L \cdot L^T$. Because the correlation matrix has a unit diagonal,
+
+\[
+  \Omega_{k,k} = L_k \cdot L_k^T = 1
+\]
+
+each row vector $L_k$ of the Cholesky factor is of unit length. The length and positivity
+constraint allow the diagonal elements of L to be calculated from the off-diagonal
+elements, so that a Cholesky factor for a $K \times K$ correlation matrix requires only $\binom{K}{2}$ unconstrained parameters.
+
+\subsection{Hyperbolic tangent transform}
+
+Suppose $y$ is a vector of $\binom{K}{2}$ unconstrained values. Let $z$ be a lower-triangular matrix with zero diagonal and below diagonal entries filled by row. For example, in the $3 \times 3$ case,
+\[
+z = \begin{bmatrix}
+0 & 0 & 0\\
+\tanh{y_1} & 0 & 0\\
+\tanh{y_2} & \tanh{y_3} & 0\\
+\end{bmatrix}
+\]
+
+The matrix $z$, with entries in the range $( -1, 1)$, is then transformed to the Cholesky factor $x$, by taking
+\[
+x_{i, j}=\left\{\begin{array}{lll}
+0 & \text { if } i<j & \text { [above diagonal] } \\
+\sqrt{1-\sum_{j^{\prime}<j} x_{i, j^{\prime}}^{2}} & \text { if } i=j & \text { [on diagonal] } \\
+z_{i, j} \sqrt{1-\sum_{j^{\prime}<j} x_{i, j^{\prime}}^{2}} & \text { if } i>j & \text { [below diagonal] }
+\end{array}\right.
+\]
+
+In the $3 \times 3$ case, this yields
+
+\[
+x=\left[\begin{array}{ccc}
+1 & 0 & 0 \\
+z_{2,1} & \sqrt{1-x_{2,1}^{2}} & 0 \\
+z_{3,1} & z_{3,2} \sqrt{1-x_{3,1}^{2}} & \sqrt{1-\left(x_{3,1}^{2}+x_{3,2}^{2}\right)}
+\end{array}\right]
+\]
+
+where the $z_{i,j}\in (-1, 1)$ are the tanh-transformed $y$.
+
+\subsection{Absolute Jacobian determinant of inverse transform}
+
+The Jacobian of the full transform is the product of the Jacobians of its component
+transforms. First, for the inverse transform $z = \tanh{y}$, the derivative is
+\[
+\frac{d}{dy} \tanh y = \frac{1}{(\cosh y)^{2}}
+\]
+Second, for the inverse transform of $z$ to $x$, the resulting Jacobian matrix $J$ is of
+dimension $\binom{K}{2} \times \binom{K}{2}$, with indexes $(i, j)$ for $(i > j)$. The Jacobian matrix is lower
+triangular, so that its determinant is the product of its diagonal entries, of which
+there is one for each $(i, j)$ pair.
+\[
+|\operatorname{det} J|=\prod_{i>j}\left|\frac{d}{d z_{i, j}} x_{i, j}\right|
+\]
+
+where
+
+\[
+\frac{d}{d z_{i, j}} x_{i, j}=\sqrt{1-\sum_{j^{\prime}<j} x_{i, j^{\prime}}^{2}}
+\]
+
+So the combined log Jacobian determinant
+of the complete inverse transform is given by
+
+\[
+\log |\operatorname{det} J|=-2 \sum_{n \leq(\substack{K \\ 2}} \log \cosh y+\frac{1}{2} \sum_{i>j} \log \left(1-\sum_{j^{\prime}<j} x_{i, j^{\prime}}^{2}\right)
+\]
+
+\section{Positive definite matrices}
 
@@
 
-An $N \times N$ positive definite (PD) matrix $X$ is a symmetric matrix that satisfies $v^\top X v \ge 0$ for all non-zero vectors $v$.
+An $N \times N$ positive definite (PD) matrix $X$ is a symmetric matrix that satisfies $v^\top X v > 0$ for all non-zero vectors $v$.
 Covariance matrices are examples of PD matrices.
 
-Density functions for distributions of PD matrices, such as the Wishart distribution, are generally defined with respect to the Lebesgue measure of the upper triangular elements $(\dv{X})^\wedge$.
+Density functions for distributions of PD matrices, such as the Wishart distribution, are generally defined with respect to the Lebesgue measure of the unique triangular elements $(\dv{X})^\wedge$.
 
-\subsubsection{Cholesky parameterization}
+\subsection{Cholesky parameterization}\label{cholesky}
 
-Given a positive (semi-)definite $X$ of rank $K$, we can uniquely decompose it as $X = Z^\top Z$ where $Z$ is an $N \times N$ upper triangular matrix where $Z_{ii} > 0$ for all $i \le K$ and $Z_{ij} = 0$ for all $i > K$.
-$Z$ has $\frac{K(K+1)}{2} + (N - K)K$ degrees of freedom.
+Given a positive (semi-)definite matrix $X$ of rank $K$, we can uniquely decompose it as $X = L L^\top$ where $L$ is an $N \times N$ lower triangular matrix with $L_{jj} > 0$ for all $j \le K$ and $L_{ij} = 0$ for all $j > K$.
+$L$ has $\frac{K(K+1)}{2} + (N - K)K$ degrees of freedom.
 When $K=N$, $X$ is positive-definite.
@@
 
-\[Z_{ij} = f_1(Y)_{ij} = \begin{cases} \exp(Y_{ij}) & \text{ if } i = j \le K \\ Y_{ij} & \text{otherwise} \end{cases}\]
+\[L_{ij} = f_1(Y)_{ij} = \begin{cases} \exp(Y_{ij}) & \text{ if } i = j \le K \\ Y_{ij} & \text{otherwise} \end{cases}\]
 
-The density correction for this map is $\absdet{J_{f_1}(Y)} = \exp(\sum_{i=1}^K Y_{ii})$.
+The Jacobian determinant for this map is $\absdet{J_{f_1}(Y)} = \exp(\sum_{i=1}^K Y_{ii})$.
 
-The second map is $f_2: Z \mapsto Z^\top Z = X$.
+The second map is $f_2: L \mapsto L L^\top = X$.
 To compute the Jacobian of the map, we use the half-vectorization $\vectu(X)$.
-Note that $X_{ij} = \sum_{k=1}^{\min(i, j)} Z_{ki} Z_{kj}$.
-$X_{ij}$ is thus only dependent on elements of $Z$ that are above and to the left of the $(i, j)$ position.
-As a result, the Jacobian of $\vectu \circ f_2 \circ \\vectu^{-1}$ is lower-triangular, and its determinant is only the product of its diagonal elements: $\absdet{J_{f_2}} = \prod_{1 \le i \le j \le N} \frac{\partial X_{ij}}{\partial Z_{ij}}$.
+Note that $X_{ij} = \sum_{k=1}^{\min(i, j)} L_{ik} L_{jk}$.
+$X_{ij}$ is thus only dependent on elements of $L$ that are below and to the left of the $(i, j)$ position.
+As a result, the Jacobian of $\vectu \circ f_2 \circ \vectu^{-1}$ is triangular, and its determinant is only the product of its diagonal elements: $\absdet{J_{f_2}} = \prod_{i=1}^N \prod_{j=1}^i \frac{\partial X_{ij}}{\partial L_{ij}}$.
 
-In this index range, $\frac{\partial X_{ij}}{\partial Z_{ij}} = (1 + \delta_{ij}) Z_{ii}$.
+In this index range, $\frac{\partial X_{ij}}{\partial L_{ij}} = (1 + \delta_{ij}) L_{ii}$.
 As a result
@@
 \[
-  \absdet{J_{f_2}(Z)} = 2^N \prod_{j = 1}^N \prod_{i = 1}^j Z_{ii} = 2^N \prod_{i=1}^N |Z_{ii}|^{N - i + 1} \\
+  \absdet{J_{f_2}(Z)} = 2^N \prod_{i = 1}^N \prod_{j = 1}^i L_{ii} = 2^N \prod_{i=1}^N |L_{ii}|^{N - i + 1} \\
 \]
 
-Combining the two bijective maps as $f = f_2 \circ f_1$, the combined density correction is
+Composing the two bijective maps as $f = f_2 \circ f_1$, the resulting Jacobian determinant is
 
@@
 
-\subsubsection{Matrix exponential parameterization}
+\subsection{Matrix exponential parameterization}
 
@@
 \[\Exp(Y) = U \Exp(\Lambda) U^\top,\]
-where $\Exp(\Lambda)_{ij} = e^{\lambda_i} \delta_{ij}$.
+where $\Exp(\Lambda)_{ij} = \exp(\lambda_i) \delta_{ij}$.
 
@@
 \]
-while the exterior product of the super-diagonal part is
+while the exterior product of the sub-diagonal part is
 \[
-  \bigwedge_{j=1}^N \bigwedge_{i=j+1}^N (U^\top \dv{U})_{ij} (f(\lambda_j) - f(\lambda_i)) = \left( \prod_{j=1}^N \prod_{i=1}^{j-1} (f(\lambda_j) - f(\lambda_i)) \right) \left(U^\top \dv{U} \right)^\wedge
+  \bigwedge_{i=1}^N \bigwedge_{j=i+1}^N (U^\top \dv{U})_{ij} (f(\lambda_j) - f(\lambda_i)) = \left( \prod_{i=1}^N \prod_{j=1}^{i-1} (f(\lambda_j) - f(\lambda_i)) \right) \left(U^\top \dv{U} \right)^\wedge
 \]
@@
 
-As a result, we can compute $(\dv(f(Y)))^\wedge$ by taking the exterior product of the super-diagonal and diagonal parts: 
-\[(\dv{(f(Y))})^\wedge = \left(\prod_{j=1}^N f'(\lambda_j) \prod_{i=1}^{j-1} (f(\lambda_j) - f(\lambda_i))\right) ((U^\top \dv{U})^\wedge \wedge \dv{\Lambda})\]
+As a result, we can compute $(\dv(f(Y)))^\wedge$ by taking the exterior product of the sub-diagonal and diagonal parts: 
+\[(\dv{(f(Y))})^\wedge = \left(\prod_{i=1}^N f'(\lambda_i) \prod_{j=1}^{i-1} (f(\lambda_j) - f(\lambda_i))\right) \left((U^\top \dv{U})^\wedge \wedge \dv{\Lambda}\right)\]
 
 If $f$ is the identity map $f(Y) = Y$, then
-\[ (\dv{Y})^\wedge = \left(\prod_{j=1}^N \prod_{i=1}^{j-1} (\lambda_j - \lambda_i)\right) \left((U^\top \dv{U})^\wedge \wedge \dv{\Lambda}\right)\]
+\[ (\dv{Y})^\wedge = \left(\prod_{i=1}^N \prod_{j=1}^{i-1} (\lambda_j - \lambda_i)\right) \left((U^\top \dv{U})^\wedge \wedge \dv{\Lambda}\right)\]
 
 Combining the latter two expressions:
-\[(\dv{(f(Y))})^\wedge = \left(\prod_{j=1}^N f'(\lambda_j) \prod_{i=1}^{j-1} \frac{f(\lambda_j) - f(\lambda_i)}{\lambda_j - \lambda_i}\right) (\dv{Y})^\wedge\]
+\[(\dv{(f(Y))})^\wedge = \left(\prod_{i=1}^N f'(\lambda_i) \prod_{j=1}^{i-1} \frac{f(\lambda_j) - f(\lambda_i)}{\lambda_j - \lambda_i}\right) (\dv{Y})^\wedge\]
 
@@
 For $X = f(Y) = \Exp(Y)$ then,
-\[ (\dv{X})^\wedge = \left(\prod_{j=1}^N e^{\lambda_j} \prod_{i=1}^{j-1} \frac{e^{\lambda_j} - e^{\lambda_i}}{\lambda_j - \lambda_i}\right) (\dv{Y})^\wedge,\]
+\[ (\dv{X})^\wedge = \left(\prod_{i=1}^N \exp(\lambda_i) \prod_{j=1}^{i-1} \frac{\exp(\lambda_j) - \exp(\lambda_i)}{\lambda_j - \lambda_i}\right) (\dv{Y})^\wedge,\]
 
-so the Jacobian determinant of the exponential map is $\absdet{J_{\Exp}(Y)} = \prod_{j=1}^N e^{\lambda_j} \prod_{i=1}^{j-1} \left|\frac{e^{\lambda_j} - e^{\lambda_i}}{\lambda_j - \lambda_i}\right|$.
+so the Jacobian determinant of the exponential map is $\absdet{J_{\Exp}(Y)} = \prod_{i=1}^N \exp(\lambda_i) \prod_{j=1}^{i-1} \left|\frac{\exp(\lambda_j) - \exp(\lambda_i)}{\lambda_j - \lambda_i}\right|$.
 
@@
 \[
-  \absdet{J_{\Exp}(Y)} = \prod_{j=1}^N \prod_{i=1}^j \begin{cases}
-    \left|\frac{e^{\lambda_j} - e^{\lambda_i}}{\lambda_j - \lambda_i}\right| &\text{ if } \lambda_i \ne \lambda_j \\ 
-    e^{\lambda_i} &\text{ if } \lambda_i = \lambda_j
+  \absdet{J_{\Exp}(Y)} = \prod_{i=1}^N \prod_{j=1}^i \begin{cases}
+    \left|\frac{\exp(\lambda_j) - \exp(\lambda_i)}{\lambda_j - \lambda_i}\right| &\text{ if } \lambda_i \ne \lambda_j \\ 
+    \exp(\lambda_i) &\text{ if } \lambda_i = \lambda_j
+  \end{cases}.
+\]
+
+If the eigenvalues are sorted in ascending order, then the log Jacobian determinant can be more stably computed as
+
+\[
+  \log\absdet{J_{\Exp}(Y)} = \sum_{j=1}^N j\lambda_j + \sum_{i=1}^j \begin{cases}
+    \log(1 - \exp(\lambda_i - \lambda_j)) - \log(\lambda_j - \lambda_i) &\text{ if } \lambda_i \ne \lambda_j \\ 
+    0 &\text{ if } \lambda_i = \lambda_j
   \end{cases}.
@@
 
-\subsubsection{Eigendecomposition parameterization}
+\subsection{Eigendecomposition parameterization}
 
 To define a transform directly from the symmetric eigendecomposition of $X$, we first need to note that the eigendecomposition is not unique; if we swap any two columns $i$ and $j$ of $U$ as well as swapping the eigenvalues $\lambda_i$ and $\lambda_j$, the result is the same matrix.
+Moreover, if we reverse the sign of any given eigenvector, it has no effect on the eigenvalues or $X$.
 
-When the eigenvalues are all unique, we can make the eigendecomposition also unique by enforcing the ordering of eigenvalues $\lambda_i > \lambda_j > 0$ for $j > i$.
-This is the ordering generally used for symmetric singular value decomposition.
+When the eigenvalues are all unique, we can make the eigendecomposition also unique by enforcing an ascending order of eigenvalues $0 < \lambda_i < \lambda_j$ for $j > i$ and by forcing the first entry of each eigenvector to be positive.
 
-So we can construct a bijective map with the maps $f_1$ that produces an ordered vector of positive eigenvalues and $f_2$ that produces an orthogonal matrix.
+So we can construct a bijective map with the maps $f_1$ that produces an ordered vector of positive eigenvalues and $f_2$ that produces an orthogonal matrix whose first row contains only positive entries.
 Consider $\lambda = f_1(y_1)$ and $U = f_2(y_2)$.
@@
   \lambda_i = \begin{cases}
-    e^{(y_1)_N} &\text{ if } i = N\\
-    \lambda_{i-1} + e^{(y_1)_i} &\text{ if } 1 \le i < N
+   \exp\bigl((y_1)_1\bigr) &\text{ if } i = 1\\
+    \lambda_{i-1} + \exp\bigl((y_1)_i\bigr) &\text{ if } 1 < i \le N
   \end{cases}
 \]
-The Jacobian determinant of this map is $\absdet{J_{f_1}(y_1)} = \exp\left(\sum_{i=1}^N (y_1)_i\right)$.
+The Jacobian of this map is triangular, so its determinant is $\absdet{J_{f_1}(y_1)} = \exp\left(\sum_{i=1}^N (y_1)_i\right)$.
 
-We can then choose for $f_2$ any of the transforms to the orthogonal matrices previously discussed.
+We can then choose for $f_2$ any of the transforms to the orthogonal matrices previously discussed, though care must be taken to avoid multimodality.
+For example, if the transform covers the entire orthogonal group, then $2^N$ matrices map to the same orthogonal matrix with positive first row, which will manifest as $2^N$ modes in the unconstrained distribution.
+If, however, we can modify one of the transforms to natively produce only matrices with this constraint, then we can eliminate these undesirable modes.
+
 The combined Jacobian determinant of $f: (y_1, y_2) \mapsto f_2(y_2) \operatorname{diag}(f_1(y_1)) f_2(y_2)^\top = X$ is then 

### 2022-07-22 ess_esp.py

@@ -1,0 +1,35 @@
+
+import sys
+sys.path.insert(1, 'utils')
+from ess import get_ess_leapfrog_ratio
+
+import pickle
+import numpy as np
+import matplotlib.pyplot as plt
+import arviz as az
+
+
+
+parameters = [{'alpha':[0.1]*10, 'N':10}, {'alpha':[0.1]*100, 'N':100},
+               {'alpha':[1]*10, 'N':10}, {'alpha':[1]*100, 'N':100},
+               {'alpha':[10]*10, 'N':10}, {'alpha':[10]*100, 'N':100}]
+
+transforms = ['stickbreaking', 'softmax', 'softmax-augmented', 'stan']
+transform_category='simplex'
+evaluating_model='dirichlet_symmetric'
+
+var_name='x'
+var_dim=0
+
+plt.rcParams["figure.figsize"] = (20,10)
+fig, axes = plt.subplots(2,3)
+for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
+    for transform in transforms:
+        x, y = get_ess_leapfrog_ratio(transform_category, transform, evaluating_model, params, var_name, var_dim, n_repeat=100)
+        ax.plot(x,y, label=transform)
+    ax.set_title(f'alpha = {params["alpha"][0]}, N = {params["N"]}')
+ax.axes.yaxis.set_ticklabels([])
+fig.supxlabel('ESS/Leapfrog')
+fig.supylabel('Density')
+plt.legend()
+plt.savefig('figures/ess.png', dpi=300)

### 2022-07-22 plot.py

@@ -1,28 +1,43 @@
-
 import os
 os.chdir('transforms')
 
 import sys
 sys.path.insert(1, 'utils')
 from sample import sample
-
+from rmse import rmse_leapfrog
 
 import numpy as np
 import matplotlib.pyplot as plt
 import arviz as az
 
-from dirichlet_symmetric import get_dirichlet_symmetric_rmse
-from rhat import calc_rhat_mixed_chains
-from ess import get_ess_leapfrog_ratio
-
 transform_category='simplex'
 evaluating_model='dirichlet_symmetric'
 
-parameters = [{'alpha': [0.1]*10, 'N': 10}, {'alpha': [0.1]*100, 'N': 100}, {'alpha': [0.1]*1000, 'N': 1000},
-{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000},
-{'alpha': [1]*10, 'N': 10}, {'alpha': [1]*100, 'N': 100}, {'alpha': [1]*1000, 'N': 1000}]
+transforms = ['stan', 'softmax-augmented','stickbreaking', 'softmax']
+parameters = [{'alpha':[0.1]*10, 'N':10}, {'alpha':[0.1]*100, 'N':100},
+               {'alpha':[1]*10, 'N':10}, {'alpha':[1]*100, 'N':100},
+               {'alpha':[10]*10, 'N':10}, {'alpha':[10]*100, 'N':100}]
 
-transforms = ['softmax', 'softmax-augmented','stickbreaking', 'stan']
+plt.rcParams["figure.figsize"] = [20,10]
+fig, axes = plt.subplots(2,3)
+fig.supxlabel('Cumulative Leapfrog Steps')
+fig.supylabel('Root Mean Squared Error')
 
-get_dirichlet_symmetric_rmse(transforms, transform_category, 
-                             parameters, fig_name='rmse_example.png')
+for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
+    for transform in transforms:
+        idata = sample(transform_category=transform_category, transform=transform,
+            evaluating_model='dirichlet_symmetric', parameters=[params],
+            auto_eval_all_params=False, n_iter = 1000,  n_chains = 4, n_repeat=100,
+                                    show_progress = True, resample=False,return_idata=True)
+        alpha = params['alpha']
+        N = params['N']
[…]

### 2022-07-22 simplex.py

@@ -1,7 +1,7 @@
 import os
 from tqdm import tqdm
 os.chdir('transforms')
-
+print(os.getcwd())
 import sys
 sys.path.insert(1, 'utils')
 
@@ -13,7 +13,7 @@ parser.add_argument('--parameters', type=int, required=True)
 parser.add_argument('--transform', type=str, required=True)
 args = parser.parse_args()
 
-with open(f"target_densities/param_map_dirichlet_symmetric.json", "rb") as f:
+with open(f"target_densities/param_map_dirichlet_symmetric.pkl", "rb") as f:
     param_map = pickle.load(f)
 
 alpha = param_map[args.parameters]
@@ -21,7 +21,6 @@ N = len(alpha)
 
 sample(transform_category='simplex', transform=args.transform, evaluating_model='dirichlet_symmetric', 
 parameters={'alpha':alpha, 'N':N}, 
-            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
+            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat=100,
                             show_progress = True, resample=True, return_idata=False)
-
-
+                            

### 2022-07-22 utils/dirichlet_symmetric.py

@@ -4,27 +4,26 @@ import matplotlib.pyplot as plt
 from rmse import rmse_leapfrog
 from sample import sample
 
-def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name, n_repeat=1, n_iter=1000, 
-                                n_chains=4, show_progress=True, resample=True):
+def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name,  n_iter=1000, 
+                                n_chains=4, n_repeat=1, show_progress=True, resample=False, output_dir='/mnt/sdceph/users/mjhajaria/'):
     
     plt.rcParams["figure.figsize"] = [20,10]
     plt.rcParams['figure.dpi'] = 300
-    fig, axes = plt.subplots(3,3)
+    fig, ax = plt.subplots(1)
     fig.supxlabel('Cumulative Leapfrog Steps')
     fig.supylabel('Root Mean Squared Error')
 
-    for ax, params in zip(axes.flatten() if len(parameters)>1 else [axes],  parameters):
-        for transform in transforms:            
-            idata = sample(transform_category=transform_category, transform=transform, 
-                evaluating_model='dirichlet_symmetric', parameters=[params], 
-                auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
-                                        show_progress = show_progress, resample=resample)
-            alpha = params['alpha']
-            N = params['N']
-            true_x = [a/sum(alpha) for a in alpha]
-            x, y = rmse_leapfrog(idata=idata, true_var=true_x, var_name='x', var_dim=0)
-            ax.plot(x,y, label = str(transform))
-
+    for transform in transforms:            
+        idata = sample(transform_category=transform_category, transform=transform, 
+            evaluating_model='dirichlet_symmetric', parameters=parameters, 
+            auto_eval_all_params=False, n_iter = n_iter,  n_chains = n_chains, n_repeat=n_repeat,
+                                    show_progress = show_progress, resample=resample, output_dir=output_dir,return_idata=True)
+        alpha = parameters[0]['alpha']
+        N = parameters[0]['N']
+        true_x = [a/sum(alpha) for a in alpha]
[…]

### 2022-07-22 utils/ess.py

@@ -4,7 +4,10 @@ import matplotlib.pyplot as plt
 from cmdstanpy import CmdStanModel
 from tqdm import tqdm
 import arviz as az
-
+import sys
+sys.path.insert(1, 'utils')
+from sample import sample
+import pickle
 
 def get_ess_leapfrog_ratio(
     transform_category,
@@ -13,25 +16,29 @@ def get_ess_leapfrog_ratio(
     params,
     var_name,
     var_dim,
-    repeat=100,
-    return_ess=False,
+    n_repeat=100,
 ):
-    model = CmdStanModel(
-        stan_file=f"stan_models/{transform}_{evaluating_model}.stan",
-        cpp_options={"STAN_THREADS": "true"},
-    )
     x = []
-    for i in tqdm(range(repeat)):
-        idata = az.from_cmdstanpy(model.sample(data=params, show_progress=False))
-        data = idata.sel(var_names=[str(var_name)], x_dim_0=var_dim)
-        ess = az.ess(data, var_names=[str(var_name)], method="bulk")[
-            str(var_name)
-        ].values
-        leapfrog = idata.sample_stats["n_steps"].sum().values
-        x.append(ess / leapfrog)
-    if return_ess == True:
-        return x
-    else:
-        kde = gaussian_kde(x)
-        dist_space = np.linspace(min(x), max(x), 1000)
-        return dist_space, kde(dist_space)
+    idata = sample(
+        transform_category=transform_category,
+        transform=transform,
+        evaluating_model=evaluating_model,
+        parameters=[params],
+        output_dir='/mnt/sdceph/users/mjhajaria',
+        auto_eval_all_params=False,
+        n_iter=1000,
+        n_chains=4,
+        n_repeat=100,
+        show_progress=True,
+        resample=False,
+        return_idata=True)
+    
+    with open(f"target_densities/param_map_{evaluating_model}.pkl", "rb") as f:
+        param_map = pickle.load(f)
+    ess = np.loadtxt(open(f'/mnt/sdceph/users/mjhajaria/sampling_results/{transform_category}/{transform}/{evaluating_model}/ess_{param_map[tuple(list(params.values())[0])]}_{n_repeat}.csv'),delimiter = ",")
+    leapfrog = np.average(idata.sample_stats['n_steps'].sum(axis=1).values.reshape(-1, 4), axis=1)
+    x=np.divide(ess, leapfrog)
+    kde = gaussian_kde(x)
[…]

### 2022-07-23 simplex.py
script fix
script fix

sample.py fix

xyz

-

-

final rmse plots

ess plot

m

--

plot

plot

commit

commit

commit

-

ess plot fix

-

df

space

@@ -1,7 +1,7 @@
 import os
 from tqdm import tqdm
 os.chdir('transforms')
-
+print(os.getcwd())
 import sys
 sys.path.insert(1, 'utils')
 
@@ -21,6 +21,6 @@ N = len(alpha)
 
 sample(transform_category='simplex', transform=args.transform, evaluating_model='dirichlet_symmetric', 
 parameters={'alpha':alpha, 'N':N}, 
-            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat = 100, 
+            auto_eval_all_params=False, n_iter = 1000, n_chains = 4, n_repeat=100,
                             show_progress = True, resample=True, return_idata=False)
                             

### 2022-07-23 utils/sample.py

@@ -32,7 +32,7 @@ def sample(
 ):
     """
     Sample from the given dictionary containing the models, transform_categories, and parameters.
-    It saves the results in a json file which is named with n_repeat, n_chains and n_iter.
+    It saves the results in a json file which is named with n_chains.
 
     Parameters
     ----------
@@ -55,14 +55,14 @@ def sample(
     auto_eval_all_params: bool
         If True, all parameterizations are evaluated.
 
-    n_repeat: int
-        Number of times to repeat the sampling.
-
     n_iter : int
         Number of samples to be drawn.
 
     n_chains : int
         Number of chains to be drawn.
+    
+    n_repeat : int 
+        Number of runs to be drawn.
 
     show_progress: bool (default = True)
         Whether to show progress bar.
@@ -79,7 +79,7 @@ def sample(
     Note
     ----
     The directory structure for storing stan files is: stan_models/{transform}_{evaluating_model}.stan
-    The directory structure for storing the results is: sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_mapping}.json
+    The directory structure for storing the results is: sampling_results/{transform_category}/{transform}/{evaluating_model}/{param_mapping}_{n_repeat}.json
     """
     with open(f"target_densities/param_map_{evaluating_model}.pkl", "rb") as f:
         param_map = pickle.load(f)
@@ -108,6 +108,7 @@ def sample(
             model = CmdStanModel(
                 stan_file=stan_filename, cpp_options={"STAN_THREADS": "true"}
             )
+            ess=[]
             idata = az.from_cmdstanpy(
                 model.sample(
                     data=params,
@@ -116,7 +117,7 @@ def sample(
                     chains=n_chains,
                 )
             )
-	    
+            ess.append(az.ess(idata)['x'][0])
             for i in tqdm(range(n_repeat - 1)):
                 fit = model.sample(
                     data=params,
@@ -125,14 +126,16 @@ def sample(
                     chains=n_chains,
                     seed=i
                 )
-                idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
+                ess.append(az.ess(az.from_cmdstanpy(fit))['x'][0])
 
+        
+                idata = az.concat(idata, az.from_cmdstanpy(fit), dim="chain")
[…]

### 2022-07-23 plot.py
plots and other fixes
* syncing stuff

* whitespace adjustment

* plot file fix

* stuff

* indent shit

* stuff

* indent shit

* indent

* add figure

* i hate indents

* whitespace

* -

* more changes

* fixmylife

* minor things

* fix bugs again

* add plots

* slight changes

* --

* add time

* replot

* plotmylife

* Revert "plotmylife"

This reverts commit bde7b392ee8ddcc5837802f7b43cb8580b74b646.

* script fix

script fix

sample.py fix

xyz

-

-

final rmse plots

ess plot

m

--

plot

plot

commit

commit

commit

-

ess plot fix

-

df

space

@@ -33,5 +33,5 @@ evaluating_model='dirichlet_symmetric'
 transforms = ['stan', 'softmax-augmented','stickbreaking', 'softmax']
 
 get_dirichlet_symmetric_rmse(transforms, transform_category, 
-                             parameters=[{'alpha':alpha, 'N':N}],resample=False, fig_name=f'rmse_{args.parameters}.png', n_repeat=100)
+                             parameters=[{'alpha':alpha, 'N':N}], fig_name=f'rmse_{args.parameters}.png', n_repeat=100)
 

### 2022-07-23 utils/dirichlet_symmetric.py

@@ -4,8 +4,8 @@ import matplotlib.pyplot as plt
 from rmse import rmse_leapfrog
 from sample import sample
 
-def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name, n_repeat=1, n_iter=1000, 
-                                n_chains=4, show_progress=True, resample=True):
+def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig_name,  n_iter=1000, 
+                                n_chains=4, n_repeat=1, show_progress=True, resample=False, output_dir='/mnt/sdceph/users/mjhajaria/'):
     
     plt.rcParams["figure.figsize"] = [20,10]
     plt.rcParams['figure.dpi'] = 300
@@ -16,8 +16,8 @@ def get_dirichlet_symmetric_rmse(transforms, transform_category, parameters, fig
     for transform in transforms:            
         idata = sample(transform_category=transform_category, transform=transform, 
             evaluating_model='dirichlet_symmetric', parameters=parameters, 
-            auto_eval_all_params=False, n_iter = n_iter, n_chains = n_chains, n_repeat = n_repeat, 
-                                    show_progress = show_progress, resample=resample)
+            auto_eval_all_params=False, n_iter = n_iter,  n_chains = n_chains, n_repeat=n_repeat,
+                                    show_progress = show_progress, resample=resample, output_dir=output_dir,return_idata=True)
         alpha = parameters[0]['alpha']
         N = parameters[0]['N']
         true_x = [a/sum(alpha) for a in alpha]

### 2022-07-23 utils/ess.py

@@ -4,7 +4,9 @@ import matplotlib.pyplot as plt
 from cmdstanpy import CmdStanModel
 from tqdm import tqdm
 import arviz as az
-
+import sys
+sys.path.insert(1, 'utils')
+from sample import sample
 
 def get_ess_leapfrog_ratio(
     transform_category,
@@ -13,25 +15,29 @@ def get_ess_leapfrog_ratio(
     params,
     var_name,
     var_dim,
-    repeat=100,
-    return_ess=False,
+    n_repeat=100,
 ):
-    model = CmdStanModel(
-        stan_file=f"stan_models/{transform}_{evaluating_model}.stan",
-        cpp_options={"STAN_THREADS": "true"},
-    )
     x = []
-    for i in tqdm(range(repeat)):
-        idata = az.from_cmdstanpy(model.sample(data=params, show_progress=False))
-        data = idata.sel(var_names=[str(var_name)], x_dim_0=var_dim)
-        ess = az.ess(data, var_names=[str(var_name)], method="bulk")[
-            str(var_name)
-        ].values
-        leapfrog = idata.sample_stats["n_steps"].sum().values
+    idata = sample(
+        transform_category=transform_category,
+        transform=transform,
+        evaluating_model=evaluating_model,
+        parameters=[params],
+        output_dir='/mnt/sdceph/users/mjhajaria',
+        auto_eval_all_params=False,
+        n_iter=1000,
+        n_chains=4,
+        show_progress=True,
+        resample=False,
+        return_idata=True)
+    for i in tqdm(range(n_repeat)):
+        x_data = idata.sel(chains=[i],var_names=[str(var_name)], x_dim_0=var_dim)
+        ess = az.ess(x_data)['x'].value
+        print(ess)
+        leapfrog = x_data.sample_stats["n_steps"].sum().values
         x.append(ess / leapfrog)
-    if return_ess == True:
-        return x
-    else:
-        kde = gaussian_kde(x)
-        dist_space = np.linspace(min(x), max(x), 1000)
-        return dist_space, kde(dist_space)
+    print(x)
+    kde = gaussian_kde(x)
+    dist_space = np.linspace(min(x), max(x), 1000)
+    print(dist_space)
+    return dist_space, kde(dist_space)
